---
title: "Step3_3_AlterationPerChromosome"
author: "Melissa"
date: "04/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
mainDir <- here::here() # the MainDir should be the root of Tumor_BRCA_P53:
source(knitr::purl(file.path(mainDir,"Scripts","global_var_current.Rmd"), quiet=TRUE))
source(knitr::purl(file.path(mainDir,"Scripts","scRNAseq_Functions.Rmd"), quiet=TRUE))
source(knitr::purl(file.path(mainDir,"Scripts","global_variables.Rmd"), quiet=TRUE))
```

#### KIP ALL THIS ¨PART TO GET THE BOXPLOT OF ALTERED GENOME ###################
#-----------------------------------------------------------------------------#


Barplot of altered chromosomes overall:
per sample : binarized or counted per chromosome:

```{r}
tlg= qs::qread(paste0(oaltDir, "/Binarized_CNT_With_Basal.qs"))
tlg2 <- qs::qread(paste0(oaltDir, "/Binarized_ST_WOBasal_SameThrasAll.qs"))
tlg3 <- qs::qread(paste0(oaltDir, "/Binarized_MT_WOBasal.qs"))
tlg4 <- qs::qread(paste0(oaltDir, "/Binarized_LT_WOBasal.qs"))
```


Split the controls per subtype:
```{r}
lp= tlg[,which(colnames(tlg) %in% rownames(cnt@meta.data)[which(cnt$subtype== "LP")]) ]
p16= tlg[,which(colnames(tlg) %in% rownames(cnt@meta.data)[which(cnt$subtype== "P16+ Pre-lesional")]) ]
avd=tlg[,which(colnames(tlg) %in% rownames(cnt@meta.data)[which(cnt$subtype== "Avd")]) ]
```
########
Add the region information FOR THE ST
```{r}
gene_pos <- as.data.frame(ic_st@gene_order)
colnames(gene_pos) <- c("chromosome","start","end")
gene_pos$chromosome =  paste0("chr",gene_pos$chromosome)
gg <- GenomicRanges::makeGRangesFromDataFrame(gene_pos)
cytoband2= read.table(file= "/cytoBand.txt", header = F)
colnames(cytoband2) <- c("chromosome", "start","stop","cytoband", "g")
## transform them to a GRanges:
library(GenomicRanges)
gcyto <- GenomicRanges::makeGRangesFromDataFrame(cytoband2)
tmp <- GenomicRanges::findOverlaps(query = gg, subject= gcyto, minoverlap = 100, type="within")
### get the correspondance overlaps between the cytobands and the gene coordinates:
gp <- gene_pos[queryHits(tmp),] 

cr <- cytoband2[subjectHits(tmp),]
##
gp$cytoband = cr$cytoband
gp$chr = paste0(gp$chromosome,"_", gp$cytoband)
rm(tmp)
rm(gg)
rm(cr)
gp$gene= rownames(gp)
gp$region=paste0(gp$chr, "_", gp$cytoband)
```

```{r}
all(rownames(avd)== rownames(ic_cnt@gene_order))
lp$chr=ic_cnt@gene_order$chr
avd$chr=ic_cnt@gene_order$chr
p16$chr=ic_cnt@gene_order$chr
```


```{r}
tlg2=tlg2[,c(1:1436)]
tlg4=tlg4[,c(1:3510)]
tlg3=tlg3[,c(1:4082)]
```

Apply per sample the aggregation per chromsome:
```{r}
chlt=aggregate(tlg4[,-1], by=list(tlg4$chromosome), function(u) {table(factor(u, levels=1))}) %>% as.data.frame()
chlt <-  chlt %>% column_to_rownames("Group.1")
#######
chm=aggregate(tlg3[,-1], by=list(tlg3$chromosome), function(u) {table(factor(u, levels=1))}) %>% as.data.frame()
chm <-  chm %>% column_to_rownames("Group.1")
####
chs=aggregate(tlg2[,-1], by=list(tlg2$chromosome), function(u) {table(factor(u, levels=1))}) %>% as.data.frame()
chs <-  chs %>% column_to_rownames("Group.1")
####
cha=aggregate(avd[,-1], by=list(avd$chr), function(u) {table(factor(u, levels=1))}) %>% as.data.frame()
cha <-  cha %>% column_to_rownames("Group.1")
###
chl=aggregate(lp[,-1], by=list(lp$chr), function(u) {table(factor(u, levels=1))}) %>% as.data.frame()
chl <-  chl %>% column_to_rownames("Group.1")
###
chp=aggregate(p16[,-1], by=list(p16$chr), function(u) {table(factor(u, levels=1))}) %>% as.data.frame()
chp <-  chp %>% column_to_rownames("Group.1")
```

Divide per the number of genes per chromosome:
```{r}
m <- table(avd$chr) %>% as.data.frame()
m=m[-1,]
##############
cha=cha[-1,]
chp=chp[-1,]
chl=chl[-1,]
###
cha$chr= rownames(cha)
chp$chr= rownames(chp)
chl$chr= rownames(chl)
chlt$chr= rownames(chlt)
chm$chr= rownames(chm)
chs$chr= rownames(chs)
######################
chlt$chr= as.character(gsub("chr","", chlt$chr))
chs$chr= as.character(gsub("chr","", chs$chr))
chm$chr=as.character( gsub("chr","", chm$chr))
#######################
cha=cha[match(m$Var1, cha$chr),]
chl=chl[match(m$Var1, chl$chr),]
chp=chp[match(m$Var1, chp$chr),]
##
chlt=chlt[match(m$Var1, chlt$chr),]
chm=chm[match(m$Var1, chm$chr),]
chs=chs[match(m$Var1, chs$chr),]
######################
stopifnot(m$Var1== chm$chr)
stopifnot(m$Var1== chlt$chr)
#####
cha <- round(apply(cha[,-201], 2, function(n){ n/m$Freq}  ),3) %>%t() %>% as.data.frame() 
chl <- round(apply(chl[,-244], 2, function(n){ n/m$Freq}  ),3) %>%t() %>% as.data.frame() 
chp <- round(apply(chp[,-152], 2, function(n){ n/m$Freq}  ),3) %>%t() %>% as.data.frame() 
##
chs <- round(apply(chs[,-1436], 2, function(n){ n/m$Freq}  ),3) %>%t() %>% as.data.frame() 

chm <- round(apply(chm[,-4082], 2, function(n){ n/m$Freq}  ),3) %>%t() %>% as.data.frame() 
chlt <- round(apply(chlt[,-3510], 2, function(n){ n/m$Freq}  ),3) %>%t() %>% as.data.frame() 
```

Combine all the alterations per chromosome:
```{r}
altchr= rbind(cha, chl, chp)
altchr2= rbind(chs, chm, chlt)
colnames(altchr)= paste0("chr", colnames(altchr))
altchr= rbind(altchr, altchr2)
qs::qsave(altchr, paste0(oaltDir, "/Matrix_PercentAlt_ChromosomePerCell.qs"))
```

##------------------------ START HERE ---------------------------------------------###
```{r}
altchr <- qs::qread(paste0(oaltDir, "/Matrix_PercentAlt_ChromosomePerCell.qs"))
```


Get the distribution :
```{r}
names(colgene)=paste0("chr", names(colgene))
p <- altchr %>% melt()  %>% ggplot(aes(y=value,  x=variable, fill=variable)) + geom_boxplot(alpha=0.7,show.legend = F) + scale_fill_manual(values=colgene[-1]) +themplot+ gtheme + xlab("chromosome") + ylab("% Alterations") + coord_flip()
##
ggsave(filename =paste0(altDir, "/BoxPlot_AlteredChr_AllTumors.pdf"), device = "pdf", plot = p, width = 300, height = 300, dpi = 500, units = "mm" )
```


