---
title: "Step2_Pseudotime_2_PAGA_Python_R"
author: "Melissa"
date: "6/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
mainDir <- here::here()
source(knitr::purl(file.path(mainDir,"Scripts","global_var_current.Rmd"), quiet=TRUE))
source(knitr::purl(file.path(mainDir,"Scripts","scRNAseq_Functions.Rmd"), quiet=TRUE))
source(knitr::purl(file.path(mainDir,"Scripts","global_variables.Rmd"), quiet=TRUE))
```


```{r}
epith= qs::qread(paste0(RDatadir,"/Epithelial_FinalAnnotation.qs"))
```


##------------------SKIP THIS PART ---------------------------------------------------#######
#--------------------------------------------------------------------------------------###
#------------------------ LOAD THE SEURAT OBJECT ON R --------------------------------#

Prepare the data for PAGA:
```{r}
seu= subset(epith, idents= c("LP","Avd","P16+ Pre-lesional", "Mki67+ ST","Cytl1+ ST","Trf+ ST","Spp1+ ST")) 

seu$subtype <- factor(seu$subtype, levels=c("Avd","LP","P16+ Pre-lesional","Mki67+ ST","Cytl1+ ST","Trf+ ST","Spp1+ ST"))
qs::qsave(seu, paste0(RDatadir, "/EpithelialSubset_forPseudotime.qs"))
```

Convert the seurat object to anndata object to be used by PAGA

The Object is stored just in case:
```{r}
seu= qs::qread(paste0(RDatadir, "/EpithelialSubset_forPseudotime.qs"))

sceasy::convertFormat(seu, from="seurat", to="anndata",
 outFile=paste0(RDatadir,"/EpithelialSubset_forPseudotime_Anndata2_WOLHS.h5ad"))
```


##NOW:

```{r}
s1= subset(epith, idents= c("LP" ,"Hist1h2ap+ T_1","Avd","p16_cycling","Emp1+ T_1", "Apod+ T_1",  "Mgp+ T_1","Isg15+ GCreP_L_7412","Luminal H-S"))

emb=Embeddings(s1, reduction = "umap") %>% as.data.frame()
emb= emb %>% dplyr::filter(., UMAP_2 < (3.5) & UMAP_1 > (-1))
#emb= emb %>% dplyr::filter(., UMAP_2 > (-5) & UMAP_2 < 5 & UMAP_1 < (-3))
s1= subset(s1, cells= rownames(emb))
```

```{r}
epith=qs::qread(file.path(RDatadir, "Epithelial_Only_KeepF_NoCreN_SubtypeAnnotated.qs"))

Idents(epith)="seurat_clusters"
epith= FindSubCluster(epith, cluster = 7, resolution = 0.25,  graph.name = "SCT_snn")
Idents(epith)="sub.cluster"
seu=subset(epith, idents=c(  "7_0", "7_1", "7_2", "11", "2", "9", "14", "3", "4", "8" )) #"Luminal H-S",

DefaultAssay(seu)="SCT"
seu= seu %>% RunPCA(., npcs = 60)  %>% RunUMAP(., reduction= "pca", dims= 1:40 ) %>% FindNeighbors(., reduction = "pca", dims = 1:40) %>% FindClusters(., resolution = 1.0,group.singletons = TRUE)

sceasy::convertFormat(seu, from="seurat", to="anndata",
 outFile=paste0(RDatadir2,"/EpithelialSubset_forPseudotime_Anndata2_WOLHS.h5ad"))
```
```{r}
library(reticulate)
scipy <- import("scipy")
```

#------------------------ RUN PAGA ON PYTHON --------------------------------#

Now, select a python chunk and load the needed libraries
```{python}
import scipy as sc
import numpy as np
import matplotlib.pyplot as pl
import os
import matplotlib as mb
import pandas as pd
sc.logging.print_versions()
```

Read the transformed ANNDATA object using scanpy
```{python}
import scanpy as sc
adata= sc.read_h5ad(paste0(RDatadir,"/EpithelialSubset_forPseudotime_Anndata2_WOLHS.h5ad"))
adata
```

Perform clustering 
```{python}
adata.obs.head()
sc.pp.neighbors(adata,n_neighbors=15, n_pcs=40)
```

```{python}
sc.tl.paga(adata, groups='subtype')
adata.obs.head()
```

Make sure the annotation is correct by plottinh the expression levels of Elf5, Cdkn2a and others :
```{python}
sc.pl.paga(adata, color=[ "Elf5","Cdkn2a",], edge_width_scale=0.5, threshold=0.15,labels= None)
```

Run UMAP on the graph based object:
```{python}
sc.pl.umap(adata, color='subtype',  title='', frameon=False,edges=True, save= 'PAGA_UMAP_Edges_RMARK.pdf', legend_loc='right')
```

RUN PAGA, and set 0.15 a threshold to keep edges between nodes (cell types)
```{python}
sc.pl.paga(adata, threshold=0.15, edge_width_scale=1, layout='fr', random_state=1)
```



#------------------- LOAD PAGA RESULTS ON R TO MANUALLY DRAW THE PAGA GRAPH AND PERFORM STATS --------------#

Need to have reticulate installed to be able to load python on r

Creation of PAGA object on r:
```{r}
a=py$adata$uns$paga$connectivities
## get PAGA POSITIONS AS WELL AS UMAP COORDINATES
pos = py$adata$uns$paga$pos %>% data.frame()
umap= py$adata$obsm["X_umap"] %>% data.frame()
```



```{r}
paga <- list(
  connectivities = py$adata$uns$paga$connectivities%>% reticulate::py_to_r() %>% data.frame(),
  connectivities_tree =py$adata$uns$paga$connectivities%>% reticulate::py_to_r() %>% data.frame(),
  group_name = levels(py$adata$obs$subtype),
  groups = levels(py$adata$obs$subtype),

  group_colors = cols, #cols are the named vector of the epithelial subtype
  position = tibble(
    group = levels(py$adata$obs$subtype),
    x = pos$X1,
    y = pos$X2
  ),
  umap = tibble(
    UMAP_1 = umap$V1,
    UMAP_2 = umap$V2
  )
)
###
rownames(paga$connectivities) <- c(1:nrow(paga$pos))
colnames(paga$connectivities) <- c(1:nrow(paga$pos))
seu@misc$paga <- paga
```


Save paga object:
```{r}
qs::qsave(paga, paste0(RDatadir, "/PAGA_object_WLHS.qs"))
```

###------------------- START HERE ----------------------------------##
V2:
```{r}
paga_connectivities <- read.csv("C:/Users/Melissa S/Desktop/PROJECTS/BRCA1_Tumorigenesis/REVIEW/output/mouse/invivo/RData/paga_connectivities.csv", header=F)

paga_connectivities=paga_connectivities[-1,]
paga_connectivities=paga_connectivities[,-1]
rownames(paga_connectivities)=seq(1, 11)
colnames(paga_connectivities)=seq(1, 11)

paga_pos <- read.csv("C:/Users/Melissa S/Desktop/PROJECTS/BRCA1_Tumorigenesis/REVIEW/output/mouse/invivo/RData/paga_pos.csv", header=FALSE)

paga_pos=paga_pos[-1,]
rownames(paga_pos)=seq(1, 11)
paga_pos=paga_pos[,-1]

paga_umap <- read.csv("C:/Users/Melissa S/Desktop/PROJECTS/BRCA1_Tumorigenesis/REVIEW/output/mouse/invivo/RData/paga_umap.csv", header=FALSE)

paga_umap=paga_umap[-1,]
rownames(paga_umap)=paga_umap$V1
paga_umap=paga_umap[,-1]
###############################################################################
col=c(brewer.pal(n = 8, name = "Dark2")) #brewer.pal(n = 11, name = "Paired"), 

paga <- list(
  connectivities = paga_connectivities,
  connectivities=paga_connectivities,
  group_name = levels(seu$subtype),
  groups = levels(seu$subtype),

  group_colors = col , #cols, #cols are the named vector of the epithelial subtype
  
  position = tibble(
    group = levels(seu$subtype)[1:11],
    x = paga_pos$V2,
    y = paga_pos$V3
  ),
  umap = tibble(
    UMAP_1 = paga_umap$V2,
    UMAP_2 = paga_umap$V3
  )
)
###
rownames(paga$connectivities) <- c(1:nrow(paga_pos))
colnames(paga$connectivities) <- c(1:nrow(paga_pos))
seu@misc$paga <- paga
####################################
paga_edges <- tibble(
  group1 = rownames(paga$connectivities)[row(paga$connectivities)[upper.tri(paga$connectivities)]],
  group2 = colnames(paga$connectivities)[col(paga$connectivities)[upper.tri(paga$connectivities)]],
  weight = paga$connectivities[upper.tri(paga$connectivities)]
)  %>%
mutate(
  x1 = paga$position$x[match(.$group1, rownames(paga$position))],
  y1 = paga$position$y[match(.$group1, rownames(paga$position))],
  x2 = paga$position$x[match(.$group2, rownames(paga$position))],
  y2 = paga$position$y[match(.$group2, rownames(paga$position))]
) %>%
dplyr::filter(weight >=0.2)


#################################

dpe= as.data.frame(paga_edges)

dpe$group1= dplyr::recode(dpe$group1, "1"= levels(seu$subtype)[1],"2"= levels(seu$subtype)[2],"3"= levels(seu$subtype)[3],"4"= levels(seu$subtype)[4],"5"= levels(seu$subtype)[5],"6"= levels(seu$subtype)[6],"7"= levels(seu$subtype)[7],"8"= levels(seu$subtype)[8], "9"= levels(seu$subtype)[9], "10"= levels(seu$subtype)[10],"11"= levels(seu$subtype)[11])
##
dpe$group2= dplyr::recode(dpe$group2, "1"= levels(seu$subtype)[1],"2"= levels(seu$subtype)[2],"3"= levels(seu$subtype)[3],"4"= levels(seu$subtype)[4],"5"= levels(seu$subtype)[5],"6"= levels(seu$subtype)[6],"7"= levels(seu$subtype)[7],"8"= levels(seu$subtype)[8], "9"= levels(seu$subtype)[9], "10"= levels(seu$subtype)[10],"11"= levels(seu$subtype)[11])
```


```{r}
 epith_cols=qs::qread("C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Backup/inVivo/Seurat/output/Epith_Cols_Subtype.qs")
col=epith_cols
paga <- qs::qread(paste0(RDatadir, "/PAGA_object_WLHS.qs"))
paga$group_colors= col[paga$groups]
```

Create a table of the connectivities and keep only those higher than 0.15
```{r}

paga_edges <- tibble(
  group1 = rownames(paga$connectivities)[row(paga$connectivities)[upper.tri(paga$connectivities)]],
  group2 = colnames(paga$connectivities)[col(paga$connectivities)[upper.tri(paga$connectivities)]],
  weight = paga$connectivities[upper.tri(paga$connectivities)]
) 

paga_edges= paga_edges %>%
mutate(
  x1 = paga$position$x[match(.$group1, rownames(paga$position))],
  y1 = paga$position$y[match(.$group1, rownames(paga$position))],
  x2 = paga$position$x[match(.$group2, rownames(paga$position))],
  y2 = paga$position$y[match(.$group2, rownames(paga$position))]
) %>%
dplyr::filter(weight >=0.2)
```

Prepare the dataframe of edges weights 

```{r}
dpe= as.data.frame(paga_edges)

dpe$group1= dplyr::recode(dpe$group1, "1"= levels(seu$subtype)[1],"2"= levels(seu$subtype)[2],"3"= levels(seu$subtype)[3],"4"= levels(seu$subtype)[4],"5"= levels(seu$subtype)[5],"6"= levels(seu$subtype)[6],"7"= levels(seu$subtype)[7],"8"= levels(seu$subtype)[8])
##
dpe$group2= dplyr::recode(dpe$group2, "1"= levels(seu$subtype)[1],"2"= levels(seu$subtype)[2],"3"= levels(seu$subtype)[3],"4"= levels(seu$subtype)[4],"5"= levels(seu$subtype)[5],"6"= levels(seu$subtype)[6],"7"= levels(seu$subtype)[7],"8"= levels(seu$subtype)[8])
```

```{r}
nbr=merge(as.data.frame(table(dpe$group1)), as.data.frame(table(dpe$group2)), by= "Var1", all=T)
nbr[is.na(nbr)] <- 0
nbr$sum= nbr$Freq.x + nbr$Freq.y
colnames(nbr)= dplyr::recode(colnames(nbr), "Var1"="group")


write.csv(nbr, file.path(RDatadir, "Nbr_Connexions_Cluster.csv"))

nbr$group=as.character(nbr$group)

paga$position= paga$position[which(paga$position$group %in% nbr$group),]

#nbr=nbr[match(paga$position$group, nbr$group),]
rownames(nbr)=nbr$group

nbr=nbr[paga$position$group,]


identical(paga$position$group, nbr$group)

paga$position$sum=nbr$sum
```



Redo the PAGA plot by adjusting the size of the dots to the number of connections:
```{r}
paga[["position"]]$sum= ifelse(is.na(paga[["position"]]$sum)== TRUE, 0.5,paga[["position"]]$sum )
qs::qsave(paga, file.path(RDatadir, "PAGA_object.qs"))

p=ggplot(paga$position, aes(x, y)) +
geom_segment(
  data = paga_edges,
  aes(x = x1, y = y1, xend = x2, yend = y2),
  size = paga_edges$weight*4,
  colour = "black",
  show.legend = FALSE) +
geom_point(data= as.data.frame(paga$position),aes(color = group, size= sum *500),  alpha = 1, show.legend = F) +
scale_color_manual(values = rev(col))+ #+ scale_size_manual(values=as.factor(sum)) +
#geom_text(aes(label = group), color = "black", fontface = "bold") + #[paga$groups]
theme_bw() +
theme(
  axis.title = element_blank(),
  axis.text = element_blank(),
  axis.ticks = element_blank(),
  axis.line = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank())
###
pdf( paste0(pagaDir, "/PAGA_plot_res015.pdf"), height = 5, width = 5)
print(p)
dev.off()
###Save the values to write on the file:

colnames(paga$connectivities)= paga$group_name
writexl::write_xlsx( paga$connectivities_tree,paste0(pagaDir, "/PAGA_ValuesConnections_WriteOnPlot.xlsx"))

qs::qsave(seu, paste0(RDatadir, "/Transitioning_Seurat_WPhate.qs"))
seu=qs::qread(paste0(RDatadir, "/Transitioning_Seurat_WPhate.qs"))
writexl::write_xlsx(paga_edges, file.path(annotDir, "PAGA_Edges_Links_Groups.xlsx"))
```

```{r}
seu=qs::qread(paste0("C:/Users/Melissa S/Desktop/PROJECTS/BRCA1_Tumorigenesis/REVIEW/output/mouse/invivo/RData/Transitioning_Seurat_WPhate.qs"))
```


```{r}
###
pdf( paste0(RDatadir, "/New_PAGA_plot_res015.pdf"), height = 5, width = 5)
print(p)
dev.off()
```


Clean and remove the unused objects:
```{r}
rm(dpe); rm(nbr); rm(paga_edges); rm(paga)
```

```{r}
qs::qsave(seu, file.path(RDatadir, "Subset_ForPAGA.qs"))
```

