---
title: "2_HU_MAYA_EPITHELIAL"
author: "Melissa"
date: "17/08/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
inputDir="C:/Users/Melissa S/Desktop/PROJECTS/Brca1_Tumorigenesis/REVIEW/input/human/snRNAseq/All_cohorts/"
outputDir= "C:/Users/Melissa S/Desktop/PROJECTS/Brca1_Tumorigenesis/REVIEW/output/human"

##
RDatadir <- file.path(outputDir,"RData") ; if(!file.exists(RDatadir)){dir.create(RDatadir)}

QCdir <- file.path(outputDir, "QC") ; if(!file.exists(QCdir)){dir.create(QCdir)}
figDir= file.path(outputDir, "Figures") ; if(!file.exists(annotDir)){dir.create(annotDir)}
cnvDir= file.path(outputDir, "CNV") ; if(!file.exists(cnvDir)){dir.create(cnvDir)}
```

```{r}
library(ggExtra)
library(cowplot)
library(ggplot2)
library(Seurat)
library(dplyr)
```

```{r}
gtheme= theme (axis.text.x = element_text(size=12, face="bold", colour = "black", angle = 0, hjust = 0.5),axis.text.y = element_text(size=12, face="bold", colour = "black"),legend.text = element_text(face = "bold", size = 12),text=element_text(size=14, face="bold", colour = "black"))
###
themplot=theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),axis.ticks.x=element_blank())

'%!in%' <- function(x,y)!('%in%'(x,y))
```

```{r}
seu=qs::qread(file.path(QCdir, "Raw_Seurat_Filtered.qs"))
df_all= seu@assays$RNA@counts
md_all=seu@meta.data
#read the MAYA All list:
md_allH=qs::qread( file.path(RDatadir, "List_MAYA_All_MdAll.qs"))[[1]]
md_all=qs::qread( file.path(RDatadir, "List_MAYA_All_MdAll.qs"))[[2]]
```

Extract the cells corresponding to epithelial cells:
```{r}

selection <- rownames(md_allH[md_allH$cell_annotation %in% c("Luminal epithelial cells","Myoepithelial cells","Epithelial cells"),])

sub_meta <- md_allH[rownames(md_allH) %in% selection,]



df_epithelial <- df_all[,selection]
md_epith <- sub_meta


md_epithH<-MAYA_pathway_analysis(expr_mat=df_epithelial,
                                            modules_list = "hallmark",
                                            is_logcpm=F)

gc()

```


perform consensus clustering on the epithelial cells only:
```{r}
bsl= md_epithH$activity_matrix
library(ConsensusClusterPlus) 

results = ConsensusClusterPlus(bsl,
                                maxK= 12,
                                reps=100,
                                pItem=0.8,
                                pFeature=0.8, 
                                title= "CC_Epith_HU_2",
                                clusterAlg="hc",
                                distance="pearson",
                                innerLinkage= "ward.D",
                                finalLinkage="ward.D",
                                seed=3.14,plot="png")
 
 qs::qsave(results, paste0(RDatadir, "/CC_Epith_HU.qs") )
 ###
  res= calcICL(results,title=paste0(RDatadir,"/CC_Epith_HU_consensus_cluster"),plot="pdf",writeTable=FALSE)
  
  qs::qsave(res, paste0(RDatadir, "/CC_Epith_HU_Results.qs"))
  
  
c3=  results[[3]]$consensusMatrix
```

cut to 4 clusters:
```{r}
library(dendextend)

mt=scale_0_1(md_epithH$activity_matrix)
dend=hclust(dist(t(mt)), method="ward.D2")

dend <- dendextend::color_branches(dend, k = 4)
dend <- dendextend::color_labels(dend, k = 4)
###
pdf(paste0(figDir, "/Fig0b_All_MAYA_Type_ScaledMat.pdf"))
plot(dend, labels=FALSE)
dev.off()


res7=cutree(tree=dend,k=4)  #5 seems the optimal number of clusters
res7=as.data.frame(res7)
colnames(res7)= "GroupID"
res7$cell=rownames(res7)

identical(rownames(res7),rownames(md_epith))

md_epith$ClustH_dendro= res7$GroupID
qs::qsave(md_epith, file.path(figDir, "md_epith_FINAL.qs"))
```



plot the heatmap with info on BRCA:
```{r}
md_epith$ClustH_dendro=paste0("C", md_epith2$ClustH_dendro)

epith_cols= met.brewer(name="Austria",n=length(unique(md_epith$ClustH_dendro)),type="discrete")
names(epith_cols)= unique(md_epith$ClustH_dendro)

metadata= data.frame("subtype"=md_epith$ClustH_dendro, "BRCA"= md_epith$BRCA)
rownames(metadata)= rownames(md_epith)
#metadata= metadata[, -1] %>% as.data.frame()

colours <- list("subtype"=epith_cols, "BRCA"= c("BRCA mut"="darkgreen", "BRCA WT"="grey"))
####---------------------##
colAnn <- HeatmapAnnotation(df = metadata,
  which = 'col',
  col = colours,
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(0.5, 'mm'))
#identical(rownames(metadata), rownames(AUCmatrix))
```


```{r}

split <- factor( metadata$subtype,levels = c("C2", "C4","C3","C1" )) #,"Apod+ T_1"

scaled_mat = scale_0_1(md_epithH$activity_matrix)



hmap <- Heatmap(
  scaled_mat,
  name = "ActMat",
  col = viridis(50, direction = 1), 
  show_row_names = TRUE,
  show_column_names = FALSE,
  cluster_rows =TRUE,
  cluster_columns = TRUE,
  show_column_dend = TRUE,
  show_row_dend = TRUE,
  row_dend_reorder = FALSE,
  column_dend_reorder = FALSE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  #width = unit(100, "mm"),
  top_annotation=colAnn,row_names_gp = gpar(fontsize = 8, fontfamily = "sans"),column_split=split,column_names_gp = gpar(fontsize = 4, fontfamily = "sans"), use_raster=TRUE)



pdf(paste0(figDir, "/Heatmap_ActivityMatrix_subtype_BRCA.pdf"), height = 10, width = 12)
draw(hmap, heatmap_legend_side="right", annotation_legend_side="right")
dev.off()
#---------
png(paste0(figDir, "/Heatmap_ActivityMatrix_subtype_BRCA.png"), height = 2000, width = 2800, res=300)
print(hmap)
dev.off()
```


Redo the UMAP:
```{r}
library(MetBrewer)

col=c("BRCA mut"="darkgreen", "BRCA WT"="grey")
##Other representation with non transparent circles, need shuffling to have a representation of all
umap_custom <- as.data.frame(md_epithH$umap$layout)
colnames(umap_custom)= c("UMAP_1","UMAP_2")

metada_cells_col <- col[match(md_epith$BRCA,names(col))]
draw <- sample(1:dim(umap_custom)[1])


dt=as.data.frame(table( md_epith$BRCA, md_epith$ClustK_dendro))

pdf(paste0(figDir, "/Barplot_Epith_BRCA.pdf"))
p <- ggplot(dt ,aes(x= Var2, y= Freq, fill= Var1))+ 
  geom_bar(stat="identity", position = "fill",show.legend = T,col="black",width=0.8) + scale_fill_manual(values=c( "darkgreen","grey"))  + themplot
print(p)
dev.off()



png(paste0(figDir,"/Fig_Umap_EPITH_BRCA.png"),height=2000, width = 2000,res=300)
p <- ggplot(umap_custom[draw,], aes(x =UMAP_1, y = UMAP_2 ))+ geom_point(size = 2.5, shape = 21,fill = metada_cells_col[draw],alpha=1) + theme_classic() + NoLegend() + NoAxes()
print(p)
dev.off()

#plot subtype:
umap_custom <- as.data.frame(md_epithH$umap$layout)
colnames(umap_custom)= c("UMAP_1","UMAP_2")

metada_cells_col <- epith_cols[match(md_epith$ClustH_dendro,names(epith_cols))]
draw <- sample(1:dim(umap_custom)[1])
p <- ggplot(umap_custom[draw,], aes(x =UMAP_1, y = UMAP_2 ))+ geom_point(size = 2.5, shape = 21,fill = metada_cells_col[draw],alpha=1) + theme_classic() + NoLegend() + NoAxes()

png(paste0(figDir,"/Fig_Umap_EPITH_SUBTYPE.png"),height=2000, width = 2000,res=300)
print(p)
dev.off()
```


DONUT PLOT DISTRIBUTION OF ALL THE EPITHELIAL CLUSTERS:

Donut plot of the distribution of C3:
```{r}
col= rev(c(met.brewer(name="Renoir",n=6,type="discrete"),met.brewer(name="Austria",n=6,type="continuous")))
 names(col)=unique(md_all$orig.ident)


sbt=unique(md_epith$ClustH_dendro)

dt <- data.frame("subtype"= md_epith$ClustH_dendro, "sample"= md_epith$orig.ident) %>% dplyr::filter( subtype %in% sbt)



for( samp in sbt){
  
dt1<-  dt %>% dplyr::filter(subtype== samp)
#same as: DescTools::Entropy(table(dt1$sample), base = 10
D= diversity(as.data.frame(table(dt1$sample))$Freq, "shannon", base= 10)
print(D)
dt1 = dt1 %>%
  group_by(sample)%>% 
  dplyr::count()%>% 
  ungroup()%>% 
  arrange(desc(sample)) %>%
  mutate(percentage = round(n/sum(n),8)*100,
         lab.pos = cumsum(percentage)-.5*percentage)
dt1$sample=as.factor(dt1$sample)


p <- ggplot(data = dt1, 
       aes(x = 2, y = percentage, fill = sample))+
  geom_bar(stat = "identity", show.legend = T)+
  coord_polar("y", start = 200) +
  geom_text(aes(y = lab.pos, label = paste(percentage,"%", sep = "")), col = "white") +
  theme_void() +
  scale_fill_manual(values= col)+
  xlim(.2,2.5) + ggtitle(D)

pdf(paste0(figDir, "/Fig_1Dbis_DonutPlot_Origin_",samp,"_F.pdf"), height = 5, width = 7)
print(p)
dev.off()
}
```


Stack violin of top markers:
```{r}

median.stat <- function(x){
    out <- quantile(x, probs = c(0.5))
    names(out) <- c("ymed")
    return(out) 
}

pdf(paste0(figDir, "/Vln_Markers2.pdf"))
p <- VlnPlot(epith, rev(c("KRT5", "KRT8", "ELF5", "KIT", "PRLR", "LDLR",  "FOSB")), group.by = "ClustH_dendro",  stack = T, pt.size = 0, same.y.lims = TRUE, log = F, fill.by="ident", cols = epith_cols) + NoLegend()  +
    stat_summary(fun.y = median.stat, geom='point', size = 2, colour = "black") 
print(p)
dev.off()
```


get the top contributing genes per pathway:

```{r}
n=10
PCA_object=md_epithH$PCA_obj
module=names(md_epithH$PCA_obj)[1]
#rownames(md_epithH$activity_matrix)

top_genes= function(PCA_object, module, n){
  nb_comp <- length(PCA_object[[module]]$expl_var)
    
top_genes <- c()
contrib <- t(PCA_object[[module]]$gene_contrib)
    if (nb_comp == 1) {
        contrib <- contrib[order(contrib, decreasing = T), ]
        names <- names(contrib)
        top_genes <- names[1:n]
        top_genes=as.data.frame(top_genes)
        colnames(top_genes)=module
    }
    else{
        comp <- sapply(colnames(contrib), function(x) strsplit(x, 
            "PC")[[1]][2])
        
        list_top_genes <- lapply(comp, function(x) {
            contrib <- contrib[order(contrib[, paste0("PC", 
                x)], decreasing = T), ]
            names <- names(contrib[, paste0("PC", x)])
            names[1:n]
        })
        
        top_genes <- do.call(cbind,list_top_genes) %>% as.data.frame()
         colnames(top_genes)=paste0(module,"_mode", comp)
    }

return(top_genes)
  
}
tg=data.frame("start"=rep(NA,10))

  
  for(i in seq(1, length(PCA_object))){
    module=names(PCA_object)[i]
    m= top_genes(PCA_object, names(PCA_object)[i],n=10)
    tg=cbind(tg,m)
  }
   writexl::write_xlsx(tg, file.path(RDatadir, "Top_Contrib_Genes_Modules_Epith_HU_Juxta.xlsx"))
```


From the activity matrix, check the most differetial:
```{r}
library(presto)
help(wilcoxauc)
X= md_epithH$activity_matrix %>% as.data.frame()

#X=t(X) %>% as.data.frame()
identical(colnames(X), colnames(epith))

cells=subset(epith, ClustH_dendro %in% c("C1", "C3", "C4") )
X=X[, colnames(cells)]

y= ifelse(cells$ClustH_dendro=="C3", "BRCA1_luminal", "Luminal")


diff= presto::wilcoxauc(X=X, y= y)
diff=diff %>% dplyr::filter(padj < 0.01)
diff=diff[which(abs(diff$logFC) > 0.4),]

diff=diff %>% dplyr::filter(group=="BRCA1_luminal")

writexl::write_xlsx(diff, file.path(RDatadir, "Wilcox_Hallmark_Brca1LumvsLum.xlsx"))
```


```{r}
dt= md_epithH$activity_matrix[which(rownames(md_epithH$activity_matrix) %in% unique(diff$feature)),]
dt= dt %>% t() %>% as.data.frame()
dt$type=epith$ClustH_dendro

pdf(file.path(figDir, "DiffPathways_Wilcox_Clusters_UpDown_Padj001.pdf"))
for(i in colnames(dt)[-ncol(dt)]){
  dt2= dt[,c(i, "type")]
  p=dt2 %>% reshape2::melt() %>% ggplot( aes(x= type, y= value, fill= type)) + geom_violin() + themplot + scale_fill_manual(values= c("darkgreen","grey")) + ggtitle(i) + 
  stat_summary(fun= median, geom='point', size = 3, colour = "black", shape = 95,aes(x = type, y = value)) + scale_fill_manual(values= epith_cols)
  print(p)
}
 dev.off()
```




```{r}
flattenCorrMatrix <- function(cormat, pmat) {
               ut <- upper.tri(cormat)
               data.frame(
               row = rownames(cormat)[row(cormat)[ut]],
               column = rownames(cormat)[col(cormat)[ut]],
               cor  =(cormat)[ut],
               p = pmat[ut]
               )
}

```

```{r}
maya=subset(seu, cells= rownames(md_epithH))
res<-rcorr(as.matrix(maya@assays$RNA@data))

col<- colorRampPalette(c("blue", "white", "red"))(20)
heatmap(x = res, col = col, symm = TRUE)


cormat <- flattenCorrMatrix(res$r, res$P)

```



plot the heatmap with info on BRCA:
```{r}
md_epith$ClustH_dendro=paste0("C", md_epith$ClustH_dendro)

epith_cols= met.brewer(name="Austria",n=length(unique(md_epith$ClustH_dendro)),type="discrete")
names(epith_cols)= unique(md_epith$ClustH_dendro)

metadata= data.frame("subtype"=md_epith$ClustH_dendro, "BRCA"= md_epith$BRCA)
rownames(metadata)= rownames(md_epith)
#metadata= metadata[, -1] %>% as.data.frame()

colours <- list("subtype"=epith_cols, "BRCA"= c("BRCA mut"="darkgreen", "BRCA WT"="grey"))
####---------------------##
colAnn <- HeatmapAnnotation(df = metadata,
  which = 'col',
  col = colours,
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(0.5, 'mm'))
#identical(rownames(metadata), rownames(AUCmatrix))
```


```{r}

split <- factor( metadata$subtype,levels = c("C2", "C4","C3","C1" )) #,"Apod+ T_1"

scaled_mat = res$r



hmap <- Heatmap(
  scaled_mat,
  name = "ActMat",
 # col = viridis(50, direction = 1), 
  show_row_names = TRUE,
  show_column_names = FALSE,
  cluster_rows =TRUE,
  cluster_columns = TRUE,
  show_column_dend = TRUE,
  show_row_dend = TRUE,
  row_dend_reorder = FALSE,
  column_dend_reorder = FALSE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  #width = unit(100, "mm"),
  top_annotation=colAnn,row_names_gp = gpar(fontsize = 8, fontfamily = "sans"),column_split=split,column_names_gp = gpar(fontsize = 4, fontfamily = "sans"), use_raster=TRUE)



pdf(paste0(figDir, "/Heatmap_PairwiseCorrelation.pdf"), height = 10, width = 12)
draw(hmap, heatmap_legend_side="right", annotation_legend_side="right")
dev.off()
#---------
png(paste0(figDir, "/Heatmap_PariwiseCorrelation.png"), height = 2000, width = 2800, res=300)
print(hmap)
dev.off()
```



```{r}
maya=epith
maya$class= paste0(maya$ClustH_dendro,"_",maya$BRCA)


c6=subset(maya, class=="C1_BRCA mut")
mat_c6= c6@assays$RNA@data
library(Hmisc)
res<-rcorr(as.matrix(mat_c6))
tmp <- flattenCorrMatrix(res$r, res$P)
tmp$clust="C1_BRCA mut"


c7=subset(maya, class=="C1_BRCA WT")
mat_c7= c7@assays$RNA@data
library(Hmisc)
res<-rcorr(as.matrix(mat_c7))
tmp7 <- flattenCorrMatrix(res$r, res$P)
tmp7$clust="C1_BRCA WT"


c4=subset(maya, class=="C2_BRCA mut")
mat_c4= c4@assays$RNA@data
library(Hmisc)
res<-rcorr(as.matrix(mat_c4))
tmp4 <- flattenCorrMatrix(res$r, res$P)
tmp4$clust="C2_BRCA mut"



c3=subset(maya, class=="C2_BRCA WT")
mat_c3= c3@assays$RNA@data
library(Hmisc)
res<-rcorr(as.matrix(mat_c3))
tmp3 <- flattenCorrMatrix(res$r, res$P)
tmp3$clust="C2_BRCA WT"

c2=subset(maya, class=="C3_BRCA mut")
mat_c2= c2@assays$RNA@data
library(Hmisc)
res<-rcorr(as.matrix(mat_c2))
tmp2 <- flattenCorrMatrix(res$r, res$P)
tmp2$clust="C3_BRCA mut"


c1=subset(maya, class=="C3_BRCA WT")
mat_c1= c1@assays$RNA@data
library(Hmisc)
res<-rcorr(as.matrix(mat_c1))
tmp1 <- flattenCorrMatrix(res$r, res$P)
tmp1$clust="C3_BRCA WT"


c0=subset(maya, class=="C4_BRCA mut")
mat_c0= c0@assays$RNA@data
library(Hmisc)
res<-rcorr(as.matrix(mat_c0))
tmp0 <- flattenCorrMatrix(res$r, res$P)
tmp0$clust="C4_BRCA mut"

c00=subset(maya, class=="C4_BRCA WT")
mat_c00= c00@assays$RNA@data
library(Hmisc)
res<-rcorr(as.matrix(mat_c00))
tmp00 <- flattenCorrMatrix(res$r, res$P)
tmp00$clust="C4_BRCA WT"
```

```{r}
datcor=list("C1_BRCA mut"=tmp$cor , "C1_BRCA WT"=tmp7$cor , "C2_BRCA mut"=tmp4$cor , "C2_BRCA WT"=tmp3$cor, "C3_BRCA mut"=tmp2$cor, "C3_BRCA WT"=tmp1$cor, "C4_BRCA mut"=tmp0$cor, "C4_BRCA WT"=tmp00$cor)  
  

res2 <- as.data.frame(do.call(cbind, datcor)) 
qs::qsave(res2, file.path(RDatadir,"PairwiseCorr_Clusters_Brca.qs"))

pdf(file.path(figDir,"IntraCorr_ClusterEpith_BRCA.pdf"))
p <-  res2 %>%  reshape2::melt() %>%  ggplot() + 
  geom_violin(aes(x = variable, y = value, col= variable, fill= variable), show.legend = F,width = 4.5) +  geom_boxplot(width = .1, aes(x = variable, y = value)) +
  stat_summary(fun= median, geom='point', size = 3, colour = "black", shape = 95,aes(x = variable, y = value)) + ylab("pairwise_IntraClust_Correlation") + themplot 
  print(p)
dev.off()
```

```{r}
median(res2$`C1_BRCA WT`)/median(res2$`C1_BRCA mut`) #0.9122284
median(res2$`C2_BRCA WT`)/median(res2$`C2_BRCA mut`) #1.004145
median(res2$`C3_BRCA WT`)/median(res2$`C3_BRCA mut`) # 1.276811
median(res2$`C4_BRCA WT`)/median(res2$`C4_BRCA mut`) # 0.9522691

wilcox.test(y=res2$`C1_BRCA WT`, x=res2$`C1_BRCA mut`, paired = FALSE, alternative = "two.sided") # p-value < 2.2e-16

wilcox.test(y=res2$`C2_BRCA WT`, x=res2$`C2_BRCA mut`, paired = FALSE, alternative = "two.sided")

wilcox.test(y=res2$`C3_BRCA WT`, x=res2$`C3_BRCA mut`, paired = FALSE, alternative = "two.sided")

wilcox.test(y=res2$`C4_BRCA WT`, x=res2$`C4_BRCA mut`, paired = FALSE, alternative = "two.sided")
```

```{r}
epith=maya
DefaultAssay(epith)="RNA"

Idents(epith)="ClustH_dendro"

comph= FindMarkers(epith, ident.1 = "3", ident.2 = c("1", "4"), only.pos = F, logfc.threshold = 0.5) %>% dplyr::filter( p_val_adj < 0.05)
comph$gene= rownames(comph)

comph=comph[-grep("^AC", comph$gene, value = F),]



comph= comph  %>% arrange(desc(avg_log2FC))

writexl::write_xlsx(comph, file.path(RDatadir, "DEG_C3_vsC14.xlsx"))
```

```{r}
l=list(md_all,md_epith,md_allH,d_epithH, md_allH, md_epithH )
qs::qsave(l,file.path(RDatadir, "/MAYA_save.qs"))
qs::qsave(epith, file.path(RDatadir, "Epithelial_MAYA_All.qs"))
```

