---
title: "Normal_Preneo_Comparison"
author: "Melissa"
date: "17/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
mainDir= here::here() #Should be the root folder Tumor_BRCA_p53

source(knitr::purl(file.path(mainDir,"Scripts","scRNAseq_Functions.Rmd"), quiet=TRUE))
source(knitr::purl(file.path(mainDir,"Scripts","global_variables.Rmd"), quiet=TRUE))

 #"C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Tumor_BRCA_p53"
inputDir= file.path(mainDir,"input" ,"human","scRNAseq","publicDatasets","input_GSE161529")
outputDir= file.path(mainDir, "output","human","public_data","EMBO")
###
plotDir= file.path(outputDir,"Plots") ; if(!file.exists(plotDir)){dir.create(plotDir)}
RdataDir= file.path(outputDir,"Rdata") ; if(!file.exists(RdataDir)){dir.create(RdataDir)}

gtheme= theme (axis.text.x = element_text(size=12, face="bold", colour = "black", angle = 0, hjust = 0.5),axis.text.y = element_text(size=12, face="bold", colour = "black"),legend.text = element_text(face = "bold", size = 12),text=element_text(size=14, face="bold", colour = "black"))
###
themplot=theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),axis.ticks.x=element_blank())
```




########################### PART3 #########################################################
Compare LP compartment from Normal & Normal preneo:
```{r}
e1 <- qs::qread(paste0(RdataDir, "/Epithelial_EMBO_Processed_Annotated.qs"))
lp <- subset(e1, subtype=="N_LP")
Idents(lp)="Condition"
lmarkers <- FindAllMarkers(lp, only.pos = TRUE, logfc.threshold = 0.5 ) %>% dplyr::filter(p_val_adj<0.05 )
lmarkers= lmarkers[-which(lmarkers$gene %in% grep("^RP", lmarkers$gene, value = TRUE)),]

writexl::write_xlsx(lmarkers, paste0(RdataDir, "/LP_DEG_4Conditions_NORP.xlsx"))
```


############################## Comparison of Normal & Preneo compartments #######################
Isolate normal samples :


```{r}
library(MetBrewer)
png(paste0(plotDir, "/UMAP_Epithelial_Clusters.png"), height = 1200, width = 1200, res = 300)
 p <- DimPlot(e1,  pt.size = 0.75, label = TRUE, label.size = 2, cols = met.brewer(name="Juarez",n=length(unique(e1$seurat_clusters)),type="continuous") ) + NoLegend()+ NoAxes()
  p[[1]]$layers[[1]]$aes_params$alpha = 0.4
  print(p)
dev.off()
####
png(paste0(plotDir, "/UMAP_Epithelial_Origin.png"), height = 1200, width = 1200, res = 300)
 p <- DimPlot(e1,  pt.size = 0.75, label = FALSE, label.size = 2, group.by = "Condition", cols = rev(met.brewer(name="Egypt",n=length(unique(e1$Condition)),type="discrete")) ) + NoLegend()+ NoAxes()
  p[[1]]$layers[[1]]$aes_params$alpha = 0.4
  print(p)
dev.off()
```


```{r}
Idents(e1)= "subtype"
normal = subset(e1, idents = c("N_LP", "N_Basal", "N_ML") )
gc()

DefaultAssay(normal)= "SCT"
normal= normal %>% RunPCA(., npcs = 60)  %>% RunUMAP(., reduction= "pca", dims= 1:60 ) %>% FindNeighbors(., reduction = "pca", dims = 1:60) %>% FindClusters(., resolution = 1.2)
#####
png(paste0(plotDir, "/UMAP_Epithelial_Condition.png"), height = 1200, width = 1200, res = 300)
 p <- DimPlot(e1,  pt.size = 0.75, label = FALSE, label.size = 2, group.by = "Condition", cols = rev(met.brewer(name="Egypt",n=length(unique(normal$Condition)),type="discrete"))[c(2,1,3,4)] ) + NoLegend()+ NoAxes()
  p[[1]]$layers[[1]]$aes_params$alpha = 0.4
  print(p)
dev.off()
########################
png(paste0(plotDir, "/UMAP_Epithelial_Samples.png"), height = 1200, width = 1200, res = 300)
 p <- DimPlot(e1,  pt.size = 0.75, label = FALSE, label.size = 2, group.by = "orig.ident", cols = rev(met.brewer(name="VanGogh2",n=length(unique(e1$orig.ident)),type="continuous")) ) + NoLegend()+ NoAxes()
  p[[1]]$layers[[1]]$aes_params$alpha = 0.4
  print(p)
dev.off()
#######################

###Legend:
pdf(paste0(plotDir, "/Legend_Condition.pdf"))
  as_ggplot(get_legend(DimPlot(normal,  pt.size = 0.75, label = FALSE, label.size = 2, group.by = "Condition", cols = rev(met.brewer(name="Egypt",n=length(unique(normal$Condition)),type="discrete"))[c(2,1,3,4)] )))
dev.off()

pdf(paste0(plotDir, "/Legend_Sample.pdf"))
  as_ggplot(get_legend(DimPlot(e1,  pt.size = 0.75, label = FALSE, label.size = 2, group.by = "orig.ident", cols = rev(met.brewer(name="VanGogh2",n=length(unique(e1$orig.ident)),type="continuous")))))
dev.off()

qs::qsave(normal, paste0(RdataDir, "/Normal_EpithelialCompartments.qs"))
normal= qs::qread(paste0(RdataDir, "/Normal_EpithelialCompartments.qs"))
```

Equilibrate epithelial cells from Normal/Preneo : get the same nuùber of cells from both conditions, then have a look at the distribution of LP/Basal/ML proportions between the two conditions
```{r}
dt=data.frame("subtype"=normal$subtype, "condition"=normal$Condition)
dt$condition=ifelse(dt$condition=="Normal BRCA1+/- pre-neoplastic", "Preneoplastic", dt$condition)
dt= dt %>% filter( condition %in% c("Preneoplastic", "Normal"))
 table(dt$condition)

dt= dt[ c(sample(rownames(dt %>% filter(condition== "Normal")), 4264 ), rownames(dt)[which(dt$condition=="Preneoplastic")]),]
```

From the equilibrated Epithelial compartments (same number of cells from Normal & Normal-Preneo):
+ Isolate only LP cells
+ check the embeddings on a PCA or UMAP plot


```{r}
equil <- subset(normal, cells= rownames(dt))
Idents(equil)="subtype"
equil= subset(equil, idents = "N_LP")
equil <- equil  %>% RunPCA(., npcs = 60)  %>% RunUMAP(., reduction= "pca", dims= 1:60 ) %>% FindNeighbors(., reduction = "pca", dims = 1:60) %>% FindClusters(., resolution = 1.2)
```




Check the differences between LPs:
```{r}
DefaultAssay(equil)="RNA"
equil=NormalizeData(equil)

Idents(equil)= "Condition"
c1 <- FindMarkers(equil, ident.1 = "Normal BRCA1+/- pre-neoplastic", ident.2 = "Normal", only.pos = TRUE, logfc.threshold = 0.3, min.diff.pct = 0.1 ) %>% dplyr::filter(p_val_adj < 0.05) 
c1$gene=rownames(c1)
writexl::write_xlsx(c1, paste0(RdataDir, "/DEG_LP_Preneo_vsNormal_Diff01_FC03.xlsx") )
```
##AUTOMATIC CALCULATION:
```{r}
pc2=equil@reductions$pca@feature.loadings[,2] %>% as.data.frame() %>% rownames_to_column("gene")
colnames(pc2)=c("gene","eigen")
pc2 = pc2 %>% dplyr::filter(eigen < 0)
##remove potential ambient RNA:
pc2= pc2 %>% dplyr::filter(! gene %in% grep("^IG", pc2$gene, value = TRUE))
pc2=pc2[order(pc2$eigen, decreasing = FALSE),]
#######################
##################################################################
pc3=equil@reductions$pca@feature.loadings[,3] %>% as.data.frame() %>% rownames_to_column("gene")
colnames(pc3)=c("gene","eigen")
pc3 = pc3 %>% dplyr::filter(eigen > 0)
##remove potential ambient RNA:
pc3= pc3 %>% dplyr::filter(! gene %in% grep("^IG", pc3$gene, value = TRUE))
pc3=pc3[order(pc3$eigen, decreasing = TRUE),]
#################################

top_enrichment= function(data1, data2, n, x, y){
  top1= data1[1:n,]
  top2= data2[1:n,]
  dt= rbind(top1, top2)
  ##Hallmarck:
  m_t2g <- msigdbr::msigdbr(species = "Homo sapiens", category = "H") %>% dplyr::select(gs_name, gene_symbol) %>% as.data.frame()
  
  p2H<- clusterProfiler::enricher(unique(dt$gene) , TERM2GENE=m_t2g) @result %>% dplyr::filter(., p.adjust< 0.05)%>% .[,c(1,6,8)]
### C2:
m_t2g <- msigdbr::msigdbr(species = "Homo sapiens", category = "C2") %>% dplyr::select(gs_name, gene_symbol) %>% as.data.frame()

p2C2<- clusterProfiler::enricher(unique(dt$gene) , TERM2GENE=m_t2g) @result %>% dplyr::filter(., p.adjust< 0.05)%>% .[,c(1,6,8)]
  
## C6 
m_t2g <- msigdbr::msigdbr(species = "Homo sapiens", category = "C6") %>% dplyr::select(gs_name, gene_symbol) %>% as.data.frame()

p2C6<- clusterProfiler::enricher(unique(dt$gene) , TERM2GENE=m_t2g) @result %>% dplyr::filter(., p.adjust< 0.05)%>% .[,c(1,6,8)]
######################
l= list( top1,top2 ,"H"= p2H, "C2"= p2C2, "C6"= p2C6)
writexl::write_xlsx(l, paste0(RdataDir, "/LP_Enrichment_",n ,"_TopGenes_PC_", x, "_and_",y,".xlsx"))
return(l)
}
 
tst= top_enrichment(pc2, pc3, n=100, x=2, y=3)
tst2= top_enrichment(pc2, pc3, n=50, x=2, y=3)
tst3= top_enrichment(pc2, pc3, n=30, x=2, y=3)
tst3= top_enrichment(pc2, pc3, n=10, x=2, y=3)
```
Baplot representation:
```{r}
h= tst3$C2 %>% filter(ID %in% grep("^REACTOME_", tst3$C2$ID, value = TRUE))
h=h[1:4,]
h$type= "Reactome"
###
k= tst3$C2 %>% filter(ID %in% grep("^KEGG_", tst3$C2$ID, value = TRUE))
k$type= "KEGG"
##
m= tst3$H
m$type="Hallmark"
###
n= tst3$C2 %>% filter(ID %in% grep("MAMMARY_", tst3$C2$ID, value = TRUE))
n$type="Mammary"
############################################
dt1= rbind(m, h, k, n)
```
  

```{r}
##########
m_t2g <- msigdbr::msigdbr(species = "Homo sapiens", category = "H") %>% dplyr::select(gs_name, gene_symbol) %>% as.data.frame()
p2H<- clusterProfiler::enricher(pc2$gene , TERM2GENE=m_t2g) @result %>% dplyr::filter(., p.adjust< 0.05)%>% .[,c(1,6,8)]
#########################
m_t2g <- msigdbr::msigdbr(species = "Homo sapiens", category = "C2") %>% dplyr::select(gs_name, gene_symbol) %>% as.data.frame()
p2C2<- clusterProfiler::enricher(pc2$gene , TERM2GENE=m_t2g) @result %>% dplyr::filter(., p.adjust< 0.05)%>% .[,c(1,6,8)]
#####
lst <- list("Hallmark"= p2H, "C2"= p2C2)
writexl::write_xlsx(lst, path = paste0(RdataDir, "/TopContribPC2_LP_Equilibrated_H_C2.xlsx"))
```

Rank gene contribution:
```{r}
pc2=pc2[order(pc2$eigen, decreasing = FALSE),]
pc2 = pc2 %>% dplyr::filter(eigen <  (-0.05))
## represent the eigen values:

pdf(paste0(plotDir, "/Contrib_Genes_Neg_PC2_005.pdf"), height = 10, width = 6)
pc2 %>% ggplot(aes(x= reorder(gene, eigen), y= eigen)) + geom_point() + 
  geom_segment(aes(xend = gene, yend = 0), size = 1.2) + coord_flip() + themplot
dev.off()

##########
m_t2g <- msigdbr::msigdbr(species = "Homo sapiens", category = "H") %>% dplyr::select(gs_name, gene_symbol) %>% as.data.frame()
p2H<- clusterProfiler::enricher(pc2$gene , TERM2GENE=m_t2g) @result %>% dplyr::filter(., p.adjust< 0.05)%>% .[,c(1,6,8)]
#########################
m_t2g <- msigdbr::msigdbr(species = "Homo sapiens", category = "C2") %>% dplyr::select(gs_name, gene_symbol) %>% as.data.frame()
p2C2<- clusterProfiler::enricher(pc2$gene , TERM2GENE=m_t2g) @result %>% dplyr::filter(., p.adjust< 0.00001)%>% .[,c(1,6,8)]
#####
dt1= p2H[, -2]
dt2= p2C2[, -2]
dt2= dt2[which(dt2$ID %in% grep("REACTOME", dt2$ID, value = TRUE)),]
dt3= p2C2[which(p2C2$ID %in% grep("SENESE_HDAC1_TARGETS_UP", p2C2$ID, value = TRUE)),]
dt2=rbind(dt2, dt3[,-2])

dt1= rbind(dt1, dt2)
dt1[,2]= gsub("/", ",", dt1[,2])
tmp<- dt1 %>%
  mutate_at(vars(ID:geneID), ~ str_split(., pattern = ",")) %>%
  unnest() 
####
pdf(paste0(plotDir, "/PEA_C2_H_LP_Equilibrated.pdf"), width = 15, height = 10)
reshape2::melt(p[, c(1,2,4)]) %>% ggplot(aes(x=reorder(factor(ID), -log10(value)), y = -log10(value))) + 
   # facet_wrap(~ type) 
   geom_bar(aes(fill = factor(type)), stat = "identity") + themplot + coord_flip() + theme(legend.position = NULL) + ylab("")
dev.off()
```


```{r}
colsE = rev(met.brewer(name="Egypt",n=length(unique(normal$Condition)),type="discrete"))[c(2,1,3,4)]
names(colsE)= unique(normal$Condition)
umap_custom <- as.data.frame(normal[["umap"]]@cell.embeddings)

metada_cells <- normal$Condition
metada_cells_col <- colsE[match(normal$Condition,names(colsE))]

draw <- sample(1:dim(umap_custom)[1])
 
png(paste0(clustDir,"/UMAP_Normal_EMBO_Condition.png"),height=1000, width = 1000,res=300)
p <- ggplot(umap_custom[draw,], aes(x =UMAP_1, y = UMAP_2 ,fill = metada_cells_col[draw] ))+ geom_point(size = 0.75, shape = 21,fill = metada_cells_col[draw],alpha=1) + theme_classic() + theme(legend.position = "right") + NoAxes()
print(p)
dev.off()
```

```{r}
p <- StackedVlnPlot(normal, features = c("CXCL8", "MMP3","IL6","CXCL1","HMGA1","FOSL1"))

pdf(paste0(plotDir, "/StackVln.pdf"))
print(p)
dev.off()
```


##------------ SCORE THE P16 HUMAN PRE TUMORAL SIGNATURE-------------##

```{r}
qs::qsave(equil, paste0(plotDir, "/LP_fromEquilibrated_Seurat.qs"))
equil=qs::qread(paste0(RdataDir,"/LP_fromEquilibrated_Seurat.qs" ))
normal= qs::qread(paste0(dtDir, "/Normal_EpithelialCompartments.qs"))
```


Read the pathway enrichment analysis file:
```{r}
epith= qs::qread(file.path(mainDir, "output", "mouse", "scRNAseq", "invivo", "/RData/Epithelial_FinalAnnotation.qs"))
sigH=readxl::read_excel(file.path(outputDir, "human", "publicDatasets", "TCGA", "RData", "/List_VarianceImportance_CPTAC_TCGA.xlsx"))
sigM= sigH
sigH= sigH %>% dplyr::filter(gene %in% seq(1, 13))
```



Test P16 human signature:
```{r}
DefaultAssay(normal)="RNA"
features=list("p16sigH"= sigH$feature)
library(UCell)
normal= UCell::AddModuleScore_UCell(normal, features = features)

normal= subset(normal, Condition %in% c("Normal","Normal BRCA1+/- pre-neoplastic"))
qs::qsave(normal,"/Normal_EpithelialCompartments.qs")

dat=normal@meta.data[, c(23,22,10)]
#Vlnplot:

p= dat %>% ggplot(aes(x= subtype, y= p16sigH_UCell)) + geom_violin() + geom_jitter(aes(color=Condition), size=2, alpha=0.4) + scale_color_manual(values=rev(met.brewer(name="Egypt",n=length(unique(dat$Condition)),type="discrete"))[c(2,1,3,4)]) 

pdf(paste0(plotDir, "/Vln_EMBO_p16Human.pdf"))
p <- VlnPlot(normal, features = "p16sigH_UCell", split.by = "Condition", group.by = "subtype",  cols = rev(met.brewer(name="Egypt",n=4,type="discrete"))[c(2,1,3,4)]) + NoLegend() + stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Normal",aes(label=..p.adj..))
print(p)
dev.off()

png(paste0(mainDir, "/Vln_EMBO_p16Human.png"), height = 1000, width = 1000, res = 300)
print(p)
dev.off()
```

Statistical comparison :
```{r}
library(rstatix)
dt= data.frame("subtype"= normal$subtype, "condition"= normal$Condition, "p16"= normal$p16sigH_UCell)

stat.test <- dt %>%
  group_by(subtype) %>%
  wilcox_test(p16 ~ condition, alternative = "less") %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")

###
dt= data.frame("subtype"= normal$subtype, "condition"= normal$Condition, "p16"= normal$p16sigH_UCell)

stat.test <- dt %>%
  group_by(subtype) %>%
  wilcox_test(p16 ~ condition, alternative = "less") %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")


dt$type=paste0(dt$subtype,"_", dt$condition)
dt=dt[,-c(1,2)]
wilcoxon_location<- wilcoxon.test(dt, 2, "type", p.adjust.method="bonferroni")
##########
df= data.frame("alt"= dat$Nbr_Alterations, "subtype"= dat$Condition)

df$subtype= factor(df$subtype, levels=c("Normal",  "Normal BRCA1+/- pre-neoplastic", "Triple negative (BRCA1) tumor"))

stat2=df %>%
  wilcox_test(alt ~ subtype, alternative = "less") %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
```



