---
title: "Informativity_PCA"
author: "Melissa"
date: "17/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
mainDir= here::here() #Should be the root folder Tumor_BRCA_p53

source(knitr::purl(file.path(mainDir,"Scripts","scRNAseq_Functions.Rmd"), quiet=TRUE))
source(knitr::purl(file.path(mainDir,"Scripts","global_variables.Rmd"), quiet=TRUE))
source(knitr::purl(file.path(mainDir,"Scripts","scRNAseq","mouse","GSE161529","PC_informativity_Functions.Rmd"), quiet=TRUE))
 #"C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Tumor_BRCA_p53"
inputDir= file.path(mainDir,"input" ,"human","scRNAseq","publicDatasets","input_GSE161529")
outputDir= file.path(mainDir, "output","human","public_data","EMBO")
###
plotDir= file.path(outputDir,"Plots") ; if(!file.exists(plotDir)){dir.create(plotDir)}
RdataDir= file.path(outputDir,"Rdata") ; if(!file.exists(RdataDir)){dir.create(RdataDir)}
pcDir=  file.path(outputDir,"PC_information") ; if(!file.exists(pcDir)){dir.create(pcDir)}

gtheme= theme (axis.text.x = element_text(size=12, face="bold", colour = "black", angle = 0, hjust = 0.5),axis.text.y = element_text(size=12, face="bold", colour = "black"),legend.text = element_text(face = "bold", size = 12),text=element_text(size=14, face="bold", colour = "black"))
###
themplot=theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),axis.ticks.x=element_blank())
```



This gives the top PCs which separate most the two :
Method1:
```{r}
equil= qs::qread(paste0(RdataDir, "/LP_fromEquilibrated_Seurat.qs"))
e1 <- qs::qread(paste0(RdataDir, "/Epithelial_EMBO_Processed_Annotated.qs"))

lp= subset(e1, annot2=="LP")
lp= subset(lp, Condition %in% c("Normal", "Normal BRCA1+/- pre-neoplastic"))
lp= subset(lp, cells= c(colnames(lp)[which(lp$Condition=="Normal BRCA1+/- pre-neoplastic")], sample(colnames(lp)[which(lp$Condition=="Normal")], 2549)))


lp= lp %>% RunPCA( npcs = 60) %>% RunUMAP(reduction= "pca", dims= 1:60 ) %>% FindNeighbors(., reduction = "pca", dims = 1:60) %>% FindClusters(., resolution = 1.2)


dtest <- PC_separability(lp, nPCs = 20, "LP")
dtest$Informativity= ifelse((dtest$KStest < quantile(dtest$KStest, 0.8) & dtest$Dvalue > quantile(dtest$Dvalue, 0.8)), "Informative", "nonInformative")
```

GET THE VARIANCE EXPLAINED
```{r}
mat <- Seurat::GetAssayData(lp, assay = "SCT", slot = "scale.data")
pca <- equil[["pca"]]
# Get the total variance:
total_variance <- sum(matrixStats::rowVars(mat))
eigValues = (pca@stdev)^2  ## EigenValues
varExplained = eigValues / total_variance
 
dtest$VarExplained<- varExplained[1:20]*100
```

```{r}
pcvar= round(lp@reductions$pca@stdev,3) %>% as.data.frame()
pcvar$PC=paste0("PC_", seq(1,nrow(pcvar)))
colnames(pcvar)=c("SD", "PC")
dtest$SD= pcvar$SD[1:20]
```


Method2: Linear regression:
```{r}
dttestlm= PC_separability_lm(equil, nPCs = 20)
dttestlm$take= ifelse(dttestlm$Rsquare > quantile(dttestlm$Rsquare, 0.8), "Informative", "NonInformative")
dttestlm$VarExplained<- varExplained[1:20]
```

```{r}
res= list(dtest, dttestlm)
names(res)=c("LP_PC_KStest", "LP_PC_LM")
writexl::write_xlsx(res, paste0(pcDir, "/nPC20_Effectiveness_LP_Normal_NormalPreneo.xlsx"))
```

From the informative PCs take the ones which perfectly separates the two conditions:
example for PC2:
```{r}
 pc <- Embeddings(lp, reduction = "pca")[, 1:20] %>% as.data.frame()
  identical(rownames(pc), rownames(lp@meta.data))
  pc$condition= lp$Condition
  pc$sample= lp$sample
  
 pdf(paste0(pcDir, "/Barplot_Distrib_LP_PC_notFlipped_BestWorst_Informative_PCs.pdf"), width = 15, height = 8)
for( i in seq(0, 30,10)[-1]){
  for(j in c(2,6,13,1,11)){ #these are informative PCs
    
    dt <- reshape2::melt(table(cut(pc[,j], seq(min(pc[,j]), max(pc[j]), i), include.lowest = TRUE), pc$condition)) 
    dt$Var2= factor(dt$Var2, levels=c("Normal BRCA1+/- pre-neoplastic", "Normal"))
    
    
    p <- dt %>% ggplot(aes(x= Var1, y=value, fill=Var2)) + geom_bar(stat="identity") + themplot + xlab(paste0("PC_",j,"_BinSize=", i)) + scale_fill_manual(values=rev(met.brewer(name="Egypt",n=4,type="discrete"))[c(1,2)]) 
    print(p)
    rm(dt)
#################
     dt <- reshape2::melt(table(cut(pc[,j], seq(min(pc[,j]), max(pc[j]), i), include.lowest = TRUE), pc$sample, pc$condition))
    
        g <- reshape2::melt(table(cut(pc[,j], seq(min(pc[,j]), max(pc[,j]), i), include.lowest = TRUE), pc$sample)) %>% ggplot(aes(x= Var1, y=value, fill=Var2)) + geom_bar(stat="identity") + themplot+ theme(legend.position = "none") + xlab(paste0("PC_",j,"_BinSize=", i)) + scale_fill_manual(values=rev(met.brewer(name="VanGogh2",n=length(unique(dt$Var2)),type="continuous"))) 
    print(g)

  }
}
 dev.off()
 ##################FOR CELL CYCLE #####################
 ####Get the intervals for which there is a shift between the two pops in favor of preneoplastic:
  dt <- as.data.frame.matrix(table(cut(pc[,2], seq(min(pc[,2]), max(pc[,2]), 20), include.lowest = TRUE), pc$condition))
  dt$ratio= round(dt$`Normal BRCA1+/- pre-neoplastic`/dt$Normal, 3)
  dt= dt %>% dplyr::filter(ratio > 1.5 | ratio < 0.5)
```



############################# BASAL CELLS ######################
```{r}
normal= qs::qread(paste0(RdataDir, "/Normal_EpithelialCompartments.qs"))
e1 <- qs::qread(paste0(RdataDir, "/Epithelial_EMBO_Processed_Annotated.qs"))

basal <- subset(e1, annot2=="Basal")
basal= subset(basal, Condition %in% c("Normal", "Normal BRCA1+/- pre-neoplastic"))
basal= basal %>% RunPCA( npcs = 60) %>% RunUMAP(reduction= "pca", dims= 1:60 ) %>% FindNeighbors(., reduction = "pca", dims = 1:60) %>% FindClusters(., resolution = 1.2)
```

```{r}
dtestb <- PC_separability(basal, nPCs = 20, "basal")
dtestb$take= ifelse((dtestb$KStest < quantile(dtestb$KStest, 0.8) & dtestb$Dvalue > quantile(dtestb$Dvalue, 0.8)), "Informative", "nonInformative")
```

GET THE VARIANCE EXPLAINED
```{r}
mat <- Seurat::GetAssayData(basal, assay = "SCT", slot = "scale.data")
pca <- basal[["pca"]]
# Get the total variance:
total_variance <- sum(matrixStats::rowVars(mat))
eigValues = (pca@stdev)^2  ## EigenValues
varExplained = eigValues / total_variance
 
dtestb$VarExplained<- varExplained[1:20]*100
```

```{r}
pcvar= round(basal@reductions$pca@stdev,3) %>% as.data.frame()
pcvar$PC=paste0("PC_", seq(1,nrow(pcvar)))
colnames(pcvar)=c("SD", "PC")
dtestb$SD= pcvar$SD[1:20]
```


```{r}
dttestlm= PC_separability_lm(basal, nPCs = 20)
dttestlm$take= ifelse(dttestlm$Rsquare > quantile(dttestlm$Rsquare, 0.8), "Informative", "NonInformative")


res= list(dtestb, dttestlm)
names(res)=c("Basal_PC_KStest", "Basal_PC_LM")
writexl::write_xlsx(res, paste0(pcDir, "/PC_Effectiveness_Basal_Normal_NormalPreneo_PC20.xlsx"))
```


```{r}
 pc <- Embeddings(basal, reduction = "pca")[, 1:20] %>% as.data.frame()
  identical(rownames(pc), rownames(basal@meta.data))
  pc$condition= basal$Condition
  pc$sample= basal$sample
  
 pdf(paste0(pcDir, "/Barplot_Distrib_BASAL_PC_notFlipped.pdf"), width = 15, height = 8)
for( i in seq(0, 30,10)[-1]){
  for(j in c(5:7)){
    
    dt <- reshape2::melt(table(cut(pc[,j], seq(min(pc[,j]), max(pc[j]), i), include.lowest = TRUE), pc$condition)) %>% as.data.frame()
    dt$Var2= factor(dt$Var2, levels=c("Normal BRCA1+/- pre-neoplastic", "Normal"))
    #pdf(paste0(RdataDir, "/Barplot_Distrib_LP_PC_", j, "_BinSize_",i,".pdf"), width = 8, height = 8)
    
    p <-  ggplot(dt,aes(x= Var1, y=value, fill=Var2)) + geom_bar(stat="identity") + themplot+ theme(legend.position = "none") + xlab(paste0("PC_",j,"_BinSize=", i)) + scale_fill_manual(values=rev(met.brewer(name="Egypt",n=4,type="discrete"))[c(1,2)]) 
    print(p)
#################
     dt <- reshape2::melt(table(cut(pc[,j], seq(min(pc[,j]), max(pc[j]), i), include.lowest = TRUE), pc$sample))
    
        g <- dt %>% ggplot(aes(x= Var1, y=value, fill=Var2)) + geom_bar(stat="identity") + themplot+ theme(legend.position = "none") + xlab(paste0("PC_",j,"_BinSize=", i)) + scale_fill_manual(values=rev(met.brewer(name="VanGogh2",n=length(unique(basal$orig.ident)),type="continuous"))) 
    print(g)

  }
}
 dev.off()
```




############################# ML CELLS ######################
```{r}
ml <- subset(e1, annot2=="ML")
ml= subset(ml, Condition %in% c("Normal", "Normal BRCA1+/- pre-neoplastic"))
##equilibrate the number of cells:
ml= subset(ml, cells= c(colnames(ml)[which(ml$Condition=="Normal BRCA1+/- pre-neoplastic")], sample(colnames(ml)[which(ml$Condition=="Normal")], 570)))
DefaultAssay(ml)="SCT"
ml= ml   %>% RunPCA( npcs = 60) %>% RunUMAP(reduction= "pca", dims= 1:60 ) %>% FindNeighbors(., reduction = "pca", dims = 1:60) %>% FindClusters(., resolution = 1.2)
```

```{r}
dtestm <- PC_separability(ml, nPCs = 20, "ML")
dtestm$take= ifelse((dtestm$KStest < quantile(dtestm$KStest, 0.8) & dtestm$Dvalue > quantile(dtestm$Dvalue, 0.8)), "Informative", "nonInformative")
```

GET THE VARIANCE EXPLAINED
```{r}
mat <- Seurat::GetAssayData(ml, assay = "SCT", slot = "scale.data")
pca <- ml[["pca"]]
# Get the total variance:
total_variance <- sum(matrixStats::rowVars(mat))
eigValues = (pca@stdev)^2  ## EigenValues
varExplained = eigValues / total_variance
 
dtestm$VarExplained<- varExplained[1:20]*100
```


```{r}
pcvar= round(ml@reductions$pca@stdev,3) %>% as.data.frame()
pcvar$PC=paste0("PC_", seq(1,nrow(pcvar)))
colnames(pcvar)=c("SD", "PC")
dtestm$SD= pcvar$SD[1:20]
```


```{r}
dttestlm= PC_separability_lm(ml, nPCs = 20)
dttestlm$take= ifelse(dttestlm$Rsquare > quantile(dttestlm$Rsquare, 0.8), "Informative", "NonInformative")


res= list(dtestm, dttestlm)
names(res)=c("ML_PC_KStest", "ML_PC_LM")
writexl::write_xlsx(res, paste0(pcDir, "/PC_Effectiveness_ML_Normal_NormalPreneo.xlsx"))
```




```{r}
 pc <- Embeddings(ml, reduction = "pca")[, 1:19] %>% as.data.frame()
  identical(rownames(pc), rownames(ml@meta.data))
  pc$condition= ml$Condition
  pc$sample= ml$sample
  
 pdf(paste0(pcDir, "/Barplot_Distrib_ML_PC_notFlipped.pdf"), width = 15, height = 8)
for( i in seq(0, 30,10)[-1]){
  for(j in c(4,12,13,19)){
    
    dt <- reshape2::melt(table(cut(pc[,j], seq(min(pc[,j]), max(pc[j]), i), include.lowest = TRUE), pc$condition)) %>% as.data.frame()
    dt$Var2= factor(dt$Var2, levels=c("Normal BRCA1+/- pre-neoplastic", "Normal"))
    #pdf(paste0(RdataDir, "/Barplot_Distrib_LP_PC_", j, "_BinSize_",i,".pdf"), width = 8, height = 8)
    
    p <-  ggplot(dt,aes(x= Var1, y=value, fill=Var2)) + geom_bar(stat="identity") + themplot+ theme(legend.position = "none") + xlab(paste0("PC_",j,"_BinSize=", i)) + scale_fill_manual(values=rev(met.brewer(name="Egypt",n=4,type="discrete"))[c(1,2)]) 
    print(p)
#################
     dt <- reshape2::melt(table(cut(pc[,j], seq(min(pc[,j]), max(pc[j]), i), include.lowest = TRUE), pc$sample)) 
    
        g <- dt %>% ggplot(aes(x= Var1, y=value, fill=Var2)) + geom_bar(stat="identity") + themplot+ theme(legend.position = "none") + xlab(paste0("PC_",j,"_BinSize=", i)) + scale_fill_manual(values=rev(met.brewer(name="VanGogh2",n=length(unique(basal$orig.ident)),type="continuous"))) 
    print(g)

  }
}
 dev.off()
 
```


```{r}
qs::qsave(lp, paste0(RdataDir, "LP_Equilibrated_New_Final.qs"))
qs::qsave(basal, paste0(RdataDir, "Basal_New_Final.qs"))
qs::qsave(ml, paste0(RdataDir, "ML_Equilibrated_New_Final.qs"))
## 
lp= qs::qread(paste0(RdataDir, "LP_Equilibrated_New_Final.qs"))
basal = qs::qread(paste0(RdataDir, "Basal_New_Final.qs"))
ml= qs::qread(paste0(RdataDir, "ML_Equilibrated_New_Final.qs"))
```

Plot the PCA fro basal and ML:
```{r}
png(paste0(RdataDir, "/PCA_BASAL_Condition.png"), height = 1200, width = 1200, res = 300)
p <-  DimPlot(basal,  pt.size = 1.2, label = FALSE, label.size = 2, group.by = "Condition", reduction= "pca", dims = c(4,8), cols = rev(met.brewer(name="Egypt",n=length(unique(e1$Condition)),type="discrete"))[c(2,1)] ) +  NoAxes() + NoLegend() + ggtitle("Basal")
 p[[1]]$layers[[1]]$aes_params$alpha = 0.4
  print(p)
dev.off()
##############
png(paste0(RdataDir, "/PCA_ML_equilibrated_Condition.png"), height = 1200, width = 1200, res = 300)
p <-  DimPlot(ml,  pt.size = 1.2, label = FALSE, label.size = 2, group.by = "Condition", reduction= "pca", dims = c(12,14), cols = rev(met.brewer(name="Egypt",n=length(unique(e1$Condition)),type="discrete"))[c(2,1)] ) +  NoAxes() + NoLegend() + ggtitle("ML equilibrated")
 p[[1]]$layers[[1]]$aes_params$alpha = 0.4
  print(p)
dev.off()
##############
png(paste0(RdataDir, "/PCA_LP_equilibrated_Condition.png"), height = 1200, width = 1200, res = 300)
p <-  DimPlot(lp,  pt.size = 1.2, label = FALSE, label.size = 2, group.by = "Condition", reduction= "pca", dims = c(2,3), cols = rev(met.brewer(name="Egypt",n=length(unique(e1$Condition)),type="discrete"))[c(2,1)] ) +  NoAxes() + NoLegend() + ggtitle("ML equilibrated")
 p[[1]]$layers[[1]]$aes_params$alpha = 0.4
  print(p)
dev.off()



```





Construct a table of the most informative PCs:
```{r}
dtest$type="LP"
dtestb$type="Basal"
dtestm$type= "ML"
```

```{r}
colnames(dtestb)= c(colnames(dtestb)[1:4],"Informativity", colnames(dtestb)[6:7],"type")
colnames(dtestm)= c(colnames(dtestm)[1:4],"Informativity", colnames(dtestm)[6:7],"type")

res= PC_important_PEA(obj= lp,dtest=dtest, subtype= "LP") 
 resb= PC_important_PEA(obj= basal,dtest=dtestb, subtype= "Basal")
resm= PC_important_PEA(obj= ml,dtest=dtestm, subtype= "ML")
```



```{r}
tmp= rbind(dtest %>% dplyr::filter(Informativity== "Informative"),dtestb %>% dplyr::filter(Informativity== "Informative"),dtestm %>% dplyr::filter(Informativity== "Informative") )
writexl::write_xlsx(tmp, paste0(pcDir, "/TopInformative_PCs_AllSubtypes.xlsx"))
tmp=readxl::read_excel(paste0(pcDir, "/TopInformative_PCs_AllSubtypes.xlsx"))
##Plot the evolution:
tmp= tmp[order(tmp$VarExplained, decreasing = TRUE),]
tmp$rank= seq(1:nrow(tmp))

pdf(paste0(pcDir, "/InformativePCs_Rank_Type.pdf"), height = 8, width = 8)
tmp %>% ggplot() + geom_point(aes(x=rank, y= VarExplained, color=type, size=3)) + themplot + xlab("Informative PCs Rank") + ylab("% Explained Variance within the epithelial subtype")
dev.off()
##Do It By Barplot:

pdf(paste0(pcDir, "/InformativePCs_Rank_Type_BARPLOT.pdf"), height = 8, width = 8)
tmp %>% ggplot() + geom_bar(aes(x=rank, y= VarExplained, fill=type), stat="identity") + themplot + xlab("Informative PCs Rank") + ylab("% Explained Variance within the epithelial subtype")
dev.off()

```

Combine the genes contributing to the informative PCs for the LP and perform pathway enrichment analysis:
```{r}
tmp1 <- lp@reductions$pca@feature.loadings[,2] %>% as.data.frame() %>% rownames_to_column("gene")
colnames(tmp1)= c("gene","eigen")
tmp1 = tmp1 %>% dplyr::filter(eigen > (0.015)) 
 ###
tmp2<- lp@reductions$pca@feature.loadings[,3] %>% as.data.frame() %>% rownames_to_column("gene")
colnames(tmp2)= c("gene","eigen")
tmp2 = tmp2 %>% dplyr::filter(eigen > (0.02))
 ###
tmp3<- lp@reductions$pca@feature.loadings[,8] %>% as.data.frame() %>% rownames_to_column("gene")
colnames(tmp3)= c("gene","eigen")
tmp3 = tmp3 %>% dplyr::filter(eigen < (-0.05))
 gn <- unique(c(tmp1$gene, tmp2$gene, tmp3$gene) )
#~~~~~~~~~~~~~~~~~~~~~~~~~~

```

Also Take the C2 gene list:
```{r}
  m_t2g <- msigdbr::msigdbr(species = "Homo sapiens", category = "H") %>% dplyr::select(gs_name, gene_symbol) %>% as.data.frame()
  
peaLP= clusterProfiler::enricher(gn , TERM2GENE=m_t2g) @result %>% dplyr::filter(., p.adjust< 0.01)%>% .[,c(1,6,8)]
 peaLP$geneID= gsub("/",",", peaLP$geneID)
 ######################
   m_t2g <- msigdbr::msigdbr(species = "Homo sapiens", category = "C2") %>% dplyr::select(gs_name, gene_symbol) %>% as.data.frame()
  
peaC2LP= clusterProfiler::enricher(gn , TERM2GENE=m_t2g) @result %>% dplyr::filter(., p.adjust< 0.001)%>% .[,c(1,6,8)]
 peaC2LP$geneID= gsub("/",",", peaC2LP$geneID)
```

Keep the info:
```{r}
l= list(peaLP, peaC2LP)
names(l)=c("LP_Equilibrated_ContribPC_H","LP_Equilibrated_ContribPC_C2")
writexl::write_xlsx(l, paste0(pcDir, "/PEA_H_C2_TopGenes_Contrib_HighPreneo.xlsx"))
rm(l)
l= readxl::read_excel(paste0(pcDir, "/PEA_H_C2_TopGenes_Contrib_HighPreneo.xlsx"))
peaLP=l
peaC2LP=readxl::read_excel(paste0(pcDir, "/PEA_H_C2_TopGenes_Contrib_HighPreneo.xlsx"), sheet = 2)
```


```{r}
p2 <- peaC2LP %>% dplyr::filter( ID %in% c( grep("MAMMARY", peaC2LP$ID, value = TRUE), grep("BREAST_CANCER", peaC2LP$ID, value = TRUE),grep("SENESCENCE", peaC2LP$ID, value = TRUE)  )) %>% dplyr::filter(! ID %in% grep("_DN", peaC2LP$ID, value = TRUE))

p2$type= ifelse(p2$ID %in% grep("MAMMARY", p2$ID, value = TRUE), "Mammary", ifelse(p2$ID %in% grep("BREAST_CANCER", p2$ID, value = TRUE), "Breast", "Senescence"))

pl <- peaC2LP[which(peaC2LP$ID %in% c("LEI_MYB_TARGETS", "ONDER_CDH1_TARGETS_2_DN","WU_CELL_MIGRATION")),]
pl$type="others"
###
p2=rbind(p2, pl)
p2= p2 %>% dplyr::filter(p.adjust < 3.008500e-04)
##take the top3 most signficant:
p <- p2 %>% group_by(type) %>%  top_n(n = 3, wt = (-p.adjust))
h <- peaLP %>%  top_n(n = 3, wt = (-p.adjust))
h=h[-4,]
h$type= "Hallmark" 
###group both:
p=rbind(p,h)
###
rm(h)
rm(pl)
rm(p2)
rm(pc2)

pdf(paste0(pcDir, "/PEA_C2_H_LP_Equilibrated_AllInformativePCs.pdf"), width = 15, height = 10)
reshape2::melt(p[, c(1,2,4)]) %>% ggplot(aes(x=reorder(factor(ID), -log10(value)), y = -log10(value))) + 
   # facet_wrap(~ type) 
   geom_bar(aes(fill = factor(type)), stat = "identity")  + coord_flip() + theme(legend.position = NULL) + ylab("") + themplot
dev.off()
```

### FIGURE IMPROVEMENT 

```{r}
res_LP=readxl::read_xlsx( paste0(pcDir, "/nPC20_Effectiveness_LP_Normal_NormalPreneo.xlsx"))
res_ML=readxl::read_xlsx(paste0(pcDir, "/PC_Effectiveness_ML_Normal_NormalPreneo.xlsx"))
res_Bsl= readxl::read_xlsx(paste0(pcDir, "/PC_Effectiveness_Basal_Normal_NormalPreneo_PC20.xlsx"))
######
#ifelse((dtest$KStest < quantile(dtest$KStest, 0.8) & dtest$Dvalue > quantile(dtest$Dvalue, 0.8))
```

Plot barplot representation:
```{r}
res_LP$KStest= ifelse(res_LP$KStest== 0, 10^(-20), res_LP$KStest)
res_LP$logKS= -log10(res_LP$KStest)
###
res_ML$KStest= ifelse(res_ML$KStest== 0, 10^(-20), res_ML$KStest)
res_ML$logKS= -log10(res_ML$KStest)
###
res_Bsl$KStest= ifelse(res_Bsl$KStest== 0, 10^(-20), res_Bsl$KStest)
res_Bsl$logKS= -log10(res_Bsl$KStest)

pdf(paste0(pcDir, "/Barplot_PC_Informativity_LP_Bsl.pdf"))

p <- res_LP %>% ggplot(aes(x=reorder(PC, VarExplained), y= varExplained, fill= Informativity)) + geom_bar(stat="identity") +  
  geom_hline(yintercept=quantile(res_LP$Dvalue, 0.8), linetype="dashed", color = "red")+ coord_flip() + scale_fill_viridis(option="B", direction= -1) + themplot + ggtitle("LP")
print(p)

g <- res_ML %>% ggplot(aes(x=reorder(PC, logKS), y= Dvalue, fill= logKS)) + geom_bar(stat="identity") +  
  geom_hline(yintercept=quantile(res_ML$Dvalue, 0.8), linetype="dashed", color = "red")+ coord_flip() + scale_fill_viridis(option="B", direction= -1) + themplot + ggtitle("ML")

h <- res_Bsl %>% ggplot(aes(x=reorder(PC, logKS), y= Dvalue, fill= logKS)) + geom_bar(stat="identity") +  
  geom_hline(yintercept=quantile(res_Bsl$Dvalue, 0.8), linetype="dashed", color = "red")+ coord_flip() + scale_fill_viridis(option="B", direction= -1) + themplot + ggtitle("Basal")
print(p)
print(g)
print(h)
dev.off()
```

```{r}
pdf(paste0(pcDir, "/Barplot_PC_Informativity_LP_Bsl.pdf"))

p <- res_LP %>% ggplot(aes(x=reorder(PC, VarExplained), y= VarExplained, color= Informativity)) + geom_bar(stat="identity", fill= "white") + coord_flip()  + themplot + ggtitle("LP") + scale_color_manual(values=c("black", "grey"))


g <- res_ML  %>% ggplot(aes(x=reorder(PC, VarExplained), y= VarExplained, color= take)) + geom_bar(stat="identity", fill= "white") + coord_flip()  + themplot + scale_color_manual(values=c("black", "grey")) + ggtitle("ML")

h <- res_Bsl %>% ggplot(aes(x=reorder(PC, VarExplained), y= VarExplained, color= take)) + geom_bar(stat="identity", fill= "white") + coord_flip()  + themplot+ scale_color_manual(values=c("black", "grey"))+ ggtitle("Basal")
print(p)
print(g)
print(h)
dev.off()
```
