---
title: "Step4_TCGA_MetadataComparison"
author: "Melissa"
date: "04/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
mainDir="C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Tumor_BRCA_p53"
mainDir <- here::here() # the MainDir should be the root of Tumor_BRCA_P53:
source(knitr::purl(file.path(mainDir,"Scripts","scRNAseq","mouse","TCGA","Global_Variables_TCGA.Rmd"), quiet=TRUE))
```



```{r}
tcga_br= readxl::read_xlsx(paste0(inputDir, "/Processed_Metadata.xlsx"))
normalized_counts= qs::qread(paste0(RDatadir, "/Normalized_Counts_Matreix_BASALvsOthers.qs"))
```

```{r}
tcga_br=readxl::read_excel("C:/Users/Melissa S/Desktop/PROJECTS/BRCA1_Tumorigenesis/Tumor_BRCA_p53/input/human/publicDatasets/inputTCGA_Breast/Processed_Metadata.xlsx")

normalized_counts= qs::qread( "C:/Users/Melissa S/Desktop/PROJECTS/BRCA1_Tumorigenesis/Tumor_BRCA_p53/output/human/public_data/TCGA/RData/Normalized_Counts_Matreix_BASALvsOthers.qs")

preT=UCell::ScoreSignatures_UCell()
```


Get only the TNBC patients
```{r}
tn <- tcga_br %>% dplyr::filter(Subtype=="BRCA_Basal")
normtn= normalized_counts[, which(colnames(normalized_counts) %in% tn$Sample.ID)]
dim(normtn)
```

```{r}
dt <- normtn[rownames(normtn)%in% c("FN1", "ITGB1", "SERPINE2", "CDKN2A", "CDK4", "RB1","E2F1", "MKI67"),] %>% t() %>% as.data.frame()
md = tcga_br %>% filter(Sample.ID %in% rownames(dt))
dt=dt[md$Sample.ID,]
stopifnot(identical(rownames(dt), md$Sample.ID))
dt$type=md$Subtype
```


Test the stage of the disease:
```{r}
rownames(scores)=scores$sample

tnscores= scores %>% dplyr::filter(rownames(scores) %in% tn$Sample.ID)
identical(rownames(tnscores), tn$Sample.ID)

tnscores$stage= tn$Neoplasm.Disease.Stage.American.Joint.Committee.on.Cancer.Code

tnscores$simplifiedStage= ifelse(tnscores$stage== "STAGE I"| tnscores$stage== "STAGE IA", "1",
                                 ifelse(tnscores$stage== "STAGE II"| tnscores$stage== "STAGE IIA" | tnscores$stage== "STAGE IIB", "2", 
                                  ifelse(tnscores$stage== "STAGE IIIA"| tnscores$stage== "STAGE IIIB" | tnscores$stage== "STAGE IIIC" , "3", "4")))
##################
tnscores=tnscores[which(is.na(tnscores$simplifiedStage)== FALSE),]
tnscores= tnscores %>% filter(!simplifiedStage== "4")
tnscores$simplifiedStage=as.factor(tnscores$simplifiedStage)    
```

```{r}
tnscores$ss= ifelse(tnscores$simplifiedStage=="1", "early", "late")
wilcox_test(data=tnscores, formula= p16UP_signature_UCell ~ as.factor(ss))

pdf(paste0(RdataDir,"/Boxplot_P16Signature_UCell_logFC1_Simplified_Disease_Stage.pdf"), height = 5, width = 5)
 tnscores %>% ggplot(aes(x=simplifiedStage, y=p16UP_signature_UCell, fill= "grey")) + geom_boxplot(aes(x=simplifiedStage, y=p16UP_signature_UCell, fill=NULL))  +theme_bw(base_size = 12) + ylab("UCELL_Score_Up_Sigature_P16")  + theme(legend.position = "none")+ gtheme +geom_jitter(color="black", size=3, alpha=0.4, fill= "grey") +   stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "1",aes(label=..p.adj..))
 dev.off()
```


```{r}
writexl::write_xlsx(tnscores, paste0(RDatadir, "/Metadata_forSurvivalanalysis.xlsx"))
writexl::write_xlsx(tn, paste0(RDatadir, "/Metadata_forSurvivalanalysis_OnlyBRCA_Basal.xlsx"))
```

####################################################################################################################
### Compute BRCA1 Comparison:
```{r}
brca1_mutation <- read.delim("C:/Users/Melissa S/Downloads/alterations_across_samples (1).tsv")
p16_mutation <- read.delim("C:/Users/Melissa S/Downloads/alterations_across_samples (2).tsv")
brca2_mutation <- read.delim("C:/Users/Melissa S/Downloads/alterations_across_samples (3).tsv")

stopifnot(identical(brca1_mutation$Sample.ID, brca2_mutation$Sample.ID))
brca1_mutation=brca1_mutation[match( brca2_mutation$Sample.ID,brca1_mutation$Sample.ID),]
p16_mutation=p16_mutation[match( brca1_mutation$Sample.ID,p16_mutation$Sample.ID),]
stopifnot(identical(brca1_mutation$Sample.ID, p16_mutation$Sample.ID))
####
brca1_mutation$brca1alt= brca1_mutation$Altered
brca2_mutation$brca2alt= brca2_mutation$Altered
p16_mutation$p16alt= p16_mutation$Altered

mutation= cbind(brca1_mutation$Sample.ID,brca1_mutation$brca1alt, brca2_mutation$brca2alt, p16_mutation$p16alt)
mutation=as.data.frame(mutation)
colnames(mutation)=c("Sample.ID", "BRCA1_alt", "BRCA2_alt", "P16_alt")

metadata= merge(tcga_br, mutation, by="Sample.ID")
#####
DRD <- read.delim2("C:/Users/Melissa S/Downloads/alterations_across_samples (4).tsv")
DRD=DRD[,1:16]
DRD[,5:16]= apply(DRD[,5:16], 2, function(x) {ifelse(x== "no alteration", 0, 1)})
#merge with the metadata:

DRD= DRD %>% dplyr::filter(Sample.ID %in% scores$sample)
scores= scores %>% dplyr::filter(sample %in%  DRD$Sample.ID)
DRD=DRD[match( scores$sample,DRD$Sample.ID),]
stopifnot(identical(scores$sample, DRD$Sample.ID))
scores= cbind(scores, DRD)

```

```{r}
rownames(scores)=scores$sample
metadata= metadata %>% dplyr::filter(Sample.ID %in% scores$sample)
scores= scores %>% dplyr::filter(sample %in%  metadata$Sample.ID)

scores= cbind(scores, metadata)
######################################
scores$type_brca1= paste0(scores$subtype, "_", scores$BRCA1_alt)

scores$type_brca2= paste0(scores$subtype, "_", scores$BRCA2_alt)


scores %>%  ggplot(aes(x=type_brca1, y=p16UP_signature_UCell)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=scores,aes(x=subtype, y=p16UP_signature_UCell,color= type_brca1), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "BRCA_Basal",aes(label=..p.adj..))  +   gtheme 
```

```{r}
rownames(scores)= scores$sample
writexl::write_xlsx(scores, paste0(SIGdir, "/Signature_Scores_All_PANCANCER_p16_EMT_Apop.xlsx"))
```

```{r}
scores <- readxl::read_excel(paste0(SIGdir, "/Signature_Scores_All_PANCANCER_p16_EMT_Apop.xlsx")) %>% as.data.frame()

DRD2 <- read.delim2("C:/Users/Melissa S/Downloads/alterations_across_samples (4).tsv")
metadata=DRD2[,c(colnames(DRD2)[1:4],grep("HOMDEL", colnames(DRD2[,5:ncol(DRD2)]), value = TRUE))]
metadata= metadata %>% dplyr::filter(Sample.ID %in% scores$sample)
scores= scores %>% dplyr::filter(sample %in%  metadata$Sample.ID)
###
metadata= metadata %>% dplyr::filter(Sample.ID %in% scores$sample)
scores= scores %>% dplyr::filter(sample %in%  metadata$Sample.ID)
metadata=metadata[match( scores$sample,metadata$Sample.ID),]
stopifnot(identical(scores$sample, metadata$Sample.ID))
scores= cbind(scores, metadata)

scores[,10:21]= apply(scores[,10:21], 2, function(x) {
  ifelse(x== "no alteration", 0, 
         ifelse(x== "AMP", 1, 
                ifelse(x== "HOMDEL (driver)", (-1),
                       NA)))})
writexl::write_xlsx(scores, paste0(SIGdir, "/Metadata_Alterations_HRDgenes.xlsx"))
```


##### START WORKING ON THE GENETIC INFORMATION ##########################
```{r}
scores <- readxl::read_excel(paste0(SIGdir, "/Signature_Scores_All_PANCANCER_p16_EMT_Apop.xlsx")) %>% as.data.frame()
DRD2 <- read.delim2("C:/Users/Melissa S/Downloads/alterations_across_samples (4).tsv")

metadata=DRD2[,c(colnames(DRD2)[1:4],grep("HOMDEL", colnames(DRD2[,5:ncol(DRD2)]), value = TRUE))]
metadata= metadata %>% dplyr::filter(Sample.ID %in% scores$sample)
scores= scores %>% dplyr::filter(sample %in%  metadata$Sample.ID)
###
metadata= metadata %>% dplyr::filter(Sample.ID %in% scores$sample)
scores= scores %>% dplyr::filter(sample %in%  metadata$Sample.ID)
metadata=metadata[match( scores$sample,metadata$Sample.ID),]
stopifnot(identical(scores$sample, metadata$Sample.ID))
scores= cbind(scores, metadata)

scores[,10:21]= apply(scores[,10:21], 2, function(x) {
  ifelse(x== "no alteration", 0, 
         ifelse(x== "AMP", 1, 
                ifelse(x== "HOMDEL (driver)", (-1),
                       NA)))})
writexl::write_xlsx(scores, paste0(SIGdir, "/Metadata_Alterations_HRDgenes_Discrete.xlsx"))
```

Representation of the alterations: Method2
```{r}
scores <- readxl::read_excel(paste0(SIGdir, "/Signature_Scores_All_PANCANCER_p16_EMT_Apop.xlsx")) %>% as.data.frame()

DRD2 <- read.delim2("C:/Users/Melissa S/Downloads/alterations_across_samples (6).tsv")

metadata=DRD2[,1:8]
metadata= metadata %>% dplyr::filter(Sample.ID %in% scores$sample)
scores= scores %>% dplyr::filter(sample %in%  metadata$Sample.ID)
###
metadata= metadata %>% dplyr::filter(Sample.ID %in% scores$sample)
scores= scores %>% dplyr::filter(sample %in%  metadata$Sample.ID)
metadata=metadata[match( scores$sample,metadata$Sample.ID),]
stopifnot(identical(scores$sample, metadata$Sample.ID))
scores= cbind(scores, metadata)

scores[,10:13]= apply(scores[,10:13], 2, function(x) {
  ifelse(x== "no alteration", 0, 
         ifelse(x== "AMP", 1, (-1)))
                })

scores$BRCA1status= scores$BRCA1
scores$BRCA1status= ifelse(scores$BRCA1status== (-1), "Alt", "NoAlt")
scores$BRCAsubtype= ifelse(scores$subtype== "BRCA_Basal", paste0("Bsl_",scores$BRCA1status), gsub( "BRCA_", "",scores$subtype))
#Do the same with BRCA2:
scores$BRCA2status= scores$BRCA2
scores$BRCA2status= ifelse(scores$BRCA2status== (-1), "Alt", "NoAlt")
scores$BRCA2subtype= ifelse(scores$subtype== "BRCA_Basal", paste0("Bsl_",scores$BRCA2status), gsub( "BRCA_", "",scores$subtype))
##  RAD51C
scores$RAD51Cstatus= scores$RAD51C
scores$RAD51Cstatus= ifelse(scores$RAD51Cstatus== (-1), "Alt", "NoAlt")
scores$RAD51Csubtype= ifelse(scores$subtype== "BRCA_Basal", paste0("Bsl_",scores$RAD51Cstatus), gsub( "BRCA_", "",scores$subtype))
##
scores$TP53status= scores$TP53
scores$TP53status= ifelse(scores$TP53status== (-1), "Alt", "No_p53Alt")
scores$TP53subtype= ifelse(scores$subtype== "BRCA_Basal", paste0("Bsl_",scores$TP53status), gsub( "BRCA_", "",scores$subtype))
```

```{r}
scores2= readxl::read_excel("C:/Users/Melissa S/Desktop/PROJECTS/BRCA1_Tumorigenesis/Tumor_BRCA_p53/output/human/public_data/TCGA/Signatures/Metadata_Alterations_HRDgenes_Discrete.xlsx")
stopifnot(identical(scores2$sample, scores$sample))
scores2$p16UP_signature_UCell=scores$p16UP_signature_UCell

scores2[,10:13]= apply(scores2[,10:13], 2, function(x) {
  ifelse(x== "no alteration", 0, 
         ifelse(x== "AMP", 1, (-1)))
                })

scores$BRCA1status= scores2$BRCA1..HOMDEL
scores$BRCA1status= ifelse(scores$BRCA1status== (-1), "Alt", "NoAlt")

scores$BRCAsubtype= ifelse(scores$subtype== "BRCA_Basal", paste0("Bsl_",scores$BRCA1status), gsub( "BRCA_", "",scores$subtype))
```



```{r}
pdf(paste0(GENEdir, "/Boxplot_Alterations_HDRgenes_P16signatures_2.pdf"), height = 5, width = 5)

scores %>%  ggplot(aes(x=BRCAsubtype, y=p16UP_signature_UCell)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=scores,aes(x=BRCAsubtype, y=p16UP_signature_UCell,color= BRCAsubtype),size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_NoAlt",aes(label=..p.adj..))  + stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Alt",aes(label=..p.adj..))  + themplot +gtheme + theme(legend.position = "none")
#####
scores %>%  ggplot(aes(x=BRCA2subtype, y=p16UP_signature_UCell)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=scores,aes(x=BRCA2subtype, y=p16UP_signature_UCell,color= BRCA2subtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_NoAlt",aes(label=..p.adj..)) +    themplot +gtheme + theme(legend.position = "none")
###
scores %>%  ggplot(aes(x=RAD51Csubtype, y=p16UP_signature_UCell)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=scores,aes(x=RAD51Csubtype, y=p16UP_signature_UCell,color= RAD51Csubtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Alt",aes(label=..p.adj..)) +    themplot +gtheme + theme(legend.position = "none")
## CDK12
scores %>%  ggplot(aes(x=scores$TP53subtype, y=p16UP_signature_UCell)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=scores,aes(x=scores$TP53subtype, y=p16UP_signature_UCell,color= scores$TP53subtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Alt",aes(label=..p.adj..)) +    themplot +gtheme + theme(legend.position = "none")
dev.off()
```

Create a combination between BRCA1/P53 mutations:
```{r}
scores$p53_BRCA1=paste0(scores$TP53subtype, "_", scores$BRCAsubtype)
scores$p53_BRCA1= dplyr::recode(scores$p53_BRCA1, "LumA_LumA"= "LumA", "Her2_Her2"="Her2", "LumB_LumB"="LumB",               "Normal_Normal" ="Normal", "Bsl_Alt_Bsl_NoAlt"="Bsl_p53mBrca1","Bsl_Alt_Bsl_Alt"="Bsl_p53mBrca1m",         "Bsl_No_p53Alt_Bsl_NoAlt"= "Bsl_NoAlt", "Bsl_No_p53Alt_Bsl_Alt"= "Bsl_p53_Brca1m" )
#############################################################

pdf(paste0(GENEdir, "/P53_mutations_DNArepair_Genes.pdf"), height = 5, width = 7)
scores %>%  ggplot(aes(x=BRCAsubtype, y=p16UP_signature_UCell)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=scores,aes(x=BRCAsubtype, y=p16UP_signature_UCell,shape= TP53subtype, color= TP53subtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Alt",aes(label=..p.adj..)) +    themplot +gtheme + theme(legend.position = "right") + ggtitle("TP53Alt from CBIO in BRCA1 alt")
#####
scores %>%  ggplot(aes(x=BRCA2subtype, y=p16UP_signature_UCell)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=scores,aes(x=BRCA2subtype, y=p16UP_signature_UCell,shape= TP53subtype, color= TP53subtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Alt",aes(label=..p.adj..)) +    themplot +gtheme + theme(legend.position = "right") + ggtitle("TP53Alt from CBIO in BRCA2 alt")

scores %>%  ggplot(aes(x=RAD51Csubtype, y=p16UP_signature_UCell)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=scores,aes(x=RAD51Csubtype, y=p16UP_signature_UCell,shape= TP53subtype, color= TP53subtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Alt",aes(label=..p.adj..)) +    themplot +gtheme + theme(legend.position = "right") + ggtitle("TP53Alt from CBIO in RAD51C alt")
dev.off()
```


BRCANESS:
```{r}
scores$sample= gsub("-", "_", scores$sample)
brca_data <- read_excel("C:/Users/Melissa S/Downloads/Sample_plan_clinical_patient_brca_NEW_celine_cut.xlsx")
brca_data$sample_ID=   str_sub(brca_data$sample_ID,1,nchar(brca_data$sample_ID)-1)
brca_data= brca_data %>% dplyr::filter(sample_ID %in% scores$sample)
##ISOLATE BRCAness status:
sub_brca= brca_data[, c(1,7,8,19,21,24)]
scores= merge(scores, sub_brca, by.x= "sample", by.y="sample_ID")

table(is.na(scores$BRCAness_code), scores$subtype)
```

Remove samples with NA values for BRCAness:
```{r}
dt2= scores[-which(is.na(scores$BRCAness_code== TRUE)),]

pdf(paste0(SIGdir, "/Boxplot_BRCAness_code_Subtype_PANCANCER.pdf"), height = 5, width = 5)
p <- dt2 %>% ggplot(aes(x= subtype, y= BRCAness_code)) + geom_boxplot() + gtheme + themplot
print(p)
dev.off()
###
dt2$BRCAnessBinary= ifelse(dt2$BRCAness_code == 0, "No_BRCAness", "BRCAness")

dt2$BRCAnessSubtype= ifelse(dt2$subtype.x== "BRCA_Basal", paste0("Bsl_", dt2$BRCAnessBinary), gsub("BRCA_", "",dt2$subtype.x))

```

```{r}
pdf(paste0(GENEdir, "/Boxplot_P16Signature_Pancancer_BRCAness.pdf"), height = 5, width = 5)
dt2 %>%  ggplot(aes(x=BRCAnessSubtype, y=p16UP_signature_UCell)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=dt2,aes(x=BRCAnessSubtype, y=p16UP_signature_UCell,color= BRCAnessSubtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_BRCAness",aes(label=..p.adj..))  +    themplot  + theme(legend.position = "none")
dev.off()
```

```{r}
pdf(paste0(GENEdir, "/ScatterPlot_P16Signature_BRCAness_BRCA1alterations.pdf"), height = 10, width = 15)
p <- dt2 %>% ggplot(aes(x=p16UP_signature_UCell , y=  BRCAness_code, fill=BRCAnessSubtype, col= BRCAnessSubtype)) + geom_point() + scale_color_manual(values=c("red", "blue", "grey", "grey", "grey", "grey")) + gtheme + themplot
print(p)
dev.off()
```

#Add the info on TP53 mutation:
```{r}
length(intersect(dt2$sample, scores$sample))
scores$sample= gsub("-", "_", scores$sample)

dt2= merge(dt2, scores, by="sample")
```

```{r}
pdf(paste0(GENEdir, "/Boxplot_BRCAness_BRCA1Mutation_Thr0.pdf"), height = 5, width = 7)
dt2 %>%  ggplot(aes(x=BRCAnessSubtype, y=p16UP_signature_UCell.x)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")    +geom_jitter(data=dt2,aes(x=BRCAnessSubtype, y=p16UP_signature_UCell.x,color= BRCAnessSubtype, shape= BRCAsubtype.x), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_BRCAness",aes(label=..p.adj..))  +    themplot  + theme(legend.position = "right")
dev.off()
```





BRCA1_BRCA2_RAD51C_genetic_alteration:
```{r}
dt2$BRCA1_BRCA2_RAD51C_genetic_alteration= ifelse(is.na(dt2$BRCA1_BRCA2_RAD51C_genetic_alteration)==TRUE, "NotAltered", "Altered")


dt2$HDRsubtype= ifelse(dt2$subtype=="BRCA_Basal", paste0("Bsl_", dt2$BRCA1_BRCA2_RAD51C_genetic_alteration), gsub( "BRCA_", "",dt2$subtype))
###
pdf(paste0(GENEdir, "/Boxplot_P16Signature_Pancancer_BRCA12RAD51_p53mut.pdf"), height = 5, width = 7)

dt2 %>%  ggplot(aes(x=HDRsubtype, y=p16UP_signature_UCell.x)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=dt2,aes(x=HDRsubtype, y=p16UP_signature_UCell.x,color= HDRsubtype, shape= TP53subtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Altered",aes(label=..p.adj..))  +    themplot  + theme(legend.position = "right") + ggtitle("BRCA1/2_RAD51C genetic alteration")
dev.off()
```


Summary Proven HRD:
```{r}
dt2$Summary_provenHRD= ifelse(is.na(dt2$Summary_provenHRD)==TRUE, "NotAltered", "Altered")
dt2$HDRsubtype= ifelse(dt2$subtype=="BRCA_Basal", paste0("Bsl_", dt2$Summary_provenHRD), gsub( "BRCA_", "",dt2$subtype))
###
pdf(paste0(GENEdir, "/Boxplot_P16Signature_Pancancer_ProvenHRD_P53mut.pdf"), height = 5, width = 5)

dt2 %>%  ggplot(aes(x=HDRsubtype, y=p16UP_signature_UCell.x)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")  + ylab("UCELL_Up in P16 Signature")  +geom_jitter(data=dt2,aes(x=HDRsubtype, y=p16UP_signature_UCell.x,color= HDRsubtype, shape= TP53subtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Altered",aes(label=..p.adj..))  +  stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Altered",aes(label=..p.adj..))  + themplot  + theme(legend.position = "none") + ggtitle("Summary_provenHRD")
dev.off()
```


#--------------------------Add CDKN2A-----------------------------------##
```{r}
dt <- normalized_counts[rownames(normalized_counts)%in% c("FN1", "CDKN2A", "CDK4"),] %>% t() %>% as.data.frame()
dt$sample= rownames(dt)
dt$sample= gsub("-", "_", dt$sample)

data= merge(dt, scores, by="sample")
```


```{r}
pdf(paste0(SIGdir, "/Boxplot_CDKN2A_BRCAness_provenHRDsubtype_TP53.pdf"), height = 5, width = 5)
data %>%  ggplot(aes(x=BRCAnessSubtype, y=log10(CDKN2A))) + geom_boxplot( fill='#DBD8D8', color="black") +  ylab("log10(CDKN2A)")  +geom_jitter(data=data,aes(x=BRCAnessSubtype, y=log10(CDKN2A),color= BRCAnessSubtype, shape= TP53subtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_BRCAness",aes(label=..p.adj..))  +    themplot +gtheme + theme(legend.position = "none")
#####

data %>%  ggplot(aes(x=HDRsubtype, y=log10(CDKN2A))) + geom_boxplot( fill='#DBD8D8', color="black")   + ylab("log10(CDKN2A)")  +geom_jitter(data=data,aes(x=HDRsubtype, y=log10(CDKN2A),color= HDRsubtype, shape= TP53subtype), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_grey()+ stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Altered",aes(label=..p.adj..))  +  stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Bsl_Altered",aes(label=..p.adj..))  + themplot  + theme(legend.position = "none") + ggtitle("Summary_provenHRD")

dev.off()
```
```{r}
pdf(paste0(GENEdir, "/Subtype_Brca1Mutation2_log10Cdkn2a.pdf"), height = 5, width = 5)

data %>%  ggplot(aes(x=subtype, y=log10(CDKN2A))) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("log10(CDKN2A)")  + geom_jitter(data=data,aes(x=subtype, y=log10(CDKN2A), color= BRCAsubtype, shape= BRCA1status), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_manual(values=c("red", "#A19999", "#C4BABA", "#C2BABA", "#D6D2D2", "#E8E8E8")) +  themplot   + ggtitle("Brca1 alterations")  + theme(legend.position = "none") #,shape= BRCAsubtype
dev.off()

x= data %>% dplyr::filter(subtype=="BRCA_Basal")
x$logp16= log10(x$CDKN2A)

stat.test <- x %>% 
  wilcox_test(CDKN2A ~ BRCAsubtype) %>%
  add_significance()
stat.test
```





```{r}
########################
brca1 <- read.delim("C:/Users/Melissa S/Desktop/PROJECTS/BRCA1_Tumorigenesis/REVIEW/output/mouse/invivo/TCGA/mutations_BRCA1.txt")
length(intersect(brca1$SAMPLE_ID, scores2$sample))
brca1= brca1 %>% dplyr::filter(SAMPLE_ID %in% scores$sample)
scores=scores %>% dplyr::filter(sample %in% brca1$SAMPLE_ID)
stopifnot(identical(brca1$SAMPLE_ID, scores$sample))

scores$BRCA1_mut= brca1$BRCA1
scores$BRCA1_status=ifelse(scores$BRCA1_mut=="WT", "0", "1")
scores$BRCA1= ifelse(scores$BRCA1_status=="1" & scores$subtype=="BRCA_Basal", "mut", "non_mut")
scores$brca1= ifelse(scores$BRCA1_status=="1" & scores$subtype=="BRCA_Basal", "mut", scores$subtype)
```

HERE :::::!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
Subtype by BRCA1 mutation:
```{r}
pdf(paste0(figDir, "/TCGA_PreTumSigSubtype_Brca1Mutation.pdf"), height = 5, width = 5)
scores %>%  ggplot(aes(x=subtype, y=p16UP_signature_UCell)) + geom_boxplot( fill='#DBD8D8', color="black") + ylab("P16Signature_UCell")   +geom_jitter(data=scores,aes(x=subtype, y=p16UP_signature_UCell, color= brca1, shape= brca1), size=1, alpha=0.9) + theme(legend.position = "none")  + scale_color_manual(values=c( "#A19999", "#C4BABA", "#C2BABA", "#D6D2D2", "#E8E8E8","red")) +  themplot   + ggtitle("Brca1 alterations_CompPval 0.014 ")  + theme(legend.position = "none") + stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "BRCA_Basal",aes(label=..p.adj..))
dev.off()

#compare the BRCA1 mutated vs Non-mutated 
library(tidyverse)
library(rstatix)
library(ggpubr)
x= scores %>% dplyr::filter(subtype=="BRCA_Basal")
stat.test <- x %>% 
  wilcox_test(p16UP_signature_UCell ~ BRCA1, alternative = "greater") %>%
  add_significance()
stat.test #pval: 0.0449

##################################################################################
x= scores
x$bsl=ifelse(x$subtype=="BRCA_Basal", "Basal", "No")
stat.test <- x %>% 
  wilcox_test(p16UP_signature_UCell ~ bsl, alternative = "greater") %>%
  add_significance()
stat.test #pval: 0.0244	
```
