---
title: "Step3_TCGA_DiffExpression_Comparison"
author: "Melissa"
date: "04/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
mainDir <- here::here() # the MainDir should be the root of Tumor_BRCA_P53:
source(knitr::purl(file.path(mainDir,"Scripts","scRNAseq","mouse","TCGA","Global_Variables_TCGA.Rmd"), quiet=TRUE))
```


READ BOTHE THE DESEQ OBJECT AND THE METADATA
```{r}
dds= qs::qread(paste0(RDatadir, "/DESEQ_Object_Basal_vsAll.qs"))
```

```{r}
tcga_br= readxl::read_xlsx(paste0(inputDir, "/Processed_Metadata.xlsx"))
```


### SKIP THIS PART
```{r}
res <- results(dds) #sub Others vs BRCA Basal 
######
rownames(res)=gsub("\\|.*", "", rownames(res)) 
res$gene=rownames(res)
res$log2FC= res$log2FoldChange* (-1)
res=as.data.frame(res)
###
write.csv(res, paste0(DEdir, "/DEG_BASAL_vsOTHERS.csv"))
```

Extract the counts and compare expression values between scRNA P16 population and basal like TCGA samples:

```{r}
library(DESeq2)
normalized_counts <- counts(dds, normalized=TRUE)
rownames(normalized_counts) =gsub("\\|.*", "", rownames(normalized_counts ))

qs::qsave(normalized_counts, paste0(RDatadir,"/Normalized_Counts_Matreix_BASALvsOthers.qs"))
##
tcga_br= tcga_br[which(tcga_br$Sample.ID %in% colnames(normalized_counts)),]
```

### START HERE IF YOU ONLY WANT THE PLOT:

```{r}
res= read.csv( paste0(DEdir, "/DEG_BASAL_vsOTHERS.csv"))
normalized_counts= qs::qread(paste0(RDatadir, "/Normalized_Counts_Matreix_BASALvsOthers.qs"))
```

normalized_counts= qs::qread( "C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Tumor_BRCA_p53/output/scRNAseq/TCGA/RData/Normalized_Counts_Matreix_BASALvsOthers.qs")

```{r}
library(Seurat)
#read the invivo epith dataset :
epith= qs::qread(paste0(RDatadir,"/Epithelial_FinalAnnotation.qs"))
epith=qs::qread("C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Tumor_BRCA_p53/output/scRNAseq/invivo/RData/Epithelial_FinalAnnotation.qs")
##Take the most significant and differential genes of the p16 population
p16d= FindMarkers(epith, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP","Avd"), only.pos = FALSE, logfc.threshold = 0.8, min.diff.pct = 0.2) %>% rownames_to_column(var="gene") %>% dplyr::filter(p_val_adj < 0.05) #
p16d$gene=toupper(p16d$gene)
```



take the genes from the DEG of BRCA vs others:

```{r}
common= res[which(rownames(res) %in% p16d$gene),] 
# plot the FC BRCAvs ALl , and P16 vs LP/AVD values of the genes:
dat= p16d[which(p16d$gene %in% rownames(common)),] 
identical(dat$gene, rownames(common))
common=common[dat$gene,]
##
colnames(common)=c(colnames(common)[1:7], "log2FC_Bulk")
dat$log2FC_Bulk= common$log2FC_Bulk


dat$ratio= round(dat$avg_log2FC/dat$log2FC_Bulk)
rm(common)
dat$pn = ifelse(dat$log2FC_Bulk < 0, "neg", "pos")
dat$pn= factor(dat$pn, levels=c("pos","neg"))
```


Visualize the common genes between the scRNAseq and the bulk RNAseq:
```{r}
p <-ggplot(data=dat) + geom_point(aes(x= avg_log2FC, y=log2FC_Bulk)) +
  ggrepel::geom_text_repel(
    data = subset(dat, ratio== 1),
    aes(label = gene,x= avg_log2FC, y=log2FC_Bulk, color=pn),
    size = 5,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines"), max.overlaps = 20) +theme_bw(base_size = 12) + theme(legend.position = "none") + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

png(paste0(DEdir,"/DotPlot_Log2FC_Bulk_singleCell.png"), height = 1500, width = 1500, res=300)
print(p)
dev.off()

pdf(paste0(DEdir, "/DotPlot_Log2FC_Bulk_singleCell.pdf"), height = 8, width = 8)
    print(p)
dev.off()
```

```{r}
p16n= p16d %>% filter(avg_log2FC < 0)  %>%  top_n(39, wt= abs(avg_log2FC))
p16d= p16d %>% filter(avg_log2FC > 0)
p16= rbind(p16d, p16n)
```

```{r}
outputDir2="C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Tumor_BRCA_p53/output/scRNAseq/TCGA/Heatmaps_Survival/"

norm.mat= normalized_counts[, which(colnames(normalized_counts) %in% scores$sample)]
#------------------- Substract the mean per gene-------------------###
norm.mat= norm.mat[which(rownames(norm.mat) %in% p16$gene),]
log_norm.mat= log(norm.mat+1)
log_centered <- sweep(log_norm.mat, MARGIN=1, STATS= rowMeans(log_norm.mat))
log_centered_scaled= apply(log_centered, 1, function(x)  x/(max(x))) 
#---------
pdf(paste0(outputDir2, "Plots_Before.pdf"))
hist(norm.mat)
hist(log(norm.mat))
hist(log_norm.mat)
hist(log_centered)
hist(log_centered_scaled)
dev.off()
#------------
```

Visualize the top genes:
```{r}
library("gplots")
library(ComplexHeatmap)
suppressPackageStartupMessages(library(ComplexHeatmap))

metadata= scores[, c(4,5)]
colours <- list("subtype"=c("BRCA_LumA"="#DCE66A","BRCA_LumB" = "#DBBA34","BRCA_Her2"= "#8BB064","BRCA_Normal"= "#6DD1C5","BRCA_Basal"= "#BD4C26"))
###############
metadata= metadata[, -1] %>% as.data.frame()
colnames(metadata)="subtype"
rownames(metadata)= scores$sample
##
metadata$subtype= factor(metadata$subtype, levels = c("BRCA_Basal", "BRCA_LumA" ,  "BRCA_LumB"  , "BRCA_Her2","BRCA_Normal" ) )
colAnn <- HeatmapAnnotation(df = metadata,
  which = 'col',
  col = colours,
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(0.5, 'mm'))
##------------- FOR THE GENE ------------------##
 p16$fc= ifelse(p16$avg_log2FC > 0, "Up", "Down")
gene= p16[, c(1,7)]
gene=gene[which(gene$gene %in% colnames(log_centered_scaled)),]
gene=gene[match(colnames(log_centered_scaled), gene$gene),]
rownames(gene)=NULL
gene= gene %>% column_to_rownames("gene") 

stopifnot(identical(rownames(gene),colnames(log_centered_scaled) ))

###
colg= list("fc"= c("Up"="red", "Down"="blue"))
rowAnn= HeatmapAnnotation(df = gene,
  which = 'row',
  col = colg,
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(0.5, 'mm'))
```

```{r}
d=log_centered_scaled[, which(colnames(log_centered_scaled) %in% p16d$gene)]
hmap <- Heatmap(
  t(d),
  name = "expression",
 # col = bluered(75), 
  show_row_names = TRUE,
  show_column_names = FALSE,
  cluster_rows =TRUE,
  cluster_columns = TRUE,
  show_column_dend = TRUE,
  show_row_dend = TRUE,
  row_dend_reorder = FALSE,
  column_dend_reorder = TRUE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  #width = unit(100, "mm"),
  top_annotation=colAnn,row_names_gp = gpar(fontsize = 5, fontfamily = "sans"))

pdf(paste0(outputDir2, "Heatmap_Genes_Up.pdf"), height = 10, width = 10)
draw(hmap, heatmap_legend_side="left", annotation_legend_side="right")
dev.off()
```

```{r}
hmap <- Heatmap(
  t(log_centered_scaled),
  name = "expression",
 # col = bluered(75), 
  show_row_names = TRUE,
  show_column_names = FALSE,
  cluster_rows =TRUE,
  cluster_columns = TRUE,
  show_column_dend = TRUE,
  show_row_dend = TRUE,
  row_dend_reorder = FALSE,
  column_dend_reorder = TRUE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  top_annotation=colAnn ,row_names_gp = gpar(fontsize = 5, fontfamily = "sans"))

pdf(paste0(outputDir2, "Heatmap_Genes_Up_Down3.pdf"), height = 10, width = 10)
draw(hmap, heatmap_legend_side="left", annotation_legend_side="right")
dev.off()
```


#-----------Basal---------------###
Focus on Basal cells:
```{r}
bsl.metadata= metadata %>% filter(subtype=="BRCA_Basal")
basal <- log_centered_scaled[which(rownames(log_centered_scaled) %in% rownames(bsl.metadata)),]
 stopifnot(identical(rownames(bsl.metadata), rownames(basal)))
stopifnot(identical(colnames(basal), rownames(gene)))
#------------------------HEATMAP----------------##

hmap <- Heatmap(
  t(basal),
  name = "expression",
 # col = bluered(75), 
  show_row_names = TRUE,
  show_column_names = FALSE,
  cluster_rows =TRUE,
  cluster_columns = TRUE,
  show_column_dend = TRUE,
  show_row_dend = TRUE,
  row_dend_reorder = FALSE,
  column_dend_reorder = TRUE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  row_names_gp = gpar(fontsize = 5, fontfamily = "sans"),
  left_annotation = rowAnn)

pdf(paste0(outputDir2, "Heatmap_Basal_Genes_Up_Down.pdf"), height = 10, width = 10)
draw(hmap, heatmap_legend_side="left", annotation_legend_side="right")
dev.off()
```


##----------------Correlation heatmap--------------------##
```{r}
cor.basal=  WGCNA::cor(basal, use = "pairwise.complete.obs")
hmap <- Heatmap(
 cor.basal,
  name = "pairwise correlation",
 # col = bluered(75), 
  show_row_names = TRUE,
  show_column_names = TRUE,
  cluster_rows =TRUE,
  cluster_columns = TRUE,
  show_column_dend = TRUE,
  show_row_dend = TRUE,
  row_dend_reorder = FALSE,
  column_dend_reorder = FALSE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  row_names_gp = gpar(fontsize = 7, fontfamily = "sans"),column_names_gp = gpar(fontsize = 7, fontfamily = "sans"),left_annotation = rowAnn)

pdf(paste0(outputDir2, "Heatmap_Pariwise_Correlation_Basal_Genes_Up_Down.pdf"), height = 10, width = 14)
draw(hmap)
dev.off()
```

#------------ STAGE ------------------------##

Add metadata:
```{r}
tn$stage= tn$Neoplasm.Disease.Stage.American.Joint.Committee.on.Cancer.Code

tn$simplifiedStage= ifelse(tn$stage== "STAGE I"| tn$stage== "STAGE IA", "1",
                                 ifelse(tn$stage== "STAGE II"| tn$stage== "STAGE IIA" | tn$stage== "STAGE IIB", "2", 
                                  ifelse(tn$stage== "STAGE IIIA"| tn$stage== "STAGE IIIB" | tn$stage== "STAGE IIIC" , "3", "3")))

bsl.metadata= tn %>% select(simplifiedStage) %>% filter(is.na(simplifiedStage)== FALSE)
rownames(bsl.metadata)= tn$Sample.ID
bsl.metadata=as.data.frame(bsl.metadata)
stopifnot(identical(rownames(bsl.metadata), rownames(basal)))
```

```{r}
colb=list("simplifiedStage"=c("1"= "#AD1818","2"= "#BA9191","3"= "#D6CECE"))

colAnn <- HeatmapAnnotation(df = bsl.metadata,
  which = 'col',
  col = colb,
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(1, 'mm'))
###----------------------------------------------------------#####

hmap <- Heatmap(
  t(basal),
  name = "Basal samples",
  show_row_names = TRUE,
  show_column_names = FALSE,
  cluster_rows =TRUE,
  cluster_columns = FALSE,
  show_column_dend = TRUE,
  show_row_dend = TRUE,
  row_dend_reorder = FALSE,
  column_dend_reorder = TRUE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  top_annotation=colAnn,left_annotation = rowAnn,
  row_names_gp = gpar(fontsize = 5, fontfamily = "sans"))

pdf(paste0(outputDir2, "Heatmap_Basal_Genes_Up_Down2_Survival.pdf"), height = 10, width = 10)
draw(hmap, heatmap_legend_side="left", annotation_legend_side="right")
dev.off()
```

#------------------ P16 Signature construction ---------------------##
Construct the signature :
```{r}
p16$sign= ifelse(p16$avg_log2FC > 0, paste0(p16$gene, "+"), paste0(p16$gene, "-"))
p16signature= list("preTumoral"= p16$sign)

library(data.table)
p16signature.bsl= UCell::ScoreSignatures_UCell(as.data.frame(t(basal)), features=p16signature) %>% as.data.frame()

p16signature.bsl$sign= ifelse(p16signature.bsl$preTumoral_UCell > quantile(p16signature.bsl$preTumoral_UCell, 0.77), "High", "Low")

p16signature.bsl= as.data.frame(p16signature.bsl[,2], row.names= rownames(p16signature.bsl))
colnames(p16signature.bsl)="Binary_p16Sig"

```



Reorder elements according to a heatmap:
```{r}
library(dplyr)
p16signature.bsl= p16signature.bsl %>% arrange(factor(Binary_p16Sig, levels = c("High", "Low")))
basal2=basal[rownames(p16signature.bsl),]
stopifnot(identical(rownames(p16signature.bsl), rownames(basal2)))

colb=list("Binary_p16Sig"=c("High"= "#AD1818","Low"= "#D6CECE"))

colAnn <- HeatmapAnnotation(df = p16signature.bsl,
  which = 'col',
  col = colb,
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(1, 'mm'))


hmap <- Heatmap(
  t(basal2),
  name = "expression",
 # col = bluered(75), 
  show_row_names = TRUE,
  show_column_names = FALSE,
  cluster_rows =FALSE,
  cluster_columns = FALSE,
  show_column_dend = FALSE,
  show_row_dend = FALSE,
  row_dend_reorder = FALSE,
  column_dend_reorder = FALSE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  top_annotation=colAnn, column_order = rownames(p16signature.bsl))

pdf(paste0(outputDir2, "Heatmap_Basal_P16Signature_BothPosNeg_Binary_Ordered_P16sig.pdf"), height = 10, width = 10)
draw(hmap, heatmap_legend_side="left", annotation_legend_side="right")
dev.off()
```
Test only positive genes: 0.74
```{r}
p16$sign= ifelse(p16$avg_log2FC > 0, paste0(p16$gene, "+"), paste0(p16$gene, "-"))

p16signature= list("preTumoral"= p16[which(p16$avg_log2FC> 0), 1])


p16signature.bsl= UCell::ScoreSignatures_UCell(t(basal), features=p16signature) %>% as.data.frame()
p16signature.bsl$sign= ifelse(p16signature.bsl$preTumoral_UCell > quantile(p16signature.bsl$preTumoral_UCell, 0.74), "High", "Low")

p16signature.bsl= as.data.frame(p16signature.bsl[,2], row.names= rownames(p16signature.bsl))
colnames(p16signature.bsl)="Binary_p16Sig"

######--------------------------------###################

library(dplyr)
p16signature.bsl= p16signature.bsl %>% arrange(factor(Binary_p16Sig, levels = c("High", "Low")))
basal2=basal[rownames(p16signature.bsl),]
stopifnot(identical(rownames(p16signature.bsl), rownames(basal2)))

colb=list("Binary_p16Sig"=c("High"= "#AD1818","Low"= "#D6CECE"))

colAnn <- HeatmapAnnotation(df = p16signature.bsl,
  which = 'col',
  col = colb,
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(1, 'mm'))


hmap <- Heatmap(
  t(basal2),
  name = "expression",
 # col = bluered(75), 
  show_row_names = TRUE,
  show_column_names = FALSE,
  cluster_rows =TRUE,
  cluster_columns = FALSE,
  show_column_dend = FALSE,
  show_row_dend = TRUE,
  row_dend_reorder = FALSE,
  column_dend_reorder = FALSE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  top_annotation=colAnn, column_order = rownames(p16signature.bsl))

pdf(paste0(outputDir2, "Heatmap_Basal_P16Signature_Pos_Binary_Ordered_P16sig.pdf"), height = 10, width = 10)
draw(hmap, heatmap_legend_side="left", annotation_legend_side="right")
dev.off()
```




Re zscale the basal data:
```{r}
norm.matB= normalized_counts[, which(colnames(normalized_counts) %in% rownames(basal2))]
#------------------- Substract the mean per gene-------------------###
norm.matB= norm.matB[which(rownames(norm.matB) %in% p16$gene),]

log_norm.matB= log(norm.matB+1)
log_centeredB <- sweep(log_norm.matB, MARGIN=1, STATS= rowMeans(log_norm.matB))
log_centered_scaledB= apply(log_centeredB, 1, function(x)  x/(max(x))) 
#---------
pdf(paste0(outputDir2, "Plots_BasalOnly.pdf"))
hist(norm.matB)
hist(log(norm.matB))
hist(log_norm.matB)
hist(log_centeredB)
hist(log_centered_scaledB)
dev.off()
#------------
```


```{r}
p16$sign= ifelse(p16$avg_log2FC > 0, paste0(p16$gene, "+"), paste0(p16$gene, "-"))
p16signature= list("preTumoral"= p16[which(p16$avg_log2FC> 0), 1])


p16signature.bsl= UCell::ScoreSignatures_UCell(t(basal), features=p16signature) %>% as.data.frame()
p16signature.bsl$sign= ifelse(p16signature.bsl$preTumoral_UCell > quantile(p16signature.bsl$preTumoral_UCell, 0.74), "High", "Low")

p16signature.bsl= as.data.frame(p16signature.bsl[,2], row.names= rownames(p16signature.bsl))
colnames(p16signature.bsl)="Binary_p16Sig"

######--------------------------------###################

library(dplyr)
p16signature.bsl= p16signature.bsl %>% arrange(factor(Binary_p16Sig, levels = c("High", "Low")))

log_centered_scaledB=log_centered_scaledB[rownames(p16signature.bsl),]

stopifnot(identical(rownames(p16signature.bsl), rownames(log_centered_scaledB)))

colb=list("Binary_p16Sig"=c("High"= "#AD1818","Low"= "#D6CECE"))

colAnn <- HeatmapAnnotation(df = p16signature.bsl,
  which = 'col',
  col = colb,
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(1, 'mm'))


hmap <- Heatmap(
  t(log_centered_scaledB),
  name = "expression",
 # col = bluered(75), 
  show_row_names = TRUE,
  show_column_names = FALSE,
  cluster_rows =TRUE,
  cluster_columns = FALSE,
  show_column_dend = FALSE,
  show_row_dend = TRUE,
  row_dend_reorder = FALSE,
  column_dend_reorder = FALSE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  top_annotation=colAnn, column_order = rownames(p16signature.bsl))

pdf(paste0(outputDir2, "Heatmap_Basal_P16Signature_Pos_Rescaled.pdf"), height = 10, width = 10)
draw(hmap, heatmap_legend_side="left", annotation_legend_side="right")
dev.off()
```

Test a gene:
```{r}
log_centered_scaledB=as.data.frame(log_centered_scaledB)
log_centered_scaledB$bon= p16signature.bsl$Binary_p16Sig
##
log_centered_scaledB %>% ggplot(aes(x= bon, y=AQP5))+ geom_boxplot()
log_centered_scaledB %>% ggplot(aes(x= bon, y=CDKN2A))+ geom_boxplot()
log_centered_scaledB %>% ggplot(aes(x= bon, y= IGFBP4))+ geom_boxplot()
log_centered_scaledB %>% ggplot(aes(x= bon, y= MUC15))+ geom_boxplot()
log_centered_scaledB %>% ggplot(aes(x= bon, y= COL9A3))+ geom_boxplot()
log_centered_scaledB %>% ggplot(aes(x= bon, y= MIA))+ geom_boxplot()
log_centered_scaledB %>% ggplot(aes(x= bon, y= LTF))+ geom_boxplot()
```

```{r}
dv <- apply(log_centered_scaledB[,-ncol(log_centered_scaledB)], 2, var) %>% as.data.frame()
```

```{r}
res=data.frame("Gene"= colnames(basal2)[-ncol(basal2)], "Pval"= NA)
for(i in seq_len(nrow(res))){
  res[i,2] <- wilcox.test(basal2[,i] ~ bon, basal2, alternative = "two.sided")$p.value
}
res$Adj_Pval= p.adjust(res$Pval, method = "BH")
```

Plot the expression values of significantly different gene patterns:
```{r}
res %>% dplyr::filter(Adj_Pval < 0.05 ) %>% dplyr::select("Gene")
gs <- res %>% dplyr::filter(Adj_Pval < 0.05) %>% dplyr::select("Gene")
```

```{r}
colAnn <- HeatmapAnnotation(df = p16signature.bsl,
  which = 'col',
  col = colb,
  annotation_width = unit(c(1, 4), 'cm'),
  gap = unit(1, 'mm'))


hmap <- Heatmap(
  t(log_centered_scaledB[,gs$Gene]),
  name = "expression",
 # col = bluered(75), 
  show_row_names = TRUE,
  show_column_names = FALSE,
  cluster_rows =TRUE,
  cluster_columns = FALSE,
  show_column_dend = FALSE,
  show_row_dend = TRUE,
  row_dend_reorder = FALSE,
  column_dend_reorder = FALSE,
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  top_annotation=colAnn, column_order = rownames(p16signature.bsl))

pdf(paste0(outputDir2, "Heatmap_Basal_TopDifferentGenes.pdf"), height = 10, width = 10)
draw(hmap, heatmap_legend_side="left", annotation_legend_side="right")
dev.off()
```





