---
title: "Step2_DownstreamPlots"
author: "Melissa"
date: "02/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r directories creation}
#the main Dir should be the root directory in which both input, output and script folders are stored
mainDir= here::here()
#
mainDir= "C:/Users/Melissa S/Desktop/PROJECTS/BRCA1_Tumorigenesis/REVIEW"
inputDir=  file.path(mainDir,"input", "mouse",    "multistaining") 
outputDir= file.path(mainDir, "output", "mouse","multiStaining")
#####################
RDatadir <- file.path(outputDir,"RData") ; if(!file.exists(RDatadir)){dir.create(RDatadir)}
QCdir <- file.path(outputDir, "QC") ; if(!file.exists(QCdir)){dir.create(QCdir)}
PLOTdir <- file.path(outputDir, "Plots") ; if(!file.exists(PLOTdir)){dir.create(PLOTdir)}
```

```{r}
library(ggplot2)
library(ggpubr)
library(dplyr)
library(reshape2)
library(qdapRegex)
library(mclust)
library(stringr)
library(data.table)
library(sunburstR)
```

```{r}
creN= epith %>% dplyr::filter(Parent=="CreN") %>% dplyr::filter(Krt8== 1  & Krt5==0 & ) #1088


d=as.data.frame(table(creN$Vim, creN$Ncad, creN$Ecad))
colnames(d)=c("Vim", "Ncad", "Ecad", "Freq")


creN$emt= ifelse(creN$Vim==1 & creN$Ncad==1, "Mes", "Ep")
```


```{r}
juxta= epith %>% dplyr::filter(Parent=="Juxta") %>% dplyr::filter(Cdkn2a==1 & Krt8==1 & Krt5==0)  
#%>% dplyr::filter(Cdkn2a== 0 ) 
d=as.data.frame(table(juxta$Vim, juxta$Ncad, juxta$Ecad, juxta$Krt8))
colnames(d)=c("Vim", "Ncad", "Ecad", "Krt8","Freq")

```

```{r}
mmT= epith %>% dplyr::filter(Parent=="mmT") %>% dplyr::filter(Krt8== 1 & Krt5==0) 

d=as.data.frame(table(mmT$Vim, mmT$Ncad, mmT$Ecad, mmT$Krt8))
colnames(d)=c("Vim", "Ncad", "Ecad", "Krt8", "Freq")
####
mat=matrix(data=c(1976,33,13664,261 ), byrow = TRUE)
```

```{r}
mat=dtest
mat=qs::qread(paste0(RDatadir, "/Thresholded_matrix_Binarized_WoStroma.qs"))
mat$name=paste0(mat$sample,"_", mat$Parent)
#epith= mat  %>% filter(! Parent== "Stroma") 
# epith= epith %>% filter(! name=="Cre_pos_3M_Duct") 

epith= mat
epith$Epithtype= ifelse(epith$Krt5== 1 & epith$Krt8== 0, "Basal", ifelse(epith$Krt5== 0 & epith$Krt8== 1, "Luminal", ifelse(epith$Krt5== 1 & epith$Krt8== 1, "Krt5Krt8p", "none")))

epith= epith %>% select(! c("Krt5", "Krt8", "Hemato"))
epith$Cdkn2a= ifelse(epith$Cdkn2a== 0, "p16neg", "p16pos")
epith$Ki67= ifelse(epith$Ki67== 0, "Ki67neg", "Ki67pos")
epith$EMstate = ifelse(epith$Ecad== 1 & epith$Ncad== 0 , "Epithelial",
                       ifelse(epith$Ncad== 1 & epith$Ecad== 0 , "Mesenchymal",
                              ifelse(epith$Ncad== 1 & epith$Ecad== 1, "DP",
                              "undetermined" )))
 

epith$Vim= ifelse(epith$Vim== 0, "Vimneg", "Vimpos")

epith= epith %>% select(! c("Ecad", "Ncad"))
```


count the number of p16 cycling EMT pos:
```{r}
poi= epith %>% filter(Cdkn2a=="p16pos" & EMstate== "Epithelial" & Epithtype=="Luminal" & Ki67=="Ki67pos") 
n= as.data.frame(table(epith$name))
colnames(n)= c("Var1", "TotFreq")
```

```{r}
n= as.data.frame(table(epith$Parent))
colnames(n)= c("Var1", "TotFreq")
epith=mat
p16p= epith %>% filter(Cdkn2a== "1")

p16p= as.data.frame(table(p16p$Parent))
```


Show the existence of p16 positive cells:
```{r}
n= as.data.frame(table(epith$name))
colnames(n)= c("Var1", "TotFreq")

p16p= epith %>% filter(Cdkn2a== "p16pos")
p16p= as.data.frame(table(p16p$name))

p16p= merge(p16p, n, by="Var1")


p16p$freq_ratio= round(p16p$Freq/p16p$TotFreq * 5000)
library(Epi)
p16p$Var1=factor(p16p$Var1, levels = c("Cre_neg_5M_Duct","Cre_pos_3M_Duct","T7071_Duct","mlesion_7058_mlesion","T99_5617_mlesion_mlesion","T99_5617_Tumeur_Tumor","T7071_Tumor"))
writexl::write_xlsx(p16p,paste0(RDatadir, "/Table_Existance_P16PositiveCells.xlsx"))

pdf(paste0(PLOTdir, "/Existence_positive_Cdkn2acells_Reordered_H.pdf"))
g <- p16p %>% ggplot(aes(x= factor(Var1), y= freq_ratio)) + geom_bar(stat="identity", color= "black",fill= "#F08DE0")  + themplot + ylab("proportion in 5000 cells ") + ggtitle("Propo_p16 positive cells")
print(g)
dev.off()

```


```{r}
tb=matrix(a=c(66, 125), b=c(938, 3622))
```

##B: existance of p16 positive cells per Condition: Duct, Lesion, Tumor:
```{r}
epith= mat  %>% filter(! Parent== "Stroma") 
epith$Epithtype= ifelse(epith$Krt5== 1 & epith$Krt8== 0, "Basal", ifelse(epith$Krt5== 0 & epith$Krt8== 1, "Luminal", ifelse(epith$Krt5== 1 & epith$Krt8== 1, "Krt5Krt8p", "none")))
epith= epith %>% select(! c("Krt5", "Krt8", "H3K27me3"))
epith$Cdkn2a= ifelse(epith$Cdkn2a== 0, "p16neg", "p16pos")
epith$Ki67= ifelse(epith$Ki67== 0, "Ki67neg", "Ki67pos")
epith$EMstate = ifelse(epith$Ecad== 1 & epith$Ncad== 0 , "Epithelial",
                       ifelse(epith$Ncad== 1 & epith$Ecad== 0 , "Mesenchymal",
                              ifelse(epith$Ncad== 1 & epith$Ecad== 1, "DP",
                              "undetermined" )))
 

epith$Vim= ifelse(epith$Vim== 0, "Vimneg", "Vimpos")

epith= epith %>% select(! c("Ecad", "Ncad"))



n= as.data.frame(table(epith$name))
colnames(n)= c("Var1" ,"TotFreq")
p16p= epith %>% filter(Cdkn2a== "p16pos")
p16p= as.data.frame(table(p16p$name))
p16p= merge(p16p, n, by="Var1")

p16p$freq_ratio= round(p16p$Freq/p16p$TotFreq * 5000)



pdf(paste0(PLOTdir, "/Existence_positive_Cdkn2acells_Parent.pdf"))
g <- p16p %>% ggplot(aes(x= factor(Var1), y= freq_ratio)) + geom_bar(stat="identity", color= "black",fill= "white") + coord_flip() + themplot + ylab("proportion in 5000 cells ") + ggtitle("Propo_p16 positive cells_Parent_5000cells")
print(g)
dev.off()

```

Existance p16 positive cells:
```{r}
n= as.data.frame(table(epith$Parent))
colnames(n)= c("Var2", "TotFreq")
p16p= epith %>% filter(Cdkn2a== "p16pos")

p16p= as.data.frame(table(p16p$sample,p16p$Parent))
p16p= merge(p16p, n, by="Var2")
p16p= p16p %>% filter(Freq > 0)


pdf(paste0(PLOTdir, "/DotPlot_RawCounts_Existence_positive_Cdkn2acells_Parent.pdf"))
g <- p16p %>% ggplot(aes(x= factor(Var2), y= log10(Freq), fill= Var1)) + geom_dotplot(binaxis = "y", stackdir = "center" ,dotsize=0.5) + themplot + ylab("log10(Raw_Count_p16positive cells)") + ggtitle("RawCounts_p16pos")
print(g)
dev.off()

##Now divide per frequency in sample size:

n= as.data.frame(table(epith$sample))
colnames(n)= c("Var1", "TotFreq")

p16p= epith %>% filter(Cdkn2a== "p16pos")

p16p= as.data.frame(table(p16p$sample,p16p$Parent))
p16p= merge(p16p, n, by="Var1")
p16p= p16p %>% filter(Freq > 0)

p16p$freq_ratio= round(p16p$Freq/p16p$TotFreq * 5000)

pdf(paste0(PLOTdir, "/DotPlot_freq5000_sampleSize_Existence_positive_Cdkn2acells_Parent.pdf"))
g <- p16p %>% ggplot(aes(x= factor(Var2), y= freq_ratio, fill= Var1)) + geom_dotplot(binaxis = "y", stackdir = "center" ,dotsize=0.5) + themplot + ylab("RFreq5K_SampleSize_p16positive cells ") + ggtitle("FreqRatioSize_5k_p16pos")
print(g)
dev.off()

pdf(paste0(PLOTdir, "/DotPlot_freq5000_sampleSize_Existence_positive_Cdkn2acells_Parent.pdf"))
g <- p16p %>% ggplot(aes(x= factor(Var2), y= freq_ratio, fill= Var1)) + geom_bar(position = "dodge") + themplot + ylab("RFreq5K_SampleSize_p16positive cells ") + ggtitle("FreqRatioSize_5k_p16pos")
print(g)
dev.off()
```


Figure B:Existance of p16 positive cells:
```{r}
epith= mat  %>% filter(! Parent== "Stroma") %>% filter(! name=="Cre_pos_3M_Duct") 

epith=mat
epith$Epithtype= ifelse(epith$Krt5== 1 & epith$Krt8== 0, "Basal", ifelse(epith$Krt5== 0 & epith$Krt8== 1, "Luminal", ifelse(epith$Krt5== 1 & epith$Krt8== 1, "Krt5Krt8p", "none")))
epith= epith %>% select(! c("Krt5", "Krt8", "Hemato"))
epith$Cdkn2a= ifelse(epith$Cdkn2a== 0, "p16neg", "p16pos")
epith$Ki67= ifelse(epith$Ki67== 0, "Ki67neg", "Ki67pos")
epith$EMstate = ifelse(epith$Ecad== 1 & epith$Ncad== 0 , "Epithelial",
                       ifelse(epith$Ncad== 1 & epith$Ecad== 0 , "Mesenchymal",
                              ifelse(epith$Ncad== 1 & epith$Ecad== 1, "DP",
                              "undetermined" )))
 

epith$Vim= ifelse(epith$Vim== 0, "Vimneg", "Vimpos")

epith= epith %>% select(! c("Ecad", "Ncad"))
###
n= as.data.frame(table(epith$name))
colnames(n)= c("Var1", "TotFreq")

p16p= epith %>% filter(Cdkn2a=="p16pos")
p16p=as.data.frame(table(p16p$name))
p16p= merge(p16p, n, by="Var1")
p16p$percentage= p16p$Freq / p16p$TotFreq * 100
##
p16p$Var1=as.factor(p16p$Var1)

#p16p$Var1= Relevel(p16p$Var1, ref= c("Cre_neg_5M_Duct", "Cre_pos_3M_Duct","T7071_Duct","mlesion_7058_mlesion","T99_5617_mlesion_mlesion","T99_5617_Tumeur_Tumor","T7071_Tumor"))
p16p$Parent= gsub(".*_", "", p16p$Var1)
library(Epi)

p16p$Parent= Relevel(p16p$Parent, ref=c("CreN", "Juxta", "mmT", "Tumor"))


m= data.frame("Var1"= c("CreN_5M_7060_CreN"), "Freq"=1, "TotFreq"=0,"percentage"=0.1, Parent="CreN")
p16p= rbind(p16p, m)

pdf(paste0(PLOTdir, "/FigB_DotPlot_Percentage_p16positive.pdf"))
p <- p16p %>% ggplot(aes(x= factor(Parent), y=percentage, fill= Var1 ))+ geom_dotplot(binaxis = "y", stackdir = "center" ,dotsize=0.5) + themplot + ylab("Percentage of sample")
print(p)
dev.off()


p16mat=p16p
```

Compare :
Compute Fischer test :
```{r}
epith2= epith
#epith2$Parent= ifelse(epith2$name== "Cre_neg_5M_Duct", "Duct_Neg", epith2$Parent)
p16p=as.data.frame.matrix(table(epith2$Parent, epith$Cdkn2a))
```

```{r}
p16p=as.data.frame.matrix(table(epith2$Parent, epith2$Cdkn2a))
### Compute pairwise Chi2 test:
cmb=combn(x= rownames(p16p), 2) %>% as.data.frame()
#cmb=cmb[,c(1,2,6)]

for(i in seq(1, ncol(cmb))){
  res=fisher.test(p16p[cmb[1:2,i],],workspace=2e8)$p.value
  cmb[3,i] = res
}
##
cmb[4,]=p.adjust(cmb[3,], method = "bonferroni", n = ncol(cmb))
rownames(cmb)=c("subtype1", "subtype2","p_val","adjust_pval")

l= list("ContingencyTable_p16numbers"= p16p, "FisherTest"=cmb)
writexl::write_xlsx(l, paste0(PLOTdir, "/FisherTest_p16_Nbr_PosNeg.xlsx"))
gc()
```


Origin of p16 positive cells:
```{r}
epith=mat

p16p= epith %>% filter(Cdkn2a== "p16pos")
p16p= as.data.frame(table(p16p$Epithtype))  #14128

p16p$percentages= p16p$Freq/ 14128 * 100


# Compute the cumulative percentages (top of each rectangle)
p16p$ymax <- cumsum(p16p$percentages)
# Compute the bottom of each rectangle
p16p$ymin <- c(0, head(p16p$ymax, n=-1))

# Compute label position
p16p$labelPosition <- (p16p$ymax + p16p$ymin) / 2
writexl::write_xlsx(p16p,paste0(RDatadir, "/P16positive_Origin_Epithelial_PercentagesFromP16pos_New.xlsx"))

# Make the plot
pdf(paste0(PLOTdir, "/FigureC_Origin_positive_Cdkn2acells_New.pdf"))
p <- ggplot(p16p, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Var1)) +
  geom_rect( color="black") +
  scale_fill_manual(values=c("Basal"= "#6DDB8C","Krt5Krt8p"= "#D076FA","Luminal"= "#E67373","none"= "#E6DADA")) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "right") + ggtitle("p16 positive_Origin_percentagep16pos") 
print(p)
dev.off()
```


Cycle:
barplot percentages:
```{r}



n= as.data.frame(table(epith$Parent))
colnames(n)= c("Var1", "TotFreq")
###


p16lum2= epith %>% filter(Cdkn2a=="p16pos") #%>% filter(! Parent== "CreN" )
dt.p16lum=as.data.frame(table(p16lum2$Parent, p16lum2$Ki67))

##
dt.p16lum= merge(dt.p16lum, n, by="Var1")

writexl::write_xlsx(dt.p16lum, paste0(RDatadir, "/C_PositiveP16cells_Cycle_Parent_WCreN.xlsx"))
dt.p16lum= readxl::read_excel(paste0(RDatadir, "/C_PositiveP16cells_Cycle_Parent_WOCre5N.xlsx"))
########
tmp2= epith #%>% filter (Parent== "CreN")
n= as.data.frame(table( tmp2$Ki67))
colnames(n)= c("Var2", "Freq")
n$Var1="Epithelial"
n$TotFreq=n$Freq
n=n[, c("Var1", "Var2", "Freq", "TotFreq")]
###
dt.p16lum= rbind(dt.p16lum, n)

pdf(paste0(PLOTdir, "/D_Barplot_P16positiveCycling_Parent_Pooled2Epith.pdf"))
g <- dt.p16lum %>% ggplot(aes(x= factor(Var1), y=Freq, fill= Var2)) + geom_bar(stat="identity",position="fill", color="black") + ylab("% from p16 positive cells") + ggtitle("Cycle_P16positive cells") + scale_fill_manual(values=c("Ki67neg"= "#839BDE","Ki67pos"= "#3C6CF0"))  +themplot
print(g)

dev.off()
```

```{r}
epith2= epith
epith2$Parent= ifelse(epith2$name== "Cre_neg_5M_Duct", "Duct_Neg", epith2$Parent)
###########
epith2$Parent= ifelse(epith2$Parent=="Duct_Juxta", "Duct", epith2$Parent)
####
p16lum2= epith2 %>% filter(Cdkn2a=="p16pos" | Parent=="Duct_Neg")
dt.p16lum=as.data.frame(table(p16lum2$Parent, p16lum2$Ki67))
library(Epi)
dt.p16lum$Var1=  Relevel(dt.p16lum$Var1, ref=c("Duct_Neg", "Duct", "mlesion", "Tumor"))
###

pdf(paste0(PLOTdir, "/D_Barplot_P16positiveCycling_Parent_WithCreneg_New.pdf"))
g <- dt.p16lum %>% ggplot(aes(x= factor(Var1), y=Freq, fill= Var2)) + geom_bar(stat="identity",position="fill", color="black") + ylab("% from p16 positive cells") + ggtitle("Cycle_P16positive cells") + scale_fill_manual(values=c("Ki67neg"= "#839BDE","Ki67pos"= "#3C6CF0"))  +themplot
print(g)
dev.off()
```


```{r}
### Compute pairwise Chi2 test:
 #dt <-aggregate(dt4[,3],by=list(dt4$Var1), FUN= sum) %>% column_to_rownames(., "Group.1")
df <- xtabs(Freq~Var1+Var2, data = dt.p16lum)
df=as.data.frame.matrix(df)
cmb=combn(x= rownames(df), 2) %>% as.data.frame()


for(i in seq(1, ncol(cmb))){
  res=fisher.test(df[cmb[1:2,i],],workspace=2e8)$p.value
  cmb[3,i] = res
}
##
cmb[4,]=p.adjust(cmb[3,], method = "bonferroni", n = ncol(cmb))
rownames(cmb)=c("subtype1", "subtype2","p_val","adjust_pval")

l= list("ContingencyTable_p16Positive_Cycle"= df, "FisherTest"=cmb)

writexl::write_xlsx(l, paste0(PLOTdir, "/FisherTest_p16_Pos_Cycle_New.xlsx"))
gc()
```

V2:
```{r}
### Compute pairwise Chi2 test:
 #dt <-aggregate(dt4[,3],by=list(dt4$Var1), FUN= sum) %>% column_to_rownames(., "Group.1")
df <- xtabs(Freq~Var1+Var2, data = dt.p16lum)
df=as.data.frame.matrix(df)
cmb=combn(x= rownames(df), 2) %>% as.data.frame()


for(i in seq(1, ncol(cmb))){
  res=fisher.test(df[cmb[1:2,i],],workspace=2e8)$p.value
  cmb[3,i] = res
}
##
cmb[4,]=p.adjust(cmb[3,], method = "bonferroni", n = ncol(cmb))
rownames(cmb)=c("subtype1", "subtype2","p_val","adjust_pval")

l= list("ContingencyTable_p16Positive_Cycle_Parent"= df, "FisherTest"=cmb)

writexl::write_xlsx(l, paste0(PLOTdir, "/FisherTest_p16_Pos_Cycle_Parent2_New.xlsx"))
gc()
```

EMT:
Density plots:
```{r}
dt$Juxta$Parent="Duct"
dt$Juxta_G7058$Parent="Duct"
dt$T99_Juxta$Parent="Duct"
dt$Cre_neg_5M$Parent="Duct_Neg"
dt$Cre_pos_3M$Parent= ifelse(dt$Cre_pos_3M$Parent=="PathAnnotationObject", "Stroma", dt$Cre_pos_3M$Parent)
unique(dt$mlesion_7058$Parent)
unique(dt$T7071$Parent)
unique(dt$T99_5617_mlesion$Parent)
unique(dt$T99_5617_Tumeur$Parent)

dt$T99_5617_Tumeur$Parent= ifelse(dt$T99_5617_Tumeur$Parent=="PathAnnotationObject", "Tumor", dt$T99_5617_Tumeur$Parent)

data=dt

tmp1= do.call(rbind,data)
tmp1$name=paste0(tmp1$sample, tmp1$Parent)

identical(rownames(mat), rownames(tmp1))
tmp1$Parent= mat$Parent
tmp1$Parent= ifelse(tmp1$Parent=="Duct_Juxta" , "Duct", tmp1$Parent)

tmp1= tmp1 %>% filter(! Parent=="Stroma")
 
pdf(paste0(PLOTdir, "/Density_Plots_Vim_Krt8_Filled_Test4.pdf"))
for(i in unique(tmp1$Parent)){
  m= tmp1 %>% filter(Parent== i)
  
 palette <- hcl.colors(50, palette = "viridis")
 smoothScatter(m$Channel.9.mean, m$Channel.7.mean, nbin=300, nrpoints = 10000,bandwidth = 25,colramp = colorRampPalette(palette))+ title(i)
 
  smoothScatter(m$Channel.9.mean, m$Channel.7.mean, nbin=300, nrpoints = 0,bandwidth = 25,colramp = colorRampPalette(palette))+ title(i)

  p <-  m %>% ggplot(aes(x=Channel.9.mean, y=Channel.7.mean, color=sample)) + geom_point()+ ggtitle(i) + themplot +
  geom_text(x = 200, y = 200, label = paste0("r²=", round(summary(lm(m$Channel.7.mean ~m$Channel.9.mean))$r.squared,3)))

 print(p)

}
dev.off()
#####################################
palette <- hcl.colors(50, palette = "viridis")
for(i in unique(tmp1$Parent)){
  m= tmp1 %>% filter(Parent== i)
  png(paste0(PLOTdir, "/Supp_Density_Plots_Vim_Krt8_W_Dots_",i  ,".png"))
    smoothScatter(m$Channel.9.mean, m$Channel.7.mean, nbin=300, nrpoints = 1000,bandwidth = 25,colramp =colorRampPalette(palette))+ title(i)
  dev.off()
}
```
VlnPlot:
```{r}
tmp1= tmp1 %>% filter(! Parent== "Duct_Neg")
pdf(paste0(PLOTdir, "/ViolinPlot_Krt8_Condition.pdf"))
p <-  ggplot(data=tmp1,aes(y=Channel.7.mean, x= Parent, fill=Parent)) + geom_violin() + stat_summary(fun.y=median, geom="point", size=2, color="black") +  geom_boxplot(width=0.1)+themplot + scale_fill_grey(start = 0.9, end = 0.7) + stat_compare_means(ref.group = "Duct")
print(p)
dev.off()
```


p16positive EMT:
```{r}
p16p = epith %>% filter(Cdkn2a=="p16pos") %>% filter(! name== "Cre_neg_5M_Duct" )
n= as.data.frame(table(p16p$name))
colnames(n)=c("Var1", "TotFreq")

p16emt=as.data.frame(table(p16p$name, p16p$EMstate))
p16emt=merge(p16emt, n, by="Var1")

writexl::write_xlsx(p16emt , paste0(PLOTdir, "/Number_P16_PositiveCells_EMTstate_PerSample.xlsx"))
```


```{r}
tmp2= epith %>% filter(! name=="Cre_neg_5M_Duct")
n= as.data.frame(table( tmp2$EMstate))
colnames(n)= c("Var2", "Freq")
n$Var1="Epithelial_All"
#n$TotFreq=n$Freq
n=n[, c("Var1", "Var2", "Freq")]
###
p16p = epith %>% filter(Cdkn2a=="p16pos") %>% filter(! name== "Cre_neg_5M_Duct" )
p16p$Parent= ifelse(p16p$Parent=="Duct_Juxta","Duct", p16p$Parent)
p16emt=as.data.frame(table(p16p$Parent, p16p$EMstate))


p16emt= rbind(p16emt, n)

pdf(paste0(PLOTdir, "/FigE_Luminal_NcadEcad_pos_V2.pdf"))
g <- p16emt %>% ggplot(aes(x= Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position = "fill", color="black") + scale_fill_manual(values= c("Mesenchymal"="#D2F1F7","Epithelial"= "#92999E","undetermined"= "#E6E6E6", "DP"="#76E0F0")) # + themplot
print(g)
##
g <- p16emt %>% ggplot(aes(x= Var1, y=Freq, fill=Var2)) + geom_bar(stat="identity", position = "fill", color="black") + themplot + scale_fill_manual(values= c("Mesenchymal"="#C9F3F7","Epithelial"= "#92999E","undetermined"= "#E6E6E6", "DP"="#76E0F0")) + coord_flip()
print(g)
dev.off()
```

p16emt= p16emt %>% filter(! Var1== "Cre_neg_5M_Duct" )
p16emt$Parent= p16emt$Var1
p16emt$Parent= dplyr::recode(p16emt$Var1, "Cre_pos_3M_Duct"="Duct",  "Juxta_Duct_Juxta"="Duct","Juxta_G7058_Duct_Juxta"="Duct",   "mlesion_7058_mlesion"="mLesion", "T7071_Duct"="Duct",    "T7071_Tumor"="Tumor", "T99_5617_mlesion_mlesion"="mLesion", "T99_5617_Tumeur_Tumor"="Tumor","T99_Juxta_Duct_Juxta"="Duct")

```{r}
df <- xtabs(Freq~Var1+Var2, data = p16emt)
df=as.data.frame.matrix(df)
cmb=combn(x= rownames(df), 2) %>% as.data.frame()
####
for(i in seq(1, ncol(cmb))){
  res=fisher.test(df[cmb[1:2,i],],workspace=2e8)$p.value
  cmb[3,i] = res
}
##
cmb[4,]=p.adjust(cmb[3,], method = "bonferroni", n = ncol(cmb))
rownames(cmb)=c("subtype1", "subtype2","p_val","adjust_pval")

l= list("ContingencyTable_p16Positive_EMT"= df, "FisherTest"=cmb)

writexl::write_xlsx(l, paste0(PLOTdir, "/FisherTest_p16_Pos_EMT.xlsx"))
```




