---
title: "CPTAC_BREAST"
author: "Melissa"
date: "20/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
mainDir="C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Tumor_BRCA_p53"
inputDir= file.path(mainDir,"input" ,"scRNAseq","mouse","inputCPTAC_Breast")
outputDir= file.path(mainDir, "output","scRNAseq", "CPTAC")
###
plotDir= file.path(outputDir,"Plots") ; if(!file.exists(plotDir)){dir.create(plotDir)}
RdataDir= file.path(outputDir,"Rdata") ; if(!file.exists(RdataDir)){dir.create(RdataDir)}
testDir= file.path(outputDir,"Test") ; if(!file.exists(testDir)){dir.create(testDir)}
testDir2= file.path(outputDir,"Test2") ; if(!file.exists(testDir2)){dir.create(testDir2)}
gsvaDir= file.path(outputDir,"GSVA") ; if(!file.exists(gsvaDir)){dir.create(gsvaDir)}
sigDir= file.path(outputDir,"P16Signature") ; if(!file.exists(sigDir)){dir.create(sigDir)}
```

```{r}
library(dplyr)
library(Seurat)
library(readr)
library(ggplot2)
library(UCell)
library(survival)
library(survminer)
```

Read the files:
```{r}
genexp <- read_delim("C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Tumor_BRCA_p53/input/scRNAseq/mouse/inputCPTAC_breast/Human__TCGA_BRCA__UNC__RNAseq__HiSeq_RNA__01_28_2016__BI__Gene__Firehose_RSEM_log2.cct.gz",  delim = "\t", escape_double = FALSE,    trim_ws = TRUE) %>% as.data.frame()

cptac_clin <- read_delim("C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Tumor_BRCA_p53/input/scRNAseq/mouse/inputCPTAC_breast/Human__TCGA_BRCA__MS__Clinical__Clinical__01_28_2016__BI__Clinical__Firehose.tsi",      delim = "\t", escape_double = FALSE,      trim_ws = TRUE) %>% as.data.frame()
length(intersect(colnames(genexp), colnames(cptac_clin)))

rownames(genexp)= make.unique(genexp$attrib_name)
#####
themplot=theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),axis.ticks.x=element_blank())
```

Check if data wwas normalized:
```{r}
require(reshape2)
violinMatrix <- reshape2::melt(t(genexp[, sample(colnames(genexp), 50)]))

colnames(violinMatrix) <- c("Sample","Gene","Expression")

library(ggplot2)
pdf(paste0(plotDir, "/BoxPlot_Samples_Genes2.pdf"), height = 6, width = 20)
p <- ggplot(violinMatrix, aes(x=Sample, y=Expression)) + geom_boxplot() + theme(axis.text.x = element_text(angle=45, hjust=1))
print(p)
dev.off()
```

```{r}
ge= exp(genexp[,-1])
dtest= log(ge)
####
violinMatrix <- reshape2::melt(t(ge[, sample(colnames(ge), 50)]))

colnames(violinMatrix) <- c("Sample","Gene","Expression")

library(ggplot2)
pdf(paste0(plotDir, "/BoxPlot_Samples_Genes3.pdf"), height = 6, width = 20)
p <- ggplot(violinMatrix, aes(x=Sample, y=log10(Expression))) + geom_boxplot() + theme(axis.text.x = element_text(angle=45, hjust=1))
print(p)
dev.off()
rm(dtest)
```

GE is the normalized not logged dataset:
```{r}
cptac_clin= cptac_clin %>% tibble::column_to_rownames("attrib_name")
cptac_clin= cptac_clin[,which(colnames(cptac_clin) %in% colnames(ge))]
cptac_clin=cptac_clin %>% t() %>% as.data.frame()
```

Heatmap of genes :
```{r}
gh <- ge[which(rownames(ge) %in% toupper(sigP$gene)),]
mdc <- cptac_clin[,9] %>% as.data.frame()
rownames(mdc)= rownames(cptac_clin)
colnames(mdc)="PAM50"
mdc= mdc %>% filter(! is.na(mdc$PAM50)==TRUE)
gh= gh[, which(colnames(gh) %in% rownames(mdc))]
###
p <- pheatmap::pheatmap(log(gh), scale = "column", annotation_col  = mdc, show_colnames = FALSE, cluster_cols = FALSE,cluster_rows = FALSE, )
```


calculate p16 signature:
```{r}
seu= qs::qread("C:/Users/Melissa S/Desktop/BRCA1_Tumorigenesis/Tumor_BRCA_p53/output/scRNAseq/invivo/RData/Epithelial_FinalAnnotation.qs")
library(Seurat)
DefaultAssay(seu)="RNA"
sig= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = TRUE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05)
```


TEST:

```{r}
bsl= cptac_clin %>% filter(PAM50== "Basal")

bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
bsl$status= as.numeric(as.character(bsl$status))
##
cptac.dt= ge[, which(colnames(ge) %in% rownames(bsl))]



p16list= list(p16s_5g= c(varimp_data$feature[1:5]), 
              p16s_6g= c(varimp_data$feature[1:6]), 
              p16s_7g= c(varimp_data$feature[1:7]), 
              p16s_8g= c(varimp_data$feature[1:8]), 
              p16s_9g= c(varimp_data$feature[1:9]), 
              p16s_10g= c(varimp_data$feature[1:10]), 
              p16s_11g= varimp_data$feature[1:11], 
              p16s_12g=varimp_data$feature[1:12],
              p16s_13g=varimp_data$feature[1:13],
              p16s_14g=varimp_data$feature[1:14],
            p16s_15g=varimp_data$feature[1:15],
            p16s_20g=varimp_data$feature[1:20],
            p16s_25g=varimp_data$feature[1:25],
              p16s_30g= varimp_data$feature[1:30],
              p16s_35g=varimp_data$feature[1:35],
             p16s_38g=varimp_data$feature[1:38])
library(data.table)
library(UCell)
cptac.dt=as.data.frame(cptac.dt)
sig.resCPTAC= UCell::ScoreSignatures_UCell(matrix = cptac.dt, features= p16list, maxRank = 15000)
sig.resCPTAC=as.data.frame(sig.resCPTAC)
identical(rownames(sig.resCPTAC), rownames(bsl))
sig.resCPTAC=sig.resCPTAC[rownames(bsl),]
sig.resCPTAC$stage= bsl$pathologic_stage
##---------------
sig.resCPTAC= sig.resCPTAC %>% filter(! is.na(sig.resCPTAC$stage))

sig.resCPTAC$stage= ifelse(sig.resCPTAC$stage== "stagei", "early", "late")
qs::qsave(sig.resCPTAC, paste0(outputDir2, "CPTAC_SignaturesAllgenes.qs"))
```

Work on the signatures:
```{r}
resTCGA= data.frame("Ngene"= colnames(sig.res)[-ncol(sig.res)], "pvalue"=NA)
sig.res$stage= factor(sig.res$stage, levels = c("early", "late"))

for(i in seq_len(ncol(sig.res)-1)){
  resTCGA[i,2]= wilcox.test(sig.res[,i] ~ sig.res$stage, alternative= "greater")$p.value
}
#---------------------------##
resCPTAC= data.frame("Ngene"= colnames(sig.resCPTAC)[-ncol(sig.resCPTAC)], "pvalue"=NA)

sig.resCPTAC$stage= factor(sig.resCPTAC$stage, levels = c("early", "late"))

for(i in seq_len(ncol(sig.resCPTAC)-1)){
  resCPTAC[i,2]= wilcox.test(sig.resCPTAC[,i] ~ sig.resCPTAC$stage, alternative= "greater")$p.value
}

resCPTAC$nbr= c(seq(5,15), 20, 25, 30, 35, 38)
resTCGA$nbr= c(seq(5,15), 20, 25, 30, 35, 38)

p <- resCPTAC  %>% ggplot(aes(x= nbr, y= pvalue)) + geom_line() + themplot + geom_hline(yintercept = 0.05, color= "red") + geom_vline(xintercept = 13, color="blue") + ggtitle("CPTAC_Pvalue_WilcoxStage")

g <- resTCGA  %>% ggplot(aes(x= nbr, y= pvalue)) + geom_line() + themplot + geom_hline(yintercept = 0.05, color= "red") + geom_vline(xintercept = 13, color="blue") + ggtitle("TCGA_Pvalue_WilcoxStage")

pdf(paste0(outputDir2, "Pvalue_Wilcox_NbrGenes_P16Signature.pdf"))
print(p)
print(g)
dev.off()
```







```{r}
gene.sets <- list(p16UP_signature = toupper(sigP$gene), maxRank = 15000)

scp= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
scp$sample= rownames(scp)
###

scp=scp[rownames(cptac_clin),]
stopifnot(identical(scp$sample, rownames(cptac_clin)))

cptac_clin=cbind(cptac_clin, scp)

stopifnot(identical(rownames(cptac_clin),colnames(ge)))

ge= ge[, rownames(cptac_clin)]
tge=as.data.frame(t(ge))

cptac_clin$CDKN2A= tge$CDKN2A
```


```{r}
cptac_clin=cptac_clin[, -c(25:30)]
cptac_clin=cptac_clin %>% filter(is.na(PAM50)==FALSE)
pdf(paste0(plotDir, "/P16_p16Sig_P16sigWOp16_CPTAC.pdf"))

p <-  cptac_clin %>% ggplot(aes(x=PAM50, y= p16UP_signature_UCell)) + geom_boxplot(fill='#DBD8D8', color="black")  + ylab("P16Signature")  +geom_jitter(data=cptac_clin,aes(x=PAM50, y= p16UP_signature_UCell,color= PAM50), size=1, alpha=0.9)   + scale_color_grey() +   gtheme + stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Basal",aes(label=..p.adj..)) + themplot + theme(legend.position = "none")

##################
g <-  cptac_clin %>% ggplot(aes(x=PAM50, y= log10(CDKN2A))) + geom_boxplot(fill='#DBD8D8', color="black")  + ylab("log10(CDKN2A)")  +geom_jitter(data=cptac_clin,aes(x=PAM50, y= log10(CDKN2A),color= PAM50), size=1, alpha=0.9)   + scale_color_grey() +   gtheme + stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Basal",aes(label=..p.adj..)) + themplot + theme(legend.position = "none")
#######
h <-   cptac_clin %>% ggplot(aes(x=PAM50, y= p16Sig_WO_UCell)) + geom_boxplot(fill='#DBD8D8', color="black")  + ylab("p16signature_WOP16")  +geom_jitter(data= cptac_clin,aes(x=PAM50, y= p16Sig_WO_UCell,color= PAM50), size=1, alpha=0.9)   + scale_color_grey() +   gtheme + stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Basal",aes(label=..p.adj..)) + themplot + theme(legend.position = "none")
print(p)
print(g)
print(h)
dev.off()
```



SURVIVAL curve:
```{r}
bsl= cptac_clin %>% filter(PAM50== "Basal")

bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
bsl$status= as.numeric(as.character(bsl$status))

qtl=list("quantile"=NA, "min_Pvalue"=NA)
for(i in seq(0.1, 0.9, 0.01)){

    bsl$p16discrete=  ifelse(bsl$p16UP_signature_UCell > quantile(bsl$p16UP_signature_UCell, i), 1, 0 )
  
    
  res <- pairwise_survdiff(Surv(overall_survival,status) ~ p16discrete, data = bsl)
  qtl$quantile= append(qtl$quantile, i)
  qtl$min_Pvalue= append(qtl$min_Pvalue, min(na.omit(res$p.value)))
  
}
dt_b=do.call( rbind, qtl) %>% t() %>%  as.data.frame()
dt_b=dt_b[-1,]
####

l= list("WildTest_Method"= dt.qtl_binary, "MinPval_PariwiseComp"= dt_qtl_b)
writexl::write_xlsx(l, paste0(RdataDir, "/SurvPlots_Methods_METABRIC.xlsx"))
```


```{r}
bsl$p16discrete= ifelse(bsl$p16UP_signature_UCell > quantile(bsl$p16UP_signature_UCell, 0.73), "high", "low" )

#P16Low should be tn$p16Up < quantile(tn$p16Up, 0.25)
library(survival)

fit <- survfit(Surv(overall_survival,status) ~ p16discrete, data = bsl)
pdf(paste0(plotDir, "/Survival_BestParam_METABRIC_Stratif_Binary.pdf"))
 p=ggsurvplot(fit,
          pval = TRUE, conf.int = FALSE,
          risk.table = FALSE, 
          risk.table.col = "strata",
          linetype = "strata", 
          surv.median.line = "hv", 
          ylab="OS_0.75",
          ggtheme = theme_bw(), 
          palette = c("#DED5D5", "#262424", "#756F6F")) 
 print(p)
 dev.off()
```

STAGE:
```{r}
bsl= cptac_clin %>% filter(PAM50== "Basal")
bsl$stage= gsub("stage", "", bsl$pathologic_stage)
bsl$stage= dplyr::recode(bsl$stage, "i"="1", "ii"="2", "iii"="3", "iv"="4")
bsl= bsl %>% filter(stage %in% c("1", "2", "3"))

#
pdf(paste0(plotDir, "/Stage_P16Signature.pdf"))
p <- bsl %>% ggplot(aes(x= stage, y= p16UP_signature_UCell)) + geom_boxplot() + geom_point() + stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "1",aes(label=..p.adj..))  + theme(legend.position = "none") + themplot
print(p)
dev.off()
```

#################### START HERE #################################################
Test the signature:
Method 1: Order by Fold change and select the top each time:
```{r}
sig= sig[order(sig$avg_log2FC, decreasing = TRUE),]
#clin= cptac_clin[,-c(21,22)]


tmp= data.frame("topGenes"= c(1,seq(5, nrow(sigP), 5), nrow(sigP)), "Best_Qtl"= NA, "Pval_Min"=NA)
rownames(tmp)= tmp$topGenes


for(j in c(1,seq(5, nrow(sigP), 5), nrow(sigP))){
  sigenes= sigP %>% top_n(n = j, wt = avg_log2FC) 
  gene.sets= list("p16"= toupper(sigenes$gene))
  dt= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
  dt$sample= rownames(dt)
  dt=dt[rownames(clin),]
  stopifnot(identical(dt$sample, rownames(clin)))
  md=cbind(clin, dt)

  #############################
  
  bsl= md %>% filter(PAM50== "Basal")
   bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
bsl$status= as.numeric(as.character(bsl$status))


#######################################################################################
qtl=list("quantile"=NA, "hazard"=NA, "waldtest"=NA,"logTestPval"=NA,"waldtestPval"=NA)

for(i in seq(0.1, 0.9, 0.01)){
  bsl$p16discrete= ifelse(bsl$p16sig > quantile(bsl$p16sig, i), 1, 0)
   res.cox <- summary(coxph(Surv(overall_survival,status ) ~ p16discrete, data = bsl))
   
  qtl$quantile= append(qtl$quantile, i)
  qtl$hazard=append(qtl$hazard,res.cox$coefficients[,2])
  ##
 qtl$logTestPval= append(qtl$logTestPval, res.cox$logtest[3])
  qtl$waldtest= append(qtl$waldtest,res.cox$waldtest[1] )
   qtl$waldtestPval= append(qtl$waldtestPval,res.cox$waldtest[3] )
} 
 dt.qtl_binary= do.call(rbind, qtl) %>% t() %>%  as.data.frame()
dt.qtl_binary=dt.qtl_binary[-1,]


bestqtl=  dt.qtl_binary[order( dt.qtl_binary$logTestPval, decreasing = FALSE),1][1]
bestpval=  dt.qtl_binary[order( dt.qtl_binary$logTestPval, decreasing = FALSE),4][1]

tmp[j,2]= bestqtl
tmp[j,3]=bestpval
####
bsl$p16discrete= ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, besthr), "high", "low" )

#P16Low should be tn$p16Up < quantile(tn$p16Up, 0.25)
library(survival)

fit <- survfit(Surv(overall_survival,status) ~ p16discrete, data = bsl)
pdf(paste0(sigDir, "/Survival_BestParam_METABRIC_Stratif", "TopG_",j,"_BestQtl_", besthr,".pdf"))
 p=ggsurvplot(fit,
          pval = TRUE, conf.int = FALSE,
          risk.table = FALSE, 
          risk.table.col = "strata",
          linetype = "strata", 
          surv.median.line = "hv", 
          ylab=paste0("OS_", besthr, "Top_n",j),
          ggtheme = theme_bw(), 
          palette = c("#DED5D5", "#262424", "#756F6F")) 
 print(p)
 dev.off()
}

tmp=tmp[ c(5, 10, 15, 20,25,30,35,39),]
tmp$topGenes= c(seq(5, nrow(sig), 5), nrow(sig))
 writexl::write_xlsx(tmp, paste0(plotDir, "/Table_TopGenes_BestParams.xlsx"))
 ###
p <- tmp %>% ggplot( aes(x= as.factor(topGenes), y= Best_Qtl)) + geom_bar(stat="identity") + themplot

g <- tmp %>% ggplot( aes(x= as.factor(topGenes), y= Pval_Min)) + geom_bar(stat="identity") + themplot

pdf(paste0(plotDir, "/Barplot_BestParams_TopGenes.pdf"))
print(p)
print(g)
dev.off()
```


  ## START PLOTS:
  p <-  md %>% filter(PAM50 %in% c("Basal", "Her2", "LumA", "LumB")) %>%
    ggplot(aes(x=PAM50, y= p16_UCell)) + geom_boxplot(fill='#DBD8D8', color="black")  + ylab("P16Signature")  +geom_jitter(data=md,aes(x=PAM50, y= p16_UCell,color= PAM50), size=1, alpha=0.9)   + scale_color_grey() + stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Basal",aes(label=..p.adj..)) + themplot + theme(legend.position = "none") + ggtitle(paste0("p16sig_Top_",j, "genes" ))
  pdf(paste0(plotDir, "/P16Signature_CPTAC_Top",j, "genes.pdf" ))
  print(p)
  dev.off()

USE MIN PVAL METHOD: WITH ORDERING GENES ACCORDING TO FOLD CHANGE:
```{r}

sig= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = TRUE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05)

sig= sig[order(sig$avg_log2FC, decreasing = TRUE),]
#clin= cptac_clin[,-c(21,22)]
clin=cptac_clin
tmpT= data.frame("topGenes"= c(1,seq(5, nrow(sig), 5), nrow(sig)), "Best_Qtl"= NA, "Pval_Min"=NA)
rownames(tmpT)= tmpT$topGenes

for(j in c(1,seq(5, nrow(sig), 5), nrow(sig))){
  sigenes= sig %>% top_n(n = j, wt = avg_log2FC) 
  gene.sets= list("p16"= toupper(sigenes$gene))
  dt= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
  dt$sample= rownames(dt)
  dt=dt[rownames(clin),]
  stopifnot(identical(dt$sample, rownames(clin)))
  md=cbind(clin, dt)


  #############################
  
  bsl= md %>% filter(PAM50== "Basal")
   bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
bsl$status= as.numeric(as.character(bsl$status))

###################################################################
 qtl2=list("quantile"=NA, "min_Pvalue"=NA)
for(i in seq(0.1, 0.9, 0.01)){
 print(i)
    bsl$p16discrete=  ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, i), 1, 0 )
    print("bsl")
  res <- pairwise_survdiff(Surv(overall_survival,status) ~ p16discrete, data = bsl)
  print("res")
  qtl2$quantile= append(qtl2$quantile, i)
  qtl2$min_Pvalue= append(qtl2$min_Pvalue, min(na.omit(res$p.value)))
}
dt_b=do.call( rbind, qtl2) %>% t() %>%  as.data.frame()
dt_b=dt_b[-1,]

#######################################################################################
#l= list("WildTest_Method"= dt.qtl_binary, "MinPval_PariwiseComp"= dt_b)
#writexl::write_xlsx(l, paste0(RdataDir, "/SurvPlots_Methods_METABRIC_TOP_",j ,"_genes.xlsx"))
##
bestqtl= dt_b[order(dt_b$min_Pvalue, decreasing = FALSE),1][1]
bestpval= dt_b[order(dt_b$min_Pvalue, decreasing = FALSE),2][1]

tmpT[j,2]= bestqtl
tmpT[j,3]=bestpval
####
bsl$p16discrete= ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, bestqtl), "high", "low" )

#P16Low should be tn$p16Up < quantile(tn$p16Up, 0.25)
library(survival)

fit <- survfit(Surv(overall_survival,status) ~ p16discrete, data = bsl)
pdf(paste0(testDir, "/Survival_BestParam_METABRIC_Stratif", "TopG_",j,"_BestQtl_", bestqtl,".pdf"))
 p=ggsurvplot(fit,
          pval = TRUE, conf.int = FALSE,
          risk.table = FALSE, 
          risk.table.col = "strata",
          linetype = "strata", 
          surv.median.line = "hv", 
          ylab=paste0("OS_", bestqtl, "Top_n",j),
          ggtheme = theme_bw(), 
          palette = c("#DED5D5", "#262424", "#756F6F")) 
 print(p)
 dev.off()
}

tmpT=tmpT[ c(1,5, 10, 15, 20,25,30,35,39),]
tmpT$topGenes= c(1,seq(5, nrow(sig), 5), nrow(sig))
 writexl::write_xlsx(tmpT, paste0(sigDir, "/Table_TopGenes_BestParams.xlsx"))
 ###
 pdf(paste0(sigDir, "/Barplot_BestParams_TOP.pdf"))
pT <- tmpT %>% ggplot( aes(x= as.factor(topGenes), y= Best_Qtl)) + geom_bar(stat="identity") + themplot

gT <- tmpT %>% ggplot( aes(x= as.factor(topGenes), y= Pval_Min)) + geom_bar(stat="identity") + themplot
print(pT)
print(gT)
dev.off()
```

## METHOD 2: RANDOMLY PICK:

USE MIN PVAL METHOD: WITH ORDERING GENES ACCORDING TO FOLD CHANGE:
```{r}
sig= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = TRUE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05)

tmp= data.frame("topGenes"= c(1,seq(5, nrow(sig), 5), nrow(sig)), "Best_Qtl"= NA, "Pval_Min"=NA)
rownames(tmp)= tmp$topGenes

for(j in c(1,seq(5, nrow(sig), 5), nrow(sig))){
  
  sigenes= sig %>% filter(gene %in% sample(sig$gene, j, replace = FALSE))   
  gene.sets= list("p16"= toupper(sigenes$gene))
  dt= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
  dt$sample= rownames(dt)
  dt=dt[rownames(clin),]
  stopifnot(identical(dt$sample, rownames(clin)))
  md=cbind(clin, dt)


  
  bsl= md %>% filter(PAM50== "Basal")
   bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
bsl$status= as.numeric(as.character(bsl$status))

###################################################################
 qtl2=list("quantile"=NA, "min_Pvalue"=NA)
for(i in seq(0.1, 0.9, 0.01)){
 
    bsl$p16discrete=  ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, i), 1, 0 )
  res <- pairwise_survdiff(Surv(overall_survival,status) ~ p16discrete, data = bsl)
  qtl2$quantile= append(qtl2$quantile, i)
  qtl2$min_Pvalue= append(qtl2$min_Pvalue, min(na.omit(res$p.value)))
}
dt_b=do.call( rbind, qtl2) %>% t() %>%  as.data.frame()
dt_b=dt_b[-1,]


bestqtl= dt_b[order(dt_b$min_Pvalue, decreasing = FALSE),1][1]
bestpval= dt_b[order(dt_b$min_Pvalue, decreasing = FALSE),2][1]

tmp[j,2]= bestqtl
tmp[j,3]=bestpval
####
bsl$p16discrete= ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, bestqtl), "high", "low" )


#P16Low should be tn$p16Up < quantile(tn$p16Up, 0.25)
library(survival)

fit <- survfit(Surv(overall_survival,status) ~ p16discrete, data = bsl)
pdf(paste0(sigDir, "/SurvivalPlot_CPTAC_", "RANDOM_G_",j,"_BestQtl_", bestqtl,".pdf"))
 p=ggsurvplot(fit,
          pval = TRUE, conf.int = FALSE,
          risk.table = FALSE, 
          risk.table.col = "strata",
          linetype = "strata", 
          surv.median.line = "hv", 
          ylab=paste0("OS_", bestqtl, "Top_n",j),
          ggtheme = theme_bw(), 
          palette = c("#DED5D5", "#262424", "#756F6F")) 
 print(p)
 dev.off()
}

tmp=tmp[ c(1,5, 10, 15, 20,25,30,35,39),]
tmp$topGenes= c(1,seq(5, nrow(sig), 5), nrow(sig))
 writexl::write_xlsx(tmp, paste0(testDir, "/Table_RandomGenes_BestParams.xlsx"))
 ###
 pdf(paste0(plotDir, "/Barplot_BestParams_Random.pdf"))
p <- tmp %>% ggplot( aes(x= as.factor(topGenes), y= Best_Qtl)) + geom_bar(stat="identity") + themplot

g <- tmp %>% ggplot( aes(x= as.factor(topGenes), y= Pval_Min)) + geom_bar(stat="identity") + themplot
print(p)
print(g)
dev.off()
```




GROUP ALL TOGETHER:
```{r}
tmpT$type= "TopGenes"
tmp$type="Random"
dt= rbind(tmpRM,tmp, tmpT)
writexl::write_xlsx(dt, paste0(testDir, "/DATASET_RandomSel_TopG_WOTopG.xlsx"))

pdf(paste0(testDir, "/Lines_Random_TopGenes_MinPval_BestQtl_Added_WO_TopGenes.pdf"))

m <- dt %>% ggplot(aes(x= topGenes, y= Pval_Min, color= type)) + geom_point() + geom_line() + themplot

n <- dt %>% ggplot(aes(x= topGenes, y= Best_Qtl, color= type)) + geom_point() + geom_line() + themplot

print(m)
print(n)
dev.off()
```

MAKE A FUNCTION:
```{r}

sig= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = TRUE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05)

test_function_CPTAC= function(analysis_mode, sig){
  
   qtl2=list("quantile"=NA, "min_Pvalue"=NA)
   tmp= data.frame("topGenes"= c(1,seq(5, nrow(sig), 5), nrow(sig)), "Best_Qtl"= NA, "Pval_Min"=NA)
   rownames(tmp)= tmp$topGenes
  
for(j in c(1,seq(5, nrow(sig), 5), nrow(sig))){

      if(analysis_mode== "TopGenes"){
        tmp$type="TopGenes"
        sig= sig[order(sig$avg_log2FC, decreasing = TRUE),]
        sigenes= sig %>% top_n(n = j, wt = avg_log2FC) 
        gene.sets= list("p16"= toupper(sigenes$gene))
          dt= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
       dt$sample= rownames(dt)
        dt=dt[rownames(clin),]
       stopifnot(identical(dt$sample, rownames(clin)))
       md=cbind(clin, dt)
       bsl= md %>% filter(PAM50== "Basal")
      bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
      bsl$status= as.numeric(as.character(bsl$status))
  }
  
        if(analysis_mode== "RandomGenes"){
            tmp$type="RandomGenes"
           sigenes= sig %>% filter(gene %in% sample(sig$gene, j, replace = FALSE))   
           gene.sets= list("p16"= toupper(sigenes$gene))
           
             dt= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
           dt$sample= rownames(dt)
          dt=dt[rownames(clin),]
          stopifnot(identical(dt$sample, rownames(clin)))
          md=cbind(clin, dt)
          bsl= md %>% filter(PAM50== "Basal")
          bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
          bsl$status= as.numeric(as.character(bsl$status))
   }
   else { #if(analysis_mode== "RmTopGenes")
     tmp$type="RmTopGenes"
     
     if(j== nrow(sig))  next
     else {
     sig= sig[order(sig$avg_log2FC, decreasing = TRUE),]
     rmgenes= sig %>% top_n(n = j, wt = avg_log2FC) 
     gene.sets= list("p16"= toupper(sig[-which(sig$gene %in% rmgenes$gene),1]))
       dt= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
     dt$sample= rownames(dt)
     dt=dt[rownames(clin),]
     stopifnot(identical(dt$sample, rownames(clin)))
     md=cbind(clin, dt)
     bsl= md %>% filter(PAM50== "Basal")
      bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
     bsl$status= as.numeric(as.character(bsl$status))}
     
   }

for(i in seq(0.1, 0.9, 0.01)){
 
    bsl$p16discrete=  ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, i), 1, 0 )
  res <- pairwise_survdiff(Surv(overall_survival,status) ~ p16discrete, data = bsl)
  qtl2$quantile= append(qtl2$quantile, i)
  qtl2$min_Pvalue= append(qtl2$min_Pvalue, min(na.omit(res$p.value)))
}
  dt_b=do.call( rbind, qtl2) %>% t() %>%  as.data.frame()
  dt_b=dt_b[-1,]

bestpv= dt_b[order(dt_b$min_Pvalue, decreasing = FALSE),2][1]
bestqtl= dt_b[order(dt_b$min_Pvalue, decreasing = FALSE),1][1]
print(bestpv)
print(bestqtl)
tmp[j,3]= dt_b[order(dt_b$min_Pvalue, decreasing = FALSE),2][1]
tmp[j,2]= dt_b[order(dt_b$min_Pvalue, decreasing = FALSE),1][1]


#tmp[j,3]= bestpv
#tmp[j,2]=bestqtl
bsl$p16discrete= ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, bestqtl), "high", "low" )

fit <- survfit(Surv(overall_survival,status) ~ p16discrete, data = bsl)

pdf(paste0(testDir2, "/Survival_OS_", tmp$type,j,"_BestQtl_", bestqtl,".pdf"))
 p=ggsurvplot(fit,
          pval = TRUE, conf.int = FALSE,
          risk.table = FALSE, 
          risk.table.col = "strata",
          linetype = "strata", 
          surv.median.line = "hv", 
          ylab=paste0("OS_", bestqtl, tmp$type,j),
          ggtheme = theme_bw(), 
          palette = c("#DED5D5", "#262424", "#756F6F")) 
 print(p)
 dev.off()
 
print(j) 
rm(dt)
}
return(tmp)  
}   
   
mel= test_function_CPTAC(analysis_mode = "TopGenes", sig=sig)
```

Include the low expressed genes:
```{r}
sig= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = FALSE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05)
sig$direction= ifelse(sig$avg_log2FC > 0, paste0(sig$gene,"+"), paste0(sig$gene,"-"))
sig$direction=toupper(sig$direction)
```



ONLY NEGATIVE GENES:

USE MIN PVAL METHOD: WITH ORDERING GENES ACCORDING TO FOLD CHANGE:
```{r}
sigN= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = FALSE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05) %>% filter(avg_log2FC < 0)
##
sigP= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = TRUE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05)
###
sig= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = FALSE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05)
#############################################################################################
tmpN= data.frame("topGenes"= c(1,seq(5,nrow(sigN), 5), 74), "Best_Qtl"= NA, "Pval_Min"=NA) #,1,seq(5, nrow(sigP), 5), nrow(sigP)
rownames(tmpN)= tmpN$topGenes

clin= cptac_clin %>% as.data.frame()


for(j in c(1,seq(5,nrow(sigN), 5), 74)){
  
  sigenes= sigN %>% top_n(n = abs(j), wt = abs(avg_log2FC)) 
  #sigenes$gene2= paste0(toupper(sigenes$gene), "-")
  gene.sets= list("p16"= toupper(sigenes$gene))
  dt= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
  dt$sample= rownames(dt)
  dt=dt[rownames(clin),]
  stopifnot(identical(dt$sample, rownames(clin)))
  md=cbind(clin, dt)

  bsl= md %>% filter(PAM50== "Basal")
   bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
bsl$status= as.numeric(as.character(bsl$status))

###################################################################
qtl=list("quantile"=NA, "hazard"=NA, "waldtest"=NA,"logTestPval"=NA,"waldtestPval"=NA)

for(i in seq(0.1, 0.9, 0.01)){
  bsl$p16discrete= ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, i), 1, 0)
   res.cox <- summary(coxph(Surv(overall_survival,status ) ~ p16discrete, data = bsl))
   
  qtl$quantile= append(qtl$quantile, i)
  qtl$hazard=append(qtl$hazard,res.cox$coefficients[,2])
  ##
 qtl$logTestPval= append(qtl$logTestPval, res.cox$logtest[3])
  qtl$waldtest= append(qtl$waldtest,res.cox$waldtest[1] )
   qtl$waldtestPval= append(qtl$waldtestPval,res.cox$waldtest[3] )
} 
 dt.qtl_binary= do.call(rbind, qtl) %>% t() %>%  as.data.frame()
dt.qtl_binary=dt.qtl_binary[-1,]


bestqtl=  dt.qtl_binary[order( dt.qtl_binary$logTestPval, decreasing = FALSE),1][1]
bestpval=  dt.qtl_binary[order( dt.qtl_binary$logTestPval, decreasing = FALSE),4][1]

tmpN[j,2]= bestqtl
tmpN[j,3]=bestpval
####
bsl$p16discrete= ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, bestqtl), "high", "low" )

library(survival)

fit <- survfit(Surv(overall_survival,status) ~ p16discrete, data = bsl)
pdf(paste0(sigDir, "/SurvivalPlot_CPTAC_Stratif", "TopNegativeFC_",j,"_BestQtl_", bestqtl,".pdf"))
 p=ggsurvplot(fit,
          pval = TRUE, conf.int = FALSE,
          risk.table = FALSE, 
          risk.table.col = "strata",
          linetype = "strata", 
          surv.median.line = "hv", 
          ylab=paste0("OS_", bestqtl, "Top_Negative_",j),
          ggtheme = theme_bw(), 
          palette = c("#DED5D5", "#262424", "#756F6F")) 
 print(p)
 dev.off()
}
tmpN2= tmpN

tmpN2=tmpN2[c(1,seq(5,nrow(sigN), 5), 74),]
tmpN2$topGenes=c(1,seq(5,nrow(sigN), 5), 74)

pdf(paste0(sigDir, "/Plot_Best_Top_Negative_Genes_p16Signature.pdf"))

m <- tmpN2 %>% ggplot(aes(x= topGenes, y= Pval_Min)) + geom_point() + geom_line() + themplot + xlab("Top_FCNeg_Genes")

n <- tmpN2 %>% ggplot(aes(x= topGenes, y= Best_Qtl)) + geom_point() + geom_line() + themplot+ xlab("Top_FCNeg_Genes")

print(m)
print(n)
dev.off()

writexl::write_xlsx(tmpN2, paste0(sigDir, "/Table_Best_TopNeg_Genes_P16Signature.xlsx"))
```

compile datasets:
```{r}
tmpN$type= "NegFC_Genes"
tmpRM$topGenes= paste0("RM_Top_", tmpRM$topGenes)
dt= rbind(tmpN, tmpT, tmp)
##################
tmpT$type= "TopGenes"
tmp$type="Random"
dt= rbind(tmpN,tmp, tmpT)
writexl::write_xlsx(dt, paste0(testDir, "/DATASET_RandomSel_TopG_TopGNeg.xlsx"))

pdf(paste0(testDir, "/Lines_Random_TopGenes_TopNegGenes_MinPval_BestQtl.pdf"))

m <- dt %>% ggplot(aes(x= topGenes, y= Pval_Min, color= type)) + geom_point() + geom_line() + themplot

n <- dt %>% ggplot(aes(x= topGenes, y= Best_Qtl, color= type)) + geom_point() + geom_line() + themplot

print(m)
print(n)
dev.off()

```
POSITIVE GENES :

```{r}
tmpP= data.frame("topGenes"= c(1,seq(5,nrow(sigP), 5), 39), "Best_Qtl"= NA, "Pval_Min"=NA) #,1,seq(5, nrow(sigP), 5), nrow(sigP)
rownames(tmpP)= tmpP$topGenes

clin= cptac_clin %>% as.data.frame()


for(j in c(1,seq(5,nrow(sigP), 5), 39)){
  
  sigenes= sigP %>% top_n(n = abs(j), wt = abs(avg_log2FC)) 
  #sigenes$gene2= paste0(toupper(sigenes$gene), "-")
  gene.sets= list("p16"= toupper(sigenes$gene))
  dt= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
  dt$sample= rownames(dt)
  dt=dt[rownames(clin),]
  stopifnot(identical(dt$sample, rownames(clin)))
  md=cbind(clin, dt)

  bsl= md %>% filter(PAM50== "Basal")
   bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
bsl$status= as.numeric(as.character(bsl$status))

###################################################################
qtl=list("quantile"=NA, "hazard"=NA, "waldtest"=NA,"logTestPval"=NA,"waldtestPval"=NA)

for(i in seq(0.1, 0.9, 0.01)){
  bsl$p16discrete= ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, i), 1, 0)
   res.cox <- summary(coxph(Surv(overall_survival,status ) ~ p16discrete, data = bsl))
   
  qtl$quantile= append(qtl$quantile, i)
  qtl$hazard=append(qtl$hazard,res.cox$coefficients[,2])
  ##
 qtl$logTestPval= append(qtl$logTestPval, res.cox$logtest[3])
  qtl$waldtest= append(qtl$waldtest,res.cox$waldtest[1] )
   qtl$waldtestPval= append(qtl$waldtestPval,res.cox$waldtest[3] )
} 
 dt.qtl_binary= do.call(rbind, qtl) %>% t() %>%  as.data.frame()
dt.qtl_binary=dt.qtl_binary[-1,]


bestqtl=  dt.qtl_binary[order( dt.qtl_binary$logTestPval, decreasing = FALSE),1][1]
bestpval=  dt.qtl_binary[order( dt.qtl_binary$logTestPval, decreasing = FALSE),4][1]

tmpP[j,2]= bestqtl
tmpP[j,3]=bestpval
####
bsl$p16discrete= ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, bestqtl), "high", "low" )

library(survival)

fit <- survfit(Surv(overall_survival,status) ~ p16discrete, data = bsl)
pdf(paste0(sigDir, "/SurvivalPlot_CPTAC_Stratif", "TopPositiveFC_",j,"_BestQtl_", bestqtl,".pdf"))
 p=ggsurvplot(fit,
          pval = TRUE, conf.int = FALSE,
          risk.table = FALSE, 
          risk.table.col = "strata",
          linetype = "strata", 
          surv.median.line = "hv", 
          ylab=paste0("OS_", bestqtl, "Top_Positive_",j),
          ggtheme = theme_bw(), 
          palette = c("#DED5D5", "#262424", "#756F6F")) 
 print(p)
 dev.off()
}
tmpP2=tmpP
tmpP= tmpP[c(1,seq(5,nrow(sigP), 5), 39),]
tmpP$topGenes=c(1,seq(5,nrow(sigP), 5), 39)
###

 ###
 pdf(paste0(sigDir, "/Barplot_BestParams_TopPositiveGenes.pdf"))
p <- tmpP %>% ggplot( aes(x= as.factor(topGenes), y= Best_Qtl)) + geom_bar(stat="identity") + themplot

g <- tmpP %>% ggplot( aes(x= as.factor(topGenes), y= Pval_Min)) + geom_bar(stat="identity") + themplot
print(p)
print(g)
dev.off()
#####################

pdf(paste0(sigDir, "/Lines_BestParams_TopPositivGenes.pdf"))

m <- tmpP %>% ggplot(aes(x= topGenes, y= Pval_Min)) + geom_point() + geom_line() + themplot + xlab("Top_Positive")

n <- tmpP %>% ggplot(aes(x= topGenes, y= Best_Qtl)) + geom_point() + geom_line() + themplot + xlab("Top_Positive")

print(m)
print(n)
dev.off()
```


Survival plots with 
```{r}
bsl2= bsl[which(rownames(bsl) %in% rownames(sig.resCPTAC)),]
identical(rownames(bsl2), rownames(sig.resCPTAC))

sig.resCPTAC$overall_survival= as.numeric(as.character(bsl2$overall_survival))
sig.resCPTAC$status= as.numeric(as.character(bsl2$status))


tmpP= data.frame("LengthGenes"= colnames(sig.resCPTAC)[1:16], "Best_Qtl"= NA, "Pval_Min"=NA) 
rownames(tmpP)= colnames(sig.resCPTAC)[1:16]

for(j in seq(1,16)){
  
  qtl=list("quantile"=NA, "hazard"=NA, "waldtest"=NA,"logTestPval"=NA,"waldtestPval"=NA)

for(i in seq(0.1, 0.9, 0.01)){
  sig.resCPTAC$p16discrete= ifelse(sig.resCPTAC[,j] > quantile(sig.resCPTAC[,j], i), 1, 0)
   res.cox <- summary(coxph(Surv(overall_survival,status ) ~ p16discrete, data = sig.resCPTAC))
  qtl$quantile= append(qtl$quantile, i)
  qtl$hazard=append(qtl$hazard,res.cox$coefficients[,2])

 qtl$logTestPval= append(qtl$logTestPval, res.cox$logtest[3])
  qtl$waldtest= append(qtl$waldtest,res.cox$waldtest[1] )
   qtl$waldtestPval= append(qtl$waldtestPval,res.cox$waldtest[3] )
} 
 
    
 dt.qtl_binary= do.call(rbind, qtl) %>% t() %>%  as.data.frame()
dt.qtl_binary=dt.qtl_binary[-1,]


bestqtl=  dt.qtl_binary[order( dt.qtl_binary$waldtestPval, decreasing = FALSE),1][1]
bestpval=  dt.qtl_binary[order( dt.qtl_binary$waldtestPval, decreasing = FALSE),5][1]
####
tmpP[j,2]=bestqtl

tmpP[j,3]=bestpval

sig.resCPTAC$p16discrete= ifelse(sig.resCPTAC[,j] > quantile(sig.resCPTAC[,j], bestqtl), "high", "low" )

library(survival)

fit <- survfit(Surv(overall_survival,status) ~ p16discrete, data = sig.resCPTAC)

pdf(paste0(outputDir2,"SurvivalPlot_CPTAC_Signature_",tmpP[j,1],"_BestQtl_", bestqtl,"_WaldTest.pdf"), height = 5, width = 5)
 p=ggsurvplot(fit,
          pval = TRUE, conf.int = FALSE,
          risk.table = FALSE, 
          risk.table.col = "strata",
          linetype = "strata", 
          surv.median.line = "hv", 
          ylab=paste0("OS_", bestqtl, "VarImp_",j),
          ggtheme = theme_bw(), 
          palette = c("#DED5D5", "#262424", "#756F6F")) 
 print(p)
 dev.off()
}

writexl::write_xlsx(tmpP, paste0(outputDir2, "CPTAC_RFE_Survival_WaldTest.xlsx"))
#------------------
tmp_Loglike=tmpP

tmpP$LengthGenes= gsub("p16s_", "", tmpP$LengthGenes)
tmpP$LengthGenes=gsub("g_UCell","", tmpP$LengthGenes)

tmpP$LengthGenes=as.numeric(as.character(tmpP$LengthGenes))

pdf(paste0(outputDir2, "Lines_CPTAC_BestParams_VarImp_WaldTest.pdf"))
m <- tmpP %>% ggplot(aes(x= LengthGenes, y= Pval_Min,group = 1))  + geom_line() + geom_point()+ themplot + xlab("Top_VarImp_RFE_CPTAC")

n <- tmpP %>% ggplot(aes(x= LengthGenes, y= Best_Qtl,group = 1)) + geom_point() + geom_line() + themplot + xlab("Top_VarImp_RFE_CPTAC")

print(m)
print(n)
dev.off()

```


for the paper:
```{r}

#P16Low should be tn$p16Up < quantile(tn$p16Up, 0.25)
#tmpP= data.frame("LengthGenes"= colnames(sig.res)[1:16], "Best_Qtl"= NA, "Pval_Min"=NA) #,1,seq(5, nrow(sigP), 5), nrow(sigP)
#rownames(tmpP)= colnames(sig.res)[1:16]
#i=0
#for(j in seq(1,16)){
  
  qtl=list("quantile"=NA, "hazard"=NA, "waldtest"=NA,"logTestPval"=NA,"waldtestPval"=NA)

for(i in seq(0.1, 0.9, 0.01)){
  sig.resCPTAC$p16discrete= ifelse(sig.resCPTAC$p16s_13g_UCell > quantile(sig.resCPTAC$p16s_13g_UCell, i), 1, 0)
  
   res.cox <- summary(coxph(Surv(overall_survival,status ) ~ p16discrete, data = sig.resCPTAC))
   
  qtl$quantile= append(qtl$quantile, i)
  qtl$hazard=append(qtl$hazard,res.cox$coefficients[,2])
  ##
 qtl$logTestPval= append(qtl$logTestPval, res.cox$logtest[3])
  qtl$waldtest= append(qtl$waldtest,res.cox$waldtest[1] )
   qtl$waldtestPval= append(qtl$waldtestPval,res.cox$waldtest[3] )
} 
  
  
 dt.qtl_binary= do.call(rbind, qtl) %>% t() %>%  as.data.frame()
dt.qtl_binary=dt.qtl_binary[-1,]

qs::qsave(dt.qtl_binary, paste0(outputDir2, "/CPTAC_Best_Qtl_preHumanSignature_13genes.qs"))



pdf(paste0( "C:/Users/Melissa S/Desktop/BestQtl_Pvalue_preHuman13gene_WaldTest_CPTAC.pdf"))
m <- dt.qtl_binary %>% ggplot(aes(x= quantile, y= waldtestPval)) + geom_bar(stat="identity") + themplot
print(m)
dev.off()

```







Plot the stage:
```{r}
unique(bsl$pathologic_stage)
sig.resCPTAC$stage= dplyr::recode(bsl2$pathologic_stage, "stageii"= "2", "stagei"= "1", "stageiii"= "3", "stageiv"="4")

pdf(paste0(outputDir2, "BoxPlot_13genes_Stage_CPTAC.pdf"))
p <- ggplot(data= sig.resCPTAC, aes(x= stage, y= p16s_13g_UCell)) +  geom_boxplot(fill="grey")  + themplot + geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.4, size=2.5)+   stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "1",aes(label=..p.adj..), alternative="greater")       

print(p)
p <- ggplot(data= sig.resCPTAC, aes(x= stage, y= p16s_38g_UCell)) + geom_boxplot()+ geom_boxplot(fill="grey")  + themplot + geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.4, size=2.5) + ggtitle("Important_10genes_CPTAC") +   stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "1",aes(label=..p.adj..), alternative="greater")
print(p)
dev.off()

sig.resCPTAC$binary_stage= ifelse(sig.resCPTAC$stage=="1", "early", "late")

p <- ggplot(data= sig.resCPTAC, aes(x= binary_stage, y= p16s_13g_UCell)) + geom_boxplot()  + themplot + geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.5) + ggtitle("Important_13genes_CPTAC") +   stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "early",aes(label=..p.adj..), alternative="greater")   #p-value = 0.0478 : greater  
```


All the boxplot :
```{r}
p16list13= list(p16s_13g=varimp_data$feature[1:13])

library(data.table)
library(UCell)
cptac.dt=as.data.frame(cptac.dt)

sig.resCPTAC2= UCell::ScoreSignatures_UCell(matrix = ge, features= p16list13, maxRank = 15000) %>% as.data.frame()

sig.resCPTAC2$sample=rownames(sig.resCPTAC2)
identical(rownames(sig.resCPTAC2), rownames(cptac_clin))
sig.resCPTAC2=sig.resCPTAC2[rownames(cptac_clin),]
sig.resCPTAC2$type= cptac_clin$PAM50
#-----------------------------#

pdf(paste0(outputDir2, "BoxPlot_Type_PAM50_CPTAC_PAM50.pdf"))
p <- ggplot(data= sig.resCPTAC2, aes(x= type, y= p16s_13g_UCell)) + geom_boxplot()+ geom_boxplot(fill="grey")  + themplot + geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.4, size=2.5) + ggtitle("Important_13genes_CPTAC") +   stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Basal",aes(label=..p.adj..))   
print(p)
dev.off()

```




















POOL POSITIVE WITH NEGATIVE VALUES:

```{r}
sig$gene2= ifelse(sig$avg_log2FC > 0, paste0(toupper(sig$gene),"+"), paste0(toupper(sig$gene),"-"))
  gene.sets= list("p16"= sig$gene2)
  dt= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
```

```{r}
sigN= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = FALSE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05) %>% filter(avg_log2FC < 0)
##
sigP= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = TRUE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05)
###
sig= FindMarkers(seu, ident.1 = "P16+ Pre-lesional", ident.2 = c("LP", "Avd"), only.pos = FALSE, logfc.threshold = 0.8, min.diff.pct = 0.2)  %>% tibble::rownames_to_column("gene") %>% dplyr::filter(p_val_adj < 0.05)
#############################################################################################


tmpB= data.frame("topGenes"= c(1,seq(5, nrow(sigP), 5), nrow(sigP)), "Best_Qtl"= NA, "Pval_Min"=NA) 
rownames(tmpB)= tmpB$topGenes

j=0
for(j in c(1,seq(5, nrow(sigP), 5), nrow(sigP))){
  
  cN= sigN %>% top_n(n = -j, wt = avg_log2FC) 
  cN$gene2= paste0(toupper(cN$gene), "-")
  ##
  cP= sigP %>% top_n(n = j, wt = avg_log2FC) 
  cP$gene2= paste0(toupper(cP$gene), "+")
  ##
  
  gene.sets= list("p16"= c(cP$gene2, cN$gene2))
  
  dt= ScoreSignatures_UCell(genexp[,-1], features= gene.sets) %>% as.data.frame()
  dt$sample= rownames(dt)
  dt=dt[rownames(clin),]
  stopifnot(identical(dt$sample, rownames(clin)))
  md=cbind(clin, dt)

  bsl= md %>% filter(PAM50== "Basal")
   bsl$overall_survival= as.numeric(as.character(bsl$overall_survival))
bsl$status= as.numeric(as.character(bsl$status))

###################################################################
qtl=list("quantile"=NA, "hazard"=NA, "waldtest"=NA,"logTestPval"=NA,"waldtestPval"=NA)

for(i in seq(0.1, 0.9, 0.01)){
  bsl$p16discrete= ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, i), 1, 0)
   res.cox <- summary(coxph(Surv(overall_survival,status ) ~ p16discrete, data = bsl))
   
  qtl$quantile= append(qtl$quantile, i)
  qtl$hazard=append(qtl$hazard,res.cox$coefficients[,2])
  ##
 qtl$logTestPval= append(qtl$logTestPval, res.cox$logtest[3])
  qtl$waldtest= append(qtl$waldtest,res.cox$waldtest[1] )
   qtl$waldtestPval= append(qtl$waldtestPval,res.cox$waldtest[3] )
} 
 dt.qtl_binary= do.call(rbind, qtl) %>% t() %>%  as.data.frame()
dt.qtl_binary=dt.qtl_binary[-1,]


bestqtl=  dt.qtl_binary[order( dt.qtl_binary$logTestPval, decreasing = FALSE),1][1]
bestpval=  dt.qtl_binary[order( dt.qtl_binary$logTestPval, decreasing = FALSE),4][1]

tmpB[j,2]= bestqtl
tmpB[j,3]=bestpval
####
bsl$p16discrete= ifelse(bsl$p16_UCell > quantile(bsl$p16_UCell, bestqtl), "high", "low" )

library(survival)

fit <- survfit(Surv(overall_survival,status) ~ p16discrete, data = bsl)
pdf(paste0(sigDir, "/SurvivalPlot_CPTAC_TopGenes_Compiled_", "Neg_Pos_",j,"_BestQtl_", bestqtl,".pdf"))
 p=ggsurvplot(fit,
          pval = TRUE, conf.int = FALSE,
          risk.table = FALSE, 
          risk.table.col = "strata",
          linetype = "strata", 
          surv.median.line = "hv", 
          ylab=paste0("OS_", bestqtl, "Top_PosNeg_",j),
          ggtheme = theme_bw(), 
          palette = c("#DED5D5", "#262424", "#756F6F")) 
 print(p)
 dev.off()
}
tmpB2=tmpB

tmpB=tmpB[ c(1,seq(5, nrow(sigP), 5), nrow(sigP)),]
tmpB$topGenes= c(1,seq(5, nrow(sigP), 5), nrow(sigP))

 writexl::write_xlsx(tmpB, paste0(sigDir, "/Table_PosNegTop_BestParams.xlsx"))
 ###
 pdf(paste0(sigDir, "/Barplot_BestParams_TopNegGenes.pdf"))
p <- tmpB %>% ggplot( aes(x= as.factor(topGenes), y= Best_Qtl)) + geom_bar(stat="identity") + themplot

g <- tmpB %>% ggplot( aes(x= as.factor(topGenes), y= Pval_Min)) + geom_bar(stat="identity") + themplot
print(p)
print(g)
dev.off()
#####################

pdf(paste0(sigDir, "/Lines_BestParams_TopNegGenes.pdf"))

m <- tmpB %>% ggplot(aes(x= topGenes, y= Pval_Min)) + geom_point() + geom_line() + themplot + xlab("Combination_Top_Positive_Negative")

n <- tmpB %>% ggplot(aes(x= topGenes, y= Best_Qtl)) + geom_point() + geom_line() + themplot + xlab("Combination_Top_Positive_Negative")

print(m)
print(n)
dev.off()
```




```{r}
writexl::write_xlsx(sig.resCPTAC, paste0(outputDir2, "CPTAC_SignatureHuman_Final.xlsx"))
sig.resCPTAC= readxl::read_excel(paste0(outputDir2, "CPTAC_SignatureHuman_Final.xlsx"))

writexl::write_xlsx(sig.res, paste0(outputDir2, "TCGA_SignatureHuman_Final.xlsx"))
```

